### 2012-05-08 paper.tex
Inital paper in taco format

% acmsmall-sample.tex, dated 4th Nov. 2011
% This is a sample file for ACM small trim journals
%
% Compilation using 'acmsmall.cls' - version 1.2, Aptara Inc.
% (c) 2010 Association for Computing Machinery (ACM)
%
% Questions/Suggestions/Feedback should be addressed to => "acmtexsupport@aptaracorp.com".
% Users can also go through the FAQs available on the journal's submission webpage.
%
% Steps to compile: latex, bibtex, latex latex
%
% For tracking purposes => this is v1.2 - Nov. 2011

\documentclass[prodmode,acmtaco]{acmsmall}

% Package to generate and customize Algorithm as per ACM style
\usepackage[ruled]{algorithm2e}
\renewcommand{\algorithmcfname}{ALGORITHM}
\SetAlFnt{\small}
\SetAlCapFnt{\small}
\SetAlCapNameFnt{\small}
\SetAlCapHSkip{0pt}
\IncMargin{-\parindent}

% Metadata Information
\acmVolume{9}
\acmNumber{4}
\acmArticle{39}
\acmYear{2010}
\acmMonth{3}

% Document starts
\begin{document}

% Page heads
\markboth{S. Muralidharan et al.}{Title goes here}

% Title portion
\title{Title goes here}
\author{
SERVESH MURALIDHARAN
\affil{Trinity College Dublin}
DAVID GREGG
\affil{Trinity College Dublin}}

\begin{abstract}

Abstract goes here

\end{abstract}

\category{C.2.2}{Computer-Communication Networks}{Network Protocols}

\terms{Design, Algorithms, Performance}

\keywords{}

\acmformat{Muralidharan, S., and Gregg, D.,  2012. Title goes here.}

\begin{bottomstuff}
This work is funded by the IRCSET Enterprise Partnership Scheme along with IBM Research Ireland.

Author's addresses: S. Muralidharan {and} D. Gregg, Computer Science Department,
Trinity College Dublin, Ireland.
\end{bottomstuff}

\maketitle

\section{Introduction}

The constant advancement in server hardware has made systems with large number of hardware threads[] readily available. In addition to this the presence of co-processors[], network processors[], Graphics Processing Units ( GPUs ) \& General Purpose GPUs ( GPGPUs ) has dramatically improved the capability of commodity hardware. These developments has paved way to a platform with immense computing potential which existing applications already take advantage of. Along with the improvements in the Multi-Core \& Many-Core processor architecture the memory and I/O speeds have also seen significant improvement. In particular the network subsystem has seen significant growth, in combination with the improvements to optical communication it has lead to 10GbE interfaces being quite common in enterprise servers. While 40GbE and 100GbE are available for backbone infrastructure but would eventually follow into commodity hardware.

Several previous work have discussed the process of parallelizing network applications[][][] to enable it to run faster in such multi threaded systems. The development of Software routers[][][] is encouraged further because it costs only a fraction of the commodity hardware routers. It has been shown that such routers can be used to process packets at line rate [][][] which has been a challenge on its own. Network applications including those that operate on few or all of the network layers have been designed with sequential processing as the target environment. The shift to multi core architectures in recent times has induced such applications to utilize parallel processing as a alternate means of computing. While this could lead to significant improvement in performance it also suffers from the same problems as any parallel application. These include scalability, load-balancing, data dependency and the presence of non parallel regions in the program apart from those that might arise due to that specific application. The scaling of software routers has been discussed extensively in RouteBricks[] which shows the problems suffered by software routers when they are scaled beyond a single system. One could argue that complexity of network applications such as Deep Packet Inspection ( DPI ), Packet Forwarding, etc., could be handled by a single many socket Multi-Core SMP system. However, as the complexity of network applications develops it would be inevitable but to move to a multi node environment in order to perform the computations and to handle more network bandwidth.

Large scale computing clusters has been used in high-performance scientific application to solve problems and applications of very large size. It uses a cluster of compute nodes and distributes the workload across them in order to solve it faster. These systems utilize parallel programs consisting of both threads and MPI to break massive problems into smaller workload that can be executed in parallel. In addition, the presence of GPUs and GPGPUs requires a mixture of parallelization constructs such as OpenCL, CUDA, etc., in order to achieve the best performance in the system. In this paper we discuss how some of the techniques used by these applications could be used to parallelize network applications.

Click Modular Software Router[] consists of extensive elements which are used to construct software routers. Click primarily targets packet processing applications and contains an extensive

Discuss about what each sections contains here.....

\section{Background}

Talk about click...

\section{Design}

Emphasize on user level execution environment

Eliminate the overhead associated with creation of dynamic objects and virtual function calls at run time

Optimize the code generated for a particular application graph

Eliminate the dependency of creating queue elements to implement task parallelism

Eliminate the creation of multiple graphs to implement data parallelism

\section{Implementation}

\section{Experimental Evaluation}

\section{Related Work}

\section{Conclusions}

% Acknowledgments
\begin{acks}
The authors would like to thank ...
\end{acks}

% Bibliography
\bibliographystyle{acmsmall}
\bibliography{}

% History dates
\received{June 2012}

\end{document}

### 2012-05-11 paper.tex
Minor corrections

@@
 
+Network applications represent those applications that operate on any of the seven network layers described in the OSI Model. \{describe more here\}
+
 Several previous work have discussed the process of parallelizing network applications[][][] to enable it to run faster in such multi threaded systems.
 The development of Software routers[][][] is encouraged further because it costs only a fraction of the commodity hardware routers.
-It has been shown that such routers can be used to process packets at line rate [][][] which has been a challenge on its own.
+Moreover the advancements in the network interfaces has made faster interfaces readily in combination with the multi core servers the performance  It has been shown that such routers can be used to process packets at line rate [][][] which has been a challenge on its own.
 Network applications including those that operate on few or all of the network layers have been designed with sequential processing as the target environment.
@@
 
-Click Modular Software Router[] consists of extensive elements which are used to construct software routers. Click primarily targets packet processing applications and contains an extensive
+Click Modular Software Router[] consists of extensive elements which are used to construct software routers.
+Click primarily targets packet processing applications especially those that operate on L4 - L7.
+\{elaborate more here\}
 
@@
 
-Talk about click...
+Talk about click...\\
+
+Describe about what would be the best way to parallelize using threads and MPI.. provide reference.. \\
 
@@
 
-Emphasize on user level execution environment
+This section describes the design principles related to our methodology.
+The objective is to determine the requirements necessary to parallelize network applications.
+Our methods focus on ways to improve the click system in some cases even extend its functionality.
 
-Eliminate the overhead associated with creation of dynamic objects and virtual function calls at run time
+\subsection{Emphasize on user level execution environment}
 
-Optimize the code generated for a particular application graph
+The existing software routers are predominantly implemented in the kernel space for the ease of access to resources and can bypass the OS overhead incurred when data is accessed through the network subsystem.
+While this alternative provides better performance it also leads to concerns regarding stability and security of the system.
+Newer techniques such as Netmap[] allows direct access to network buffers which makes it possible to write high performance packet processing applications in userspace itself.
+Typically network applications which operate on L4 - L7 are developed in the kernel space in order to intercept the network packets before it flows through the OS.
+Whereas those applications that utilize L1 - L3 are present in the user space.
+The objective here is to completely switch to user space to solve any of the stability or security concerns.
+In addition, scenarios where the applications that operate across most or all the layers could be improved by combining the operations across the different layers in user space itself.
 
-Eliminate the dependency of creating queue elements to implement task parallelism
+\subsection{Eliminate the overhead associated with creation of dynamic objects}
 
-Eliminate the creation of multiple graphs to implement data parallelism
+Click constructs application graphs based on dynamic creation of objects and passing the data packets across the virtual functions.
+Though this is a very elegant approach in creating the software router it does incur the penalty of the virtual function tables.
+The click-tool[] addresses this issue but it does not solve the problem entirely.
+The construction of the software router at run time is the primary reason behind it.
+This mainly arises due to the presence of kernel module which installs the router as a kernel module based on the input configuration.
+However in the user level it would not be necessary for a separate process to initialize the router instead we could generate the entire program statically including the necessary flow associated with the specific application.
+This not only eliminates the overhead associated but also opens up means to optimize the graph further.
+
+\subsection{Optimize based on application graph}
+
+Network applications especially that operate on the lower layers are very sensitive to the overhead associated with operating on the data separately on each layer.
+Instead given an application graph if we were to provide a sequential flow of the different functions at the various layers then it would be possible for the compiler to optimize the regions of program into a single and optimized block of code.
+This would also improve the cache locality of the data accessed by consecutive functions.
+This also presents us with the ability to perform other types of stream graph optimizations[] which would not be possible if the specific graph was constructed dynamically at run time.
+
+\subsection{Automated data parallelism}
+
+Network traffic is a classic example by which data parallelism can be utilized to achieve scalable parallelism.
+Click's support for parallelism in user level is described as experimental, it is implemented by using a task graph consisting of each work associated with each element which is executed in the threads.
+The support for parallelism in the kernel mode is better in comparison to that of user mode.
+Since our target environment is in the user space the proposed design involves a different approach to which we can extract data parallelism.
+We first split the data across multiple queues and then replicate the application graph across to process the data the queues in parallel.
+An important condition here is to make sure packets of the same flow should pass through the same queue.
+Since the application graph can be replicated it also makes it possible to dynamically load balance several applications based on the workload.
+
+\subsection{Efficient implementation of task parallelism}
+
+In generating the application graph as a sequence of operations it also makes it possible to process different stages of the operation in a pipelined manner.
+This could be combined with the stream optimizations and could be used to determine which steps of the graph is best suited to form the stages of a pipeline.
+The condition to balance here is the weight associated with computation in that stage with that of the overhead incurred in computing the data in two separate stages.
+It would be necessary to determine whether the additional over head involved in executing the task in parallel does provide any significant improvement in performance.
 
 \section{Implementation}
+
+--Click was designed as a means to easily implement a software router on top of a Linux OS.
+Click achieves this by implementing it in two different ways one is by installing a kernel module and the other is a program that executes in the userlevel.
+\\
+
+--Talk about MPI \& threads\\
+
+--Our goal here is to focus on the network application as a workload and characterize it as such without involving any kind of overhead that might occur due to the network.\\
 

### 2012-05-11 script.sh

@@ -1,0 +1,4 @@
+#!/bin/sh
+while inotifywait paper.tex; do
+    pdflatex -interaction=nonstopmode paper.tex
+done

### 2012-05-21 paper.tex
Editted the paper for IEEE Infocom '13 format

@@
-% acmsmall-sample.tex, dated 4th Nov. 2011
-% This is a sample file for ACM small trim journals
+%% bare_conf.tex
+%% V1.3
+%% 2007/01/11
+%% by Michael Shell
+%% See:
+%% http://www.michaelshell.org/
+%% for current contact information.
+%%
+%% This is a skeleton file demonstrating the use of IEEEtran.cls
+%% (requires IEEEtran.cls version 1.7 or later) with an IEEE conference paper.
+%%
+%% Support sites:
+%% http://www.michaelshell.org/tex/ieeetran/
+%% http://www.ctan.org/tex-archive/macros/latex/contrib/IEEEtran/
+%% and
+%% http://www.ieee.org/
+
+%%*************************************************************************
+%% Legal Notice:
+%% This code is offered as-is without any warranty either expressed or
+%% implied; without even the implied warranty of MERCHANTABILITY or
+%% FITNESS FOR A PARTICULAR PURPOSE! 
+%% User assumes all risk.
+%% In no event shall IEEE or any contributor to this code be liable for
+%% any damages or losses, including, but not limited to, incidental,
+%% consequential, or any other damages, resulting from the use or misuse
+%% of any information contained here.
+%%
+%% All comments are the opinions of their respective authors and are not
+%% necessarily endorsed by the IEEE.
+%%
+%% This work is distributed under the LaTeX Project Public License (LPPL)
+%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
+%% distributed and modified. A copy of the LPPL, version 1.3, is included
+%% in the base LaTeX documentation of all distributions of LaTeX released
+%% 2003/12/01 or later.
+%% Retain all contribution notices and credits.
+%% ** Modified files should be clearly indicated as such, including  **
+%% ** renaming them and changing author support contact information. **
+%%
+%% File list of work: IEEEtran.cls, IEEEtran_HOWTO.pdf, bare_adv.tex,
+%%                    bare_conf.tex, bare_jrnl.tex, bare_jrnl_compsoc.tex
+%%*************************************************************************
+
+% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
+% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
+% *** with production work. IEEE's font choices can trigger bugs that do  ***
+% *** not appear when using other class files.                            ***
+% The testflow support page is at:
+% http://www.michaelshell.org/tex/testflow/
+
+% Note that the a4paper option is mainly intended so that authors in
+% countries using A4 can easily print to A4 and see how their papers will
+% look in print - the typesetting of the document will not typically be
+% affected with changes in paper size (but the bottom and side margins will).
+% Use the testflow package mentioned above to verify correct handling of
+% both paper sizes by the user's LaTeX system.
 %
-% Compilation using 'acmsmall.cls' - version 1.2, Aptara Inc.
-% (c) 2010 Association for Computing Machinery (ACM)
+% Also note that the "draftcls" or "draftclsnofoot", not "draft", option
+% should be used if it is desired that the figures are to be displayed in
+% draft mode.
 %
-% Questions/Suggestions/Feedback should be addressed to => "acmtexsupport@aptaracorp.com".
-% Users can also go through the FAQs available on the journal's submission webpage.
+\documentclass[conference]{IEEEtran}
+%% INFOCOM 2013 addition:
+\makeatletter
+\def\ps@headings{%
+\def\@oddhead{\mbox{}\scriptsize\rightmark \hfil \thepage}%
+\def\@evenhead{\scriptsize\thepage \hfil \leftmark\mbox{}}%
+\def\@oddfoot{}%
+\def\@evenfoot{}}
+\makeatother
+\pagestyle{headings}
+% Add the compsoc option for Computer Society conferences.
 %
-% Steps to compile: latex, bibtex, latex latex
+% If IEEEtran.cls has not been installed into the LaTeX system files,
+% manually specify the path to it like:
+% \documentclass[conference]{../sty/IEEEtran}
+
+% Some very useful LaTeX packages include:
+% (uncomment the ones you want to load)
+
+% *** MISC UTILITY PACKAGES ***
 %
-% For tracking purposes => this is v1.2 - Nov. 2011
+%\usepackage{ifpdf}
+% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
+% compilation based on whether the output is pdf or dvi.
+% usage:
+% \ifpdf
+%   % pdf code
+% \else
+%   % dvi code
+% \fi
+% The latest version of ifpdf.sty can be obtained from:
+% http://www.ctan.org/tex-archive/macros/latex/contrib/oberdiek/
+% Also, note that IEEEtran.cls V1.7 and later provides a builtin
+% \ifCLASSINFOpdf conditional that works the same way.
+% When switching from latex to pdflatex and vice-versa, the compiler may
+% have to be run twice to clear warning/error messages.
 
-\documentclass[prodmode,acmtaco]{acmsmall}
+% *** CITATION PACKAGES ***
+%
+%\usepackage{cite}
+% cite.sty was written by Donald Arseneau
+% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
+% \cite{} output to follow that of IEEE. Loading the cite package will
+% result in citation numbers being automatically sorted and properly
+% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
+% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
+% \cite will automatically add leading space, if needed. Use cite.sty's
+% noadjust option (cite.sty V3.8 and later) if you want to turn this off.
+% cite.sty is already installed on most LaTeX systems. Be sure and use
+% version 4.0 (2003-05-27) and later if using hyperref.sty. cite.sty does
+% not currently provide for hyperlinked citations.
+% The latest version can be obtained at:
+% http://www.ctan.org/tex-archive/macros/latex/contrib/cite/
+% The documentation is contained in the cite.sty file itself.
 
-% Package to generate and customize Algorithm as per ACM style
-\usepackage[ruled]{algorithm2e}
-\renewcommand{\algorithmcfname}{ALGORITHM}
-\SetAlFnt{\small}
-\SetAlCapFnt{\small}
-\SetAlCapNameFnt{\small}
-\SetAlCapHSkip{0pt}
-\IncMargin{-\parindent}
+% *** GRAPHICS RELATED PACKAGES ***
+%
+\ifCLASSINFOpdf
+  % \usepackage[pdftex]{graphicx}
+  % declare the path(s) where your graphic files are
+  % \graphicspath{{../pdf/}{../jpeg/}}
+  % and their extensions so you won't have to specify these with
+  % every instance of \includegraphics
+  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
+\else
+  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
+  % will default to the driver specified in the system graphics.cfg if no
+  % driver is specified.
+  % \usepackage[dvips]{graphicx}
+  % declare the path(s) where your graphic files are
+  % \graphicspath{{../eps/}}
+  % and their extensions so you won't have to specify these with
+  % every instance of \includegraphics
+  % \DeclareGraphicsExtensions{.eps}
+\fi
+% graphicx was written by David Carlisle and Sebastian Rahtz. It is
+% required if you want graphics, photos, etc. graphicx.sty is already
+% installed on most LaTeX systems. The latest version and documentation can
+% be obtained at: 
+% http://www.ctan.org/tex-archive/macros/latex/required/graphics/
+% Another good source of documentation is "Using Imported Graphics in
+% LaTeX2e" by Keith Reckdahl which can be found as epslatex.ps or
+% epslatex.pdf at: http://www.ctan.org/tex-archive/info/
+%
+% latex, and pdflatex in dvi mode, support graphics in encapsulated
+% postscript (.eps) format. pdflatex in pdf mode supports graphics
+% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
+% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
+% not a bitmapped formats (.jpeg, .png). IEEE frowns on bitmapped formats
+% which can result in "jaggedy"/blurry rendering of lines and letters as
+% well as large increases in file sizes.
+%
+% You can find documentation about the pdfTeX application at:
+% http://www.tug.org/applications/pdftex
 
-% Metadata Information
-\acmVolume{9}
-\acmNumber{4}
-\acmArticle{39}
-\acmYear{2010}
-\acmMonth{3}
+% *** MATH PACKAGES ***
+%
+%\usepackage[cmex10]{amsmath}
+% A popular package from the American Mathematical Society that provides
+% many useful and powerful commands for dealing with mathematics. If using
+% it, be sure to load this package with the cmex10 option to ensure that
+% only type 1 fonts will utilized at all point sizes. Without this option,
+% it is possible that some math symbols, particularly those within
+% footnotes, will be rendered in bitmap form which will result in a
+% document that can not be IEEE Xplore compliant!
+%
+% Also, note that the amsmath package sets \interdisplaylinepenalty to 10000
+% thus preventing page breaks from occurring within multiline equations. Use:
+%\interdisplaylinepenalty=2500
+% after loading amsmath to restore such page breaks as IEEEtran.cls normally
+% does. amsmath.sty is already installed on most LaTeX systems. The latest
+% version and documentation can be obtained at:
+% http://www.ctan.org/tex-archive/macros/latex/required/amslatex/math/
 
-% Document starts
+% *** SPECIALIZED LIST PACKAGES ***
+%
+%\usepackage{algorithmic}
+% algorithmic.sty was written by Peter Williams and Rogerio Brito.
+% This package provides an algorithmic environment fo describing algorithms.
+% You can use the algorithmic environment in-text or within a figure
+% environment to provide for a floating algorithm. Do NOT use the algorithm
+% floating environment provided by algorithm.sty (by the same authors) or
+% algorithm2e.sty (by Christophe Fiorio) as IEEE does not use dedicated
+% algorithm float types and packages that provide these will not provide
+% correct IEEE style captions. The latest version and documentation of
+% algorithmic.sty can be obtained at:
+% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithms/
+% There is also a support site at:
+% http://algorithms.berlios.de/index.html
+% Also of interest may be the (relatively newer and more customizable)
+% algorithmicx.sty package by Szasz Janos:
+% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithmicx/
+
+% *** ALIGNMENT PACKAGES ***
+%
+%\usepackage{array}
+% Frank Mittelbach's and David Carlisle's array.sty patches and improves
+% the standard LaTeX2e array and tabular environments to provide better
+% appearance and additional user controls. As the default LaTeX2e table
+% generation code is lacking to the point of almost being broken with
+% respect to the quality of the end results, all users are strongly
+% advised to use an enhanced (at the very least that provided by array.sty)
+% set of table tools. array.sty is already installed on most systems. The
+% latest version and documentation can be obtained at:
+% http://www.ctan.org/tex-archive/macros/latex/required/tools/
+
+%\usepackage{mdwmath}
+%\usepackage{mdwtab}
+% Also highly recommended is Mark Wooding's extremely powerful MDW tools,
+% especially mdwmath.sty and mdwtab.sty which are used to format equations
+% and tables, respectively. The MDWtools set is already installed on most
+% LaTeX systems. The lastest version and documentation is available at:
+% http://www.ctan.org/tex-archive/macros/latex/contrib/mdwtools/
+
+% IEEEtran contains the IEEEeqnarray family of commands that can be used to
+% generate multiline equations as well as matrices, tables, etc., of high
+% quality.
+
+%\usepackage{eqparbox}
+% Also of notable interest is Scott Pakin's eqparbox package for creating
+% (automatically sized) equal width boxes - aka "natural width parboxes".
+% Available at:
+% http://www.ctan.org/tex-archive/macros/latex/contrib/eqparbox/
+
+% *** SUBFIGURE PACKAGES ***
+%\usepackage[tight,footnotesize]{subfigure}
+% subfigure.sty was written by Steven Douglas Cochran. This package makes it
+% easy to put subfigures in your figures. e.g., "Figure 1a and 1b". For IEEE
+% work, it is a good idea to load it with the tight package option to reduce
+% the amount of white space around the subfigures. subfigure.sty is already
+% installed on most LaTeX systems. The latest version and documentation can
+% be obtained at:
+% http://www.ctan.org/tex-archive/obsolete/macros/latex/contrib/subfigure/
+% subfigure.sty has been superceeded by subfig.sty.
+
+%\usepackage[caption=false]{caption}
+%\usepackage[font=footnotesize]{subfig}
+% subfig.sty, also written by Steven Douglas Cochran, is the modern
+% replacement for subfigure.sty. However, subfig.sty requires and
+% automatically loads Axel Sommerfeldt's caption.sty which will override
+% IEEEtran.cls handling of captions and this will result in nonIEEE style
+% figure/table captions. To prevent this problem, be sure and preload
+% caption.sty with its "caption=false" package option. This is will preserve
+% IEEEtran.cls handing of captions. Version 1.3 (2005/06/28) and later 
+% (recommended due to many improvements over 1.2) of subfig.sty supports
+% the caption=false option directly:
+%\usepackage[caption=false,font=footnotesize]{subfig}
+%
+% The latest version and documentation can be obtained at:
+% http://www.ctan.org/tex-archive/macros/latex/contrib/subfig/
+% The latest version and documentation of caption.sty can be obtained at:
+% http://www.ctan.org/tex-archive/macros/latex/contrib/caption/
+
+% *** FLOAT PACKAGES ***
+%
+%\usepackage{fixltx2e}
+% fixltx2e, the successor to the earlier fix2col.sty, was written by
+% Frank Mittelbach and David Carlisle. This package corrects a few problems
+% in the LaTeX2e kernel, the most notable of which is that in current
+% LaTeX2e releases, the ordering of single and double column floats is not
+% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
+% single column figure to be placed prior to an earlier double column
+% figure. The latest version and documentation can be found at:
+% http://www.ctan.org/tex-archive/macros/latex/base/
+
+%\usepackage{stfloats}
+% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
+% the ability to do double column floats at the bottom of the page as well
+% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
+% LaTeX2e). It also provides a command:
+%\fnbelowfloat
+% to enable the placement of footnotes below bottom floats (the standard
+% LaTeX2e kernel puts them above bottom floats). This is an invasive package
+% which rewrites many portions of the LaTeX2e float routines. It may not work
+% with other packages that modify the LaTeX2e float routines. The latest
+% version and documentation can be obtained at:
+% http://www.ctan.org/tex-archive/macros/latex/contrib/sttools/
+% Documentation is contained in the stfloats.sty comments as well as in the
+% presfull.pdf file. Do not use the stfloats baselinefloat ability as IEEE
+% does not allow \baselineskip to stretch. Authors submitting work to the
+% IEEE should note that IEEE rarely uses double column equations and
+% that authors should try to avoid such use. Do not be tempted to use the
+% cuted.sty or midfloat.sty packages (also by Sigitas Tolusis) as IEEE does
+% not format its papers in such ways.
+
+% *** PDF, URL AND HYPERLINK PACKAGES ***
+%
+%\usepackage{url}
+% url.sty was written by Donald Arseneau. It provides better support for
+% handling and breaking URLs. url.sty is already installed on most LaTeX
+% systems. The latest version can be obtained at:
+% http://www.ctan.org/tex-archive/macros/latex/contrib/misc/
+% Read the url.sty source comments for usage information. Basically,
+% \url{my_url_here}.
+
+% *** Do not adjust lengths that control margins, column widths, etc. ***
+% *** Do not use packages that alter fonts (such as pslatex).         ***
+% There should be no need to do such things with IEEEtran.cls V1.6 and later.
+% (Unless specifically asked to do so by the journal or conference you plan
+% to submit to, of course. )
+
+% correct bad hyphenation here
+\hyphenation{op-tical net-works semi-conduc-tor}
+
 \begin{document}
+%
+% paper title
+% can use linebreaks \\ within to get better formatting as desired
+\title{Title goes here}
 
-% Page heads
-\markboth{S. Muralidharan et al.}{Title goes here}
+% author names and affiliations
+% use a multiple column layout for up to three different
+% affiliations
+\author{\IEEEauthorblockN{Servesh Muralidharan}
+\IEEEauthorblockA{School of Computer Science and\\Statistics\\
+Trinity College Dublin\\
+Dublin 2, Ireland\\
+Email: muralis@tcd.ie}
+\and
+\IEEEauthorblockN{David Gregg}
+\IEEEauthorblockA{School of Computer Science and\\Statistics\\
+Trinity College Dublin\\
+Dublin 2, Ireland\\
+Email: David.Gregg@cs.tcd.ie}}
 
-% Title portion
-\title{Title goes here}
-\author{
-SERVESH MURALIDHARAN
-\affil{Trinity College Dublin}
-DAVID GREGG
-\affil{Trinity College Dublin}}
+% conference papers do not typically use \thanks and this command
+% is locked out in conference mode. If really needed, such as for
+% the acknowledgment of grants, issue a \IEEEoverridecommandlockouts
+% after \documentclass
+
+% for over three affiliations, or if they all won't fit within the width
+% of the page, use this alternative format:
+% 
+%\author{\IEEEauthorblockN{Michael Shell\IEEEauthorrefmark{1},
+%Homer Simpson\IEEEauthorrefmark{2},
+%James Kirk\IEEEauthorrefmark{3}, 
+%Montgomery Scott\IEEEauthorrefmark{3} and
+%Eldon Tyrell\IEEEauthorrefmark{4}}
+%\IEEEauthorblockA{\IEEEauthorrefmark{1}School of Electrical and Computer Engineering\\
+%Georgia Institute of Technology,
+%Atlanta, Georgia 30332--0250\\ Email: see http://www.michaelshell.org/contact.html}
+%\IEEEauthorblockA{\IEEEauthorrefmark{2}Twentieth Century Fox, Springfield, USA\\
+%Email: homer@thesimpsons.com}
+%\IEEEauthorblockA{\IEEEauthorrefmark{3}Starfleet Academy, San Francisco, California 96678-2391\\
+%Telephone: (800) 555--1212, Fax: (888) 555--1212}
+%\IEEEauthorblockA{\IEEEauthorrefmark{4}Tyrell Inc., 123 Replicant Street, Los Angeles, California 90210--4321}}
+
+% use for special paper notices
+%\IEEEspecialpapernotice{(Invited Paper)}
+
+% make the title area
+\maketitle
 
 \begin{abstract}
+%\boldmath
+The abstract goes here.
+\end{abstract}
+% IEEEtran.cls defaults to using nonbold math in the Abstract.
+% This preserves the distinction between vectors and scalars. However,
+% if the conference you are submitting to favors bold math in the abstract,
+% then you can use LaTeX's standard command \boldmath at the very start
+% of the abstract to achieve this. Many IEEE journals/conferences frown on
+% math in the abstract anyway.
 
-Abstract goes here
+% no keywords
 
-\end{abstract}
+% For peer review papers, you can put extra information on the cover
+% page as needed:
+% \ifCLASSOPTIONpeerreview
+% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
+% \fi
+%
+% For peerreview papers, this IEEEtran command inserts a page break and
+% creates the second title. It will be ignored for other modes.
+\IEEEpeerreviewmaketitle
 
-\category{C.2.2}{Computer-Communication Networks}{Network Protocols}
-
-\terms{Design, Algorithms, Performance}
-
-\keywords{}
-
-\acmformat{Muralidharan, S., and Gregg, D.,  2012. Title goes here.}
-
-\begin{bottomstuff}
-This work is funded by the IRCSET Enterprise Partnership Scheme along with IBM Research Ireland.
-
-Author's addresses: S. Muralidharan {and} D. Gregg, Computer Science Department,
-Trinity College Dublin, Ireland.
-\end{bottomstuff}
-
-\maketitle
-
-\section{Introduction}
+\section{Motivation}
 
@@
 In generating the application graph as a sequence of operations it also makes it possible to process different stages of the operation in a pipelined manner.
-This could be combined with the stream optimizations and could be used to determine which steps of the graph is best suited to form the stages of a pipeline.
+This when combined with stream optimizations and could be used to determine which steps of the graph is best suited to form the stages of a pipeline.
 The condition to balance here is the weight associated with computation in that stage with that of the overhead incurred in computing the data in two separate stages.
-It would be necessary to determine whether the additional over head involved in executing the task in parallel does provide any significant improvement in performance.
+It would be necessary to determine whether the additional over head involved in executing the task in parallel provides any significant improvement in performance.
 
@@
 
---Click was designed as a means to easily implement a software router on top of a Linux OS.
-Click achieves this by implementing it in two different ways one is by installing a kernel module and the other is a program that executes in the userlevel.
-\\
+Click was designed as a means to easily implement a software router on top of a Linux OS.
+This is implemented in two different ways one is by installing a kernel module and the other is a program that executes in the user level.
+Both of them accept a configuration file described in click's language specification.
+This is then used to construct the router by linking the virtual functions for the push and pull operations of the different elements in accordance with the specified graph.
+
+\subsection{Modifications to click system}
+
+\subsubsection{Static Generation Of ???}
+The first step in the design was to overcome the limitation of using click's infrastructure to construct the software router which consists of dynamically initialized objects and virtual functions.
+This was necessary due to two important conditions, one was to eliminate the overhead associated with the dynamic construction of the objects and virtual functions and the other was to provide the compiler a chance to optimize the program during the compilation phase.
+This would however result in not being able to construct the software router using click's infrastructure but would require the application specific code to be recompiled.
+This could be a significant hurdle in the case of running the software router in kernel mode where the router configuration is installed by the click kernel module in which case the entire kernel module should be recompiled and reinitialized.
+Whereas in user space this would not be an issue since routers of different configurations are basically treated as separate applications and would just have to go through an additional compilation phase before being used.
+The added advantage of this would be to enable network applications that operate on L1 - L3 layers to use the packet processing ( L4 - L7 ) infrastructure provided by click elements.
+
+\subsubsection{Implementation of data parallelism}
 
@@
 
-% Acknowledgments
-\begin{acks}
-The authors would like to thank ...
-\end{acks}
+% An example of a floating figure using the graphicx package.
+% Note that \label must occur AFTER (or within) \caption.
+% For figures, \caption should occur after the \includegraphics.
+% Note that IEEEtran v1.7 and later has special internal code that
+% is designed to preserve the operation of \label within \caption
+% even when the captionsoff option is in effect. However, because
+% of issues like this, it may be the safest practice to put all your
+% \label just after \caption rather than within \caption{}.
+%
+% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
+% option should be used if it is desired that the figures are to be
+% displayed while in draft mode.
+%
+%\begin{figure}[!t]
+%\centering
+%\includegraphics[width=2.5in]{myfigure}
+% where an .eps filename suffix will be assumed under latex, 
+% and a .pdf suffix will be assumed for pdflatex; or what has been declared
+% via \DeclareGraphicsExtensions.
+%\caption{Simulation Results}
+%\label{fig_sim}
+%\end{figure}
 
-% Bibliography
-\bibliographystyle{acmsmall}
-\bibliography{}
+% Note that IEEE typically puts floats only at the top, even when this
+% results in a large percentage of a column being occupied by floats.
 
-% History dates
-\received{June 2012}
+% An example of a double column floating figure using two subfigures.
+% (The subfig.sty package must be loaded for this to work.)
+% The subfigure \label commands are set within each subfloat command, the
+% \label for the overall figure must come after \caption.
+% \hfil must be used as a separator to get equal spacing.
+% The subfigure.sty package works much the same way, except \subfigure is
+% used instead of \subfloat.
+%
+%\begin{figure*}[!t]
+%\centerline{\subfloat[Case I]\includegraphics[width=2.5in]{subfigcase1}%
+%\label{fig_first_case}}
+%\hfil
+%\subfloat[Case II]{\includegraphics[width=2.5in]{subfigcase2}%
+%\label{fig_second_case}}}
+%\caption{Simulation results}
+%\label{fig_sim}
+%\end{figure*}
+%
+% Note that often IEEE papers with subfigures do not employ subfigure
+% captions (using the optional argument to \subfloat), but instead will
+% reference/describe all of them (a), (b), etc., within the main caption.
 
+% An example of a floating table. Note that, for IEEE style tables, the 
+% \caption command should come BEFORE the table. Table text will default to
+% \footnotesize as IEEE normally uses this smaller font for tables.
+% The \label must come after \caption as always.
+%
+%\begin{table}[!t]
+%% increase table row spacing, adjust to taste
+%\renewcommand{\arraystretch}{1.3}
+% if using array.sty, it might be a good idea to tweak the value of
+% \extrarowheight as needed to properly center the text within the cells
+%\caption{An Example of a Table}
+%\label{table_example}
+%\centering
+%% Some packages, such as MDW tools, offer better commands for making tables
+%% than the plain LaTeX2e tabular which is used here.
+%\begin{tabular}{|c||c|}
+%\hline
+%One & Two\\
+%\hline
+%Three & Four\\
+%\hline
+%\end{tabular}
+%\end{table}
+
+% Note that IEEE does not put floats in the very first column - or typically
+% anywhere on the first page for that matter. Also, in-text middle ("here")
+% positioning is not used. Most IEEE journals/conferences use top floats
+% exclusively. Note that, LaTeX2e, unlike IEEE journals/conferences, places
+% footnotes above bottom floats. This can be corrected via the \fnbelowfloat
+% command of the stfloats package.
+
+% conference papers do not normally have an appendix
+
+% use section* for acknowledgement
+\section*{Acknowledgment}
+This work is funded by the IRCSET Enterprise Partnership Scheme along with IBM Research Ireland.
+
+% trigger a \newpage just before the given reference
+% number - used to balance the columns on the last page
+% adjust value as needed - may need to be readjusted if
+% the document is modified later
+%\IEEEtriggeratref{8}
+% The "triggered" command can be changed if desired:
+%\IEEEtriggercmd{\enlargethispage{-5in}}
+
+% references section
+
+% can use a bibliography generated by BibTeX as a .bbl file
+% BibTeX documentation can be easily obtained at:
+% http://www.ctan.org/tex-archive/biblio/bibtex/contrib/doc/
+% The IEEEtran BibTeX style support page is at:
+% http://www.michaelshell.org/tex/ieeetran/bibtex/
+%\bibliographystyle{IEEEtran}
+% argument is your BibTeX string definitions and bibliography database(s)
+%\bibliography{IEEEabrv,../bib/paper}
+%
+% <OR> manually copy in the resultant .bbl file
+% set second argument of \begin to the number of references
+% (used to reserve space for the reference number labels box)
+\begin{thebibliography}{1}
+
+\bibitem{IEEEhowto:kopka}
+H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
+  0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.
+
+\end{thebibliography}
+
+% that's all folks
 \end{document}

### 2012-06-26 paper.tex
Wrote motivation, commit before david's update

@@
 %\boldmath
-The abstract goes here.
+the abstract goes here.
+
 \end{abstract}
@@
 \section{Motivation}
+In the continued race for improvement of multi core and many core architectures, computer networks are becoming a fundamental area for performance bottleneck especially in systems which are used for distributed processing of parallel applications.
+In order to handle the requirements posed by these applications the interconnects are becoming more and more complex.
+Such systems utilize a wide range of custom interconnects ranging from infiniband, myrinet to the more readily available ones such as 1Gbe and 10Gbe Ethernet standards and sophisticated routers and switches to handle the added challenge in routing data.
+While a small group of the specialized systems employ custom routers and topology configurations the majority of them especially those that use 1Gbe or 10Gbe use the commercial routers.
+In addition the growing complexity of routing protocols has resulted in advanced switches such as OpenVSwitch \cite{openvswitch}, etc., has resulted in these routers being upgraded frequently to keep up with the growing requirements.
+Software routers such as click modular router have been gaining popularity due to their flexible nature and cheaper alternative to the more expensive commercial routers and are widely used to prototype newer standards.
+In using software routers in the above scenario it would be possible to not only benefit from the newer developments in switching and routing protocols but it would require little effort to implement it in existing systems.
+One might argue that it would be a challenge to handle such complex workloads in software routers but with the presence of multicore and many core architectures and stream processors it is feasible to parallelize the packet processing workloads to improve the overall performance of the system.
+Several research projects are working on improving the performance by parallelizing the packet processing tasks on these architectures \cite{routebricks} \cite{packetshader}.
+
+In a computing cluster which is used in the execution of parallel applications it is common to dedicate a core or processor in each node for operating system related tasks in order to minimize its effects on the application.
+Similarly, it should also be possible to utilize an additional core or processor to perform packet processing operations.
+While this is would be a challenging scenario it does provide the added benefits of a cheaper alternative to commercial routers.
+Also in bringing the network system into a software controlled domain it should be possible to perform additional optimizations from the application such as prioritized data flow, intelligent routing, etc., that would otherwise be difficult to implement due to the abstraction of the network system.
+
+The focus of this paper is the development of a parallel processing engine that could be integrated with the application and can perform both the actual computations and the packet processing related tasks in parallel.
+The entire system can be divided into key parts as follows,
+
+\begin{itemize}
+
+\item A suitable methodology by which we can access the data in the userspace without interacting with the operating system.
+To do this (1)We need to read the data from the Network Interface Controller (NIC) before the OS and (2)Using this, establish a efficient manner for sending and receiving the data.(Section \ref{sec:userspace-packet})
+
+\item Given a graph that represents the workload to be done on the packet flow, we divide it into tasks that can be executed in parallel and determine the regions that could be used to extract data and task parallelism.(Section \ref{sec:workload})
+
+\item The parallel processing engine that utilizes the above and does the actual computations on the multi core or many core processor.
+It also handles load balancing and scheduling of the work across the available computing resources.
+(Section \ref{sec:ppe})
+
+\end{itemize}
+
+We concentrate more on the framework that can integrate the packet processing and application tasks into a parallel processing engine and the extraction of parallelism from the packet processing related tasks to improve the overall performance.
+Given an application graph and its partition consisting of parallel tasks we the combine it with the packet processing operations.
+In addition, Section \ref{sec:} we show the challeneges involved in using a system consisting of threads and process that are used to actually perform the work.
+The extent to which we extract parallelism ranges not only to the packet tasks but to the low level access of data I/O to the network interface hardware through certain additions to the netmap APIs.
+Finally in Section \ref{sec: } we present certain pseudo applications that involves operations such as compression and encryption to evaluate our parallel processing engine.
+
+\section{Background}
+
+An important part of our system is the tasks related to packets such as routing, forwarding, etc., Since our design focuses more on the aspects of the parallel framework we utilize the click modular router for the packet processing tasks.
+Click modular router \cite{click} is a commonly used software router that includes a library of elements which is used in building flexible and reconfigurable routers.
+It offers a platform that is flexible for rapid development of newer network protocols and achieves good performance.
+An improved version known as SMP Click \cite{smpclick} also supports thread based parallel processing of network related tasks.
+Until recently click's performance in the user space was not as good as its kernel space equivalent.
+This is mainly due to its extensive use of polling driver in the kernel space.
+Netmap I/O \cite{netmap} provides a set of APIs that could directly access the packets from the NIC and it helped in the recent improvement of click's user space performance.
+
+Click dynamically constructs routers in which the configuration file is passed as a parameter for the kernel module or user space binary.
+As a result most of the actual components are dynamically allocated and connected at run time.
+While this may be an advantage considering the alternate is to recompile the linux kernel, it does act as significant bottleneck in the case of user space.
+Moreover some of the restrictions on how the configuration can be parallelized limits the actual performance that could be achieved in combination with netmap.
+By integrating the ability to access click's elements within our parallel processing engine we were able to scale significantly better.
+This does require a limited amount of modifications to some of click's libraries and elements.
+In generating the actual code statically, compilers such as gnu c, etc., can perform optimizations such as loop fusions, in-lining, dead code optimization etc., which were not possible previously.
+
+Our framework primarily focuses on user space and netmap APIs forms a crucial part in this.
+
+\section{Background}
 
@@
 \section*{Acknowledgment}
-This work is funded by the IRCSET Enterprise Partnership Scheme along with IBM Research Ireland.
+This work is funded by the IRCSET Enterprise Partnership Scheme along with IBM Research, Ireland.
 
@@
 
-\bibitem{IEEEhowto:kopka}
-H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
-  0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.
-
 \end{thebibliography}

### 2012-06-27 paper.tex
Added new background section

@@
 %
-%\usepackage{cite}
+\usepackage{cite}
 % cite.sty was written by Donald Arseneau
@@
 \ifCLASSINFOpdf
-  % \usepackage[pdftex]{graphicx}
+  \usepackage[pdftex]{graphicx}
   % declare the path(s) where your graphic files are
-  % \graphicspath{{../pdf/}{../jpeg/}}
+  \graphicspath{{./eps/}}
   % and their extensions so you won't have to specify these with
   % every instance of \includegraphics
-  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
+  \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
 \else
@@
 % the caption=false option directly:
-%\usepackage[caption=false,font=footnotesize]{subfig}
+\usepackage[caption=false,font=footnotesize]{subfig}
 %
@@
 We concentrate more on the framework that can integrate the packet processing and application tasks into a parallel processing engine and the extraction of parallelism from the packet processing related tasks to improve the overall performance.
-Given an application graph and its partition consisting of parallel tasks we the combine it with the packet processing operations.
-In addition, Section \ref{sec:} we show the challeneges involved in using a system consisting of threads and process that are used to actually perform the work.
+Given an application graph and its partition consisting of parallel tasks we then combine it with the packet processing operations.
+In addition, Section \ref{sec:} we show the challenges involved in using a system consisting of threads and process that are used to actually perform the work.
 The extent to which we extract parallelism ranges not only to the packet tasks but to the low level access of data I/O to the network interface hardware through certain additions to the netmap APIs.
@@
 \section{Background}
+
+Let us consider a simple file storage system that is based in a data center.
+The application handles sensitive information by performing encryption on the data that matches predefined rules like file names, access level of user, etc., otherwise the default operation is to perform compression and store the information in the location specified.
+For the purpose of this example let us assume that the filesystem is located on a NFS server and it is responsible for concurrent operations on the data.
+A group of servers act as the interface to the storage system and the requests are evenly distributed over these nodes.
+The important operations that these nodes perform on the data are represented in the stream graph shown in figure \ref{fig1}.
+In the data retrieval part the information fetched from the filesystem is checked to see if it matches a set of predefined rules to check if it has to be encrypted otherwise passes it to the compression operation.
+During data storage the input data is checked whether it has to be decrypted or decompressed and then stored on to the filesystem.
+
+\begin{figure}[ht]
+\centering
+\subfigure
+{
+	\includegraphics[width=3.5in]{ip-data-proc}
+	\label{fig1.1}
+}
+
+\subfigure
+{
+	\includegraphics[width=3in]{op-data-proc}
+	\label{fig1.2}
+}
+\caption{Operations performed on the data stored and retrieved}
+\label{fig1}
+\end{figure}
+
+The nodes that are connected in the cluster are linked through a switch or router.
+In such a scenario the number of routers increase with the size of the cluster and the bandwidth required by the application.
+Due to this the performance of the application is directly dependent on the network especially the functionality of the routers.
+Moreover with a fixed infrastructure such as this it is not only difficult to upgrade the routers but it would also be a expensive proposition.
+In such a situation it makes more sense to use software routers which are much more flexible and cheaper and could prove to be a more effective than or in addition to commercial switches or routers.
+
+In implementing the software routers if one were to dedicate several nodes for this purpose it might prove to be inefficient use of resources since the network traffic generated is highly dependent on the applications being executed in such a closed environment.
+Moreover such systems might act as a communication bottleneck and the routing operations are focused on these nodes.
+A more effective approach would be to bring the software routers into the nodes that are used for executing the applications itself.
+In doing this we dedicate a specific set of resources of each node for the packet processing operations and in essence make each node act as a router.
+All the nodes in such a cluster are connected with each other based on network topology suitable for its requirements.
+This would make it possible to divide the workload generated due to the network operations across all the nodes in the cluster.
+In this paper we propose such a system that not only brings the software routers as separate entity into the actual application nodes but takes it a step further and integrates the network operations with the application itself.
+
+There are several challenges arises due to this mainly how do we tackle such large number of computations both from the application and the network effectively.
+Two important constraints would enable us to satisfy the requirements posed by such a system.
+First, the operations of both the application and the network is parallel in nature and second, the presence of co-processors[], network processors[], Graphics Processing Units ( GPUs ) \& General Purpose GPUs ( GPGPUs ) has dramatically improved the capability of commodity hardware.
+With this theoretically, we should be able to meet the computational demand but if such a system were to succeed an efficient parallel processing engine is crucial.\\
+
+\textit{write a para forward referencing other sections}\\
 
@@
 Netmap I/O \cite{netmap} provides a set of APIs that could directly access the packets from the NIC and it helped in the recent improvement of click's user space performance.
+We found the simpler construction of netmap easier to adapt to our framework  compared to other user space packet access APIs.
 
@@
 
-Our framework primarily focuses on user space and netmap APIs forms a crucial part in this.
-
-\section{Background}
+\section{Old Background}
 
@@
 Discuss about what each sections contains here.....
-
-\section{Background}
-
-Talk about click...\\
-
-Describe about what would be the best way to parallelize using threads and MPI.. provide reference.. \\
 
@@
 % (used to reserve space for the reference number labels box)
-\begin{thebibliography}{1}
+%\begin{thebibliography}{1}
 
-\end{thebibliography}
+%\end{thebibliography}
 

### 2012-06-27 script.sh

@@ -1,4 +1,5 @@
 #!/bin/sh
 while inotifywait paper.tex; do
-    pdflatex -interaction=nonstopmode paper.tex
+#    pdflatex -interaction=nonstopmode paper.tex
+     make
 done

### 2013-03-18 paper.tex
Re commit

--- 2012-05-21 paper.tex
@@
 
+\usepackage{listings}
+\usepackage{color}
+
+\definecolor{dkgreen}{rgb}{0,0.6,0}
+\definecolor{gray}{rgb}{0.5,0.5,0.5}
+\definecolor{mauve}{rgb}{0.58,0,0.82}
+
+\lstset{ %
+  language=C,                % the language of the code
+  basicstyle=\footnotesize\ttfamily,           % the size of the fonts that are used for the code
+  numbers=right,                   % where to put the line-numbers
+  numberstyle=\tiny\color{gray},  % the style that is used for the line-numbers
+  stepnumber=0,                   % the step between two line-numbers. If it's 1, each line 
+                                  % will be numbered
+  numbersep=0pt,                  % how far the line-numbers are from the code
+  %backgroundcolor=\color{white},      % choose the background color. You must add \usepackage{color}
+  showspaces=false,               % show spaces adding particular underscores
+  showstringspaces=false,         % underline spaces within strings
+  showtabs=false,                 % show tabs within strings adding particular underscores
+  %frame=single,                   % adds a frame around the code
+  rulecolor=\color{black},        % if not set, the frame-color may be changed on line-breaks within not-black text (e.g. commens (green here))
+  tabsize=2,                      % sets default tabsize to 2 spaces
+  captionpos=b,                   % sets the caption-position to bottom
+  breaklines=true,                % sets automatic line breaking
+  breakatwhitespace=false,        % sets if automatic breaks should only happen at whitespace
+  title=\lstname,                   % show the filename of files included with \lstinputlisting;
+                                  % also try caption instead of title
+  keywordstyle=\bfseries,          % keyword style
+  commentstyle=\color{dkgreen},       % comment style
+  stringstyle=\color{mauve},         % string literal style
+  escapeinside={\%*}{*)},            % if you want to add LaTeX within your code
+  morekeywords={*,...}               % if you want to add more keywords to the set
+}
+
 % *** SPECIALIZED LIST PACKAGES ***
 %
+\usepackage{algpseudocode}
+\usepackage{algorithm}
 %\usepackage{algorithmic}
+%\usepackage{algorithmicx}
 % algorithmic.sty was written by Peter Williams and Rogerio Brito.
@@
 
-%\usepackage{mdwmath}
-%\usepackage{mdwtab}
+\usepackage{mdwmath}
+\usepackage{mdwtab}
 % Also highly recommended is Mark Wooding's extremely powerful MDW tools,
@@
 % *** SUBFIGURE PACKAGES ***
-%\usepackage[tight,footnotesize]{subfigure}
+\usepackage[tight,footnotesize]{subfigure}
 % subfigure.sty was written by Steven Douglas Cochran. This package makes it
@@
 
-%\usepackage[caption=false]{caption}
+\usepackage[font=small,center]{caption}
 %\usepackage[font=footnotesize]{subfig}
@@
 %\usepackage[caption=false,font=footnotesize]{subfig}
+
 %
@@
 
+\newcommand{\INDSTATE}[1][1]{\STATE\hspace{#1\algorithmicindent}}
+\newcommand{\comment}[1]{}
+
 % correct bad hyphenation here
@@
 % can use linebreaks \\ within to get better formatting as desired
-\title{Title goes here}
+\title{A high-level programming framework to support parallel processing of communication intensive \\ stream applications}
 
@@
 %\boldmath
-The abstract goes here.
+Stream applications are often limited in their performance by the communication system employed.
+In systems that use conventional 1GbE or 10GbE standard, the Operating System (OS) handles most if not all of the network operations.
+In such cases, the communication stack, which was not designed to handle such tremendous amounts of data, acts as a bottleneck and restricts the performance of the application.
+An optimum framework for computing and communicating in parallel is essential for the continued improvement in performance.
+
+We propose a high-level programming framework that integrates the packet processing operations with the stream applications and provides a common interface to execute both the application and packet processing operations in parallel on the multicore processors.
+
+We utilize the ability of network interface controller (NIC) to classify packets on to multiple hardware queues.
+We combine this with system software that allows direct access to the hardware queues.
+We develop an interface in our framework for application processes to read or write to in parallel.
+Our system integrates this with the application by a series of packet processing operations that transform the data from the application to packets suitable for transmission.
+Representing these tasks as stages in the application itself, we are able to parallelize them using techniques similar to those used for parallelizing regular computations.
+
+All the operations required for the communication are performed in the application, bypassing the OS to send and receive data directly over the network interface.
+By overcoming the bottleneck induced by performing network operations in the OS and by parallelizing both the computations and packet operations together, we are able to boost the performance of the application.
+
 \end{abstract}
@@
 
-The constant advancement in server hardware has made systems with large number of hardware threads[] readily available.
-In addition to this the presence of co-processors[], network processors[], Graphics Processing Units ( GPUs ) \& General Purpose GPUs ( GPGPUs ) has dramatically improved the capability of commodity hardware.
-These developments has paved way to a platform with immense computing potential which existing applications already take advantage of.
-Along with the improvements in the Multi-Core \& Many-Core processor architecture the memory and I/O speeds have also seen significant improvement.
-In particular the network subsystem has seen significant growth, in combination with the improvements to optical communication it has lead to 10GbE interfaces being quite common in enterprise servers.
-While 40GbE and 100GbE are available for backbone infrastructure but would eventually follow into commodity hardware.
+Stream applications are designed in a manner in which they operate on data that flows as a \textit{stream}.
+This data stream passes through different stages of computations, to be transformed from query to result \cite{Thies:2002:SLS:647478.727935} \cite{Halbwachs91thesynchronous} \cite{Stephens95asurvey}.
+This is different from conventional computation model where data is stored and then processed.
+By eliminating the need to store data and processing it on the go, it makes it possible for stream applications to compute on large amounts of data.
 
-Network applications represent those applications that operate on any of the seven network layers described in the OSI Model. \{describe more here\}
+Stream applications such as file storage systems, database query management, data mining, etc., that operate on vast amount of data are often distributed over multiple systems.
+Traditional stream applications do target multicore processors for performing their computations in parallel.
+Due to this and the exponential rate at which multicore processors are growing the amount of computations performed by these stream applications would further increase and eventually lead to processing even larger amounts of data.
+With the growth in the complexity of these applications it is necessary to leverage on a multicore environment to improve their communication performance as well.
+Such stream applications that operate on large amounts of data at very high bandwidth are often limited in their performance by the communication stack of the OS kernel\cite{Wu:2007:PPB:1323954.1323957}\cite{Wu:2007:PAL:1227865.1228021}\cite{Dobrescu09routebricks:exploiting}\cite{Han:2010:PGS:1851275.1851207}\cite{Kohler2000}.
+In the continued race for improvement of multicore and many core architectures, it is becoming essential to improve the processing speed of packet operations along with those of the application in order to scale the performance of the application.
 
-Several previous work have discussed the process of parallelizing network applications[][][] to enable it to run faster in such multi threaded systems.
-The development of Software routers[][][] is encouraged further because it costs only a fraction of the commodity hardware routers.
-Moreover the advancements in the network interfaces has made faster interfaces readily in combination with the multi core servers the performance  It has been shown that such routers can be used to process packets at line rate [][][] which has been a challenge on its own.
-Network applications including those that operate on few or all of the network layers have been designed with sequential processing as the target environment.
-The shift to multi core architectures in recent times has induced such applications to utilize parallel processing as a alternate means of computing.
-While this could lead to significant improvement in performance it also suffers from the same problems as any parallel application.
-These include scalability, load-balancing, data dependency and the presence of non parallel regions in the program apart from those that might arise due to that specific application.
-The scaling of software routers has been discussed extensively in RouteBricks[] which shows the problems suffered by software routers when they are scaled beyond a single system.
-One could argue that complexity of network applications such as Deep Packet Inspection ( DPI ), Packet Forwarding, etc., could be handled by a single many socket Multi-Core SMP system.
-However, as the complexity of network applications develops it would be inevitable but to move to a multi node environment in order to perform the computations and to handle more network bandwidth.
+When performing packet processing operations for communication intensive stream applications, a major challenge is the very high levels of performance required.
+These applications process large numbers of data streams in real time.
+If the communication system cannot keep up with the rate at which packets arrive due to these data streams, then they must be dropped or remain unprocessed.
 
-Large scale computing clusters has been used in high-performance scientific application to solve problems and applications of very large size.
-It uses a cluster of compute nodes and distributes the workload across them in order to solve it faster.
-These systems utilize parallel programs consisting of both threads and MPI to break massive problems into smaller workload that can be executed in parallel.
-In addition, the presence of GPUs and GPGPUs requires a mixture of parallelization constructs such as OpenCL, CUDA, etc., in order to achieve the best performance in the system.
-In this paper we discuss how some of the techniques used by these applications could be used to parallelize network applications.
+In building a framework for such applications on general purpose computers, it may be possible to exploit two hardware features which are multicore processors and multi queue network controller that could allow dramatically improved performance.
+Using this it makes it possible to do much of the packet processing in parallel using the multiple cores that are available on modern processors.
 
-Click Modular Software Router[] consists of extensive elements which are used to construct software routers.
-Click primarily targets packet processing applications especially those that operate on L4 - L7.
-\{elaborate more here\}
+A major bottleneck for packet processing on general purpose computers is the overhead of the operating system's network stack.
+In particular, for applications that run as user processes, data must be copied between OS kernel space and user space.
+Previous research use alternate mechanism to access packets from the network hardware directly due to this\cite{Dobrescu09routebricks:exploiting}\cite{Han:2010:PGS:1851275.1851207}\cite{Kohler2000}.
+Most of these use modern network hardware in combination with a particular systems software that allows direct access to hardware-controlled packet queues to dramatically reduce overheads and improve performance.
+However, constructing stream applications that use these features is difficult.
+Furthermore, many stream applications have a very specific communication pattern, typically operating on data between a source and sink in the form a \textit{data stream}.
+Data in a stream must be processed sequentially, whereas those from separate streams can often be processed in parallel.
 
-Discuss about what each sections contains here.....
+Modern network hardware typically has a mechanism for separating arriving packets into multiple queues, where packets that belong to the same flow are guaranteed to be placed in the same queue\cite{micro2008} \cite{intel2010}.
+In many stream applications we can utilize this to operate on multiple streams of data by processing each of the queues in parallel.
+However, building such applications is complex because it requires explicit parallel programming, and it requires explicit interaction with the low-level network interfaces.
+Research focuses on improving the performance of network applications by parallelizing the packet processing tasks on multicore architectures \cite{Dobrescu09routebricks:exploiting} \cite{Han:2010:PGS:1851275.1851207}.
+We parallelize both the application and packet operations to gain significant improvement in performance.
+
+\section{Contribution}
+
+In this paper we propose a high-level programming framework for building communication intensive stream applications.
+We focus on integrating the packet processing and application tasks into a single graph and how we use this to extract data parallelism.
+The developer describes their program as a graph of \textit{actors}.
+Each actor consumes one or more \textit{tokens} of data, which may be simple data such as integers, or more complex data such as packets, arrays or objects and produces one or more tokens.
+The framework can be used to automatically map the application over multiple parallel cores, and interface with the \texttt{netmap} APIs \cite{Rizzo:2012:RNI:2090147.2103536} for accessing data through the network buffers.
+
+The system can be divided into three main parts:
+
+\begin{itemize} 
+\item A subsystem for interfacing with the network hardware through the \texttt{netmap} APIs to send and receive data from the application in user-space without going through the operating system.
+(Section \ref{ppio})
+\item Given a graph that represents the workload to be performed by the application, we construct and integrate the packet processing operations represented as a directed acyclic graph (DAG) to it.
+(Section \ref{ppe})
+\item A run-time system that executes these tasks in parallel and explicitly schedules them on the multicore processor. (Section \ref{ppt})
+\end{itemize}
+
+Even though we integrate the system with the application the techniques for extracting parallelism from the application is not in scope of this paper.
+The application is considered to be in a streaming model and the tasks corresponding to the various functions are considered to be parallel in nature.
 
@@
 
-Talk about click...\\
+Let us consider a sample application that operates on data streams.
+The important operations of the application are represented in the stream graph shown in figure \ref{fig1}.
+The application operates by performing encryption on the data that matches predefined rules like file names, access level of user, etc., otherwise it performs compression and store the information in the location specified.
+Considering a scenario in which, a group of servers are used to perform these operations by evenly distributing them over the different nodes.
 
-Describe about what would be the best way to parallelize using threads and MPI.. provide reference.. \\
+\begin{figure}[ht]
+\centering
+\subfigure
+{
+	\includegraphics[width=3.5in]{ip-data-proc}
+}
 
-\section{Design}
+\subfigure
+{
+	\includegraphics[width=3in]{op-data-proc}
+}
+\caption{Sample stream application}
+\label{fig1}
+\end{figure}
 
-This section describes the design principles related to our methodology.
-The objective is to determine the requirements necessary to parallelize network applications.
-Our methods focus on ways to improve the click system in some cases even extend its functionality.
+The main challenge that arises in such a situation is how to tackle such large number of computations from the application and the network operations due to the intensive communication.
+Two important features enable us to tackle this challenge posed by such a system.
+Firstly, the operations of both the application and the network is parallel in nature and secondly, the presence of multicore processors to execute the application.
 
-\subsection{Emphasize on user level execution environment}
+In our system we use an efficient parallel processing framework that could leverage on multicore processors and multi queue network controllers to achieve this.
+The performance of multicore processors has dramatically improved the capability of commodity hardware.
+With this we would be able to meet the computational demand of the application.
+In addition, the presence of multi queue network interface controller which was designed to handle multiple streams of data could help us achieve the necessary communication performance.
 
-The existing software routers are predominantly implemented in the kernel space for the ease of access to resources and can bypass the OS overhead incurred when data is accessed through the network subsystem.
-While this alternative provides better performance it also leads to concerns regarding stability and security of the system.
-Newer techniques such as Netmap[] allows direct access to network buffers which makes it possible to write high performance packet processing applications in userspace itself.
-Typically network applications which operate on L4 - L7 are developed in the kernel space in order to intercept the network packets before it flows through the OS.
-Whereas those applications that utilize L1 - L3 are present in the user space.
-The objective here is to completely switch to user space to solve any of the stability or security concerns.
-In addition, scenarios where the applications that operate across most or all the layers could be improved by combining the operations across the different layers in user space itself.
+An important part of our system is the tasks related to packets such as Userdatagram Protocol (UDP), Internet Protocol (IP), etc., We utilize the click modular router for these packet processing tasks.
+Click modular router is a tool that includes a library of elements which is used in building flexible, re-configurable and modular software routers \cite{Kohler2000}.
+It offers a platform that is flexible for rapid development of newer network protocols and achieves good performance.
+An improved version known as SMP Click supports thread based parallel processing of network related tasks\cite{Chen:2001:FCP:647055.759948}.
+Until recently click's performance in the user space was not as good as its kernel space equivalent.
+This is mainly due to its extensive use of polling driver in the kernel space that was not possible in user space.
+Netmap I/O\cite{Rizzo:2012:RNI:2090147.2103536} provides a set of APIs that could directly access the packets from the NIC and it helped in the recent improvement of click's user space performance.
+We found the simpler construction of \texttt{netmap} easier to adapt to our framework compared to other user space packet access APIs \cite{Rizzo:2012:RNI:2090147.2103536} \cite{1564468} \cite{Krasnyansky}.
 
-\subsection{Eliminate the overhead associated with creation of dynamic objects}
+Section \ref{dppl} represents the design principles that we follow to conceptualize our framework.
+Section \ref{imple} describes the implementation details used in the construction of the system.
+In section \ref{ppe}, we show the challenges involved in using a system consisting of processes that are used to actually perform the work.
+The extent to which we extract parallelism ranges not only to the packet tasks but to the low level access of data I/O to the network interface hardware through certain additions to the \texttt{netmap} APIs.
+Finally in Section \ref{app-const} we present a simple application constructed from operations such as compression and encryption in order to evaluate our high-level framework.
 
-Click constructs application graphs based on dynamic creation of objects and passing the data packets across the virtual functions.
-Though this is a very elegant approach in creating the software router it does incur the penalty of the virtual function tables.
-The click-tool[] addresses this issue but it does not solve the problem entirely.
-The construction of the software router at run time is the primary reason behind it.
-This mainly arises due to the presence of kernel module which installs the router as a kernel module based on the input configuration.
-However in the user level it would not be necessary for a separate process to initialize the router instead we could generate the entire program statically including the necessary flow associated with the specific application.
-This not only eliminates the overhead associated but also opens up means to optimize the graph further.
+\section{Design Principles}
+\label{dppl}
 
-\subsection{Optimize based on application graph}
+The design of a high-level programming framework that integrates packet processing tasks into the application requires overcoming several challenges.
+\comment{The entire system is driven by the extent to which parallelism can be extracted.
+Some of the techniques that can be applied to the packet processing tasks can be used for the tasks related to the application.} In this section, we describe these challenges and the principles that are used to tackle them.
+\comment{We focus on the methodologies related to the packet processing tasks and the parallel processing engine itself.}
 
-Network applications especially that operate on the lower layers are very sensitive to the overhead associated with operating on the data separately on each layer.
-Instead given an application graph if we were to provide a sequential flow of the different functions at the various layers then it would be possible for the compiler to optimize the regions of program into a single and optimized block of code.
-This would also improve the cache locality of the data accessed by consecutive functions.
-This also presents us with the ability to perform other types of stream graph optimizations[] which would not be possible if the specific graph was constructed dynamically at run time.
+\subsection{Userspace Packet I/O}
 
-\subsection{Automated data parallelism}
+For a packet processing system to be integrated with an application it is necessary to have an efficient mechanism for accessing the data from the NIC without interacting with the operating system.
+Rapid data access here is crucial for the extraction of data parallelism at later stages.
+\comment{In using the techniques described below we exploit the modern NIC architecture to its full potential.}
 
-Network traffic is a classic example by which data parallelism can be utilized to achieve scalable parallelism.
-Click's support for parallelism in user level is described as experimental, it is implemented by using a task graph consisting of each work associated with each element which is executed in the threads.
-The support for parallelism in the kernel mode is better in comparison to that of user mode.
-Since our target environment is in the user space the proposed design involves a different approach to which we can extract data parallelism.
-We first split the data across multiple queues and then replicate the application graph across to process the data the queues in parallel.
-An important condition here is to make sure packets of the same flow should pass through the same queue.
-Since the application graph can be replicated it also makes it possible to dynamically load balance several applications based on the workload.
+\subsubsection{Handling Multiple Hardware Queues}
 
-\subsection{Efficient implementation of task parallelism}
+The development of virtualization and improved network flow handling has lead to hardware level multi queue support in the NICs.
+During transmission concurrent writes to different queues are possible, enabling multiple processes to send data simultaneously.
+During reception the NIC classifies the packet on to one of the receive queues using a technique known as Receive Side Scaling (RSS)\cite{micro2008} \cite{intel2010}.
+It uses the header information or tuples to classify packets on to different hardware queues.
+Multi queue support can handle several streams of data in parallel and it is exploited extensively in our framework to process data concurrently.
+By offloading the classification of packets to the NIC, it removes any overhead associated with packet classification in the application.
+We use these hardware queues as a means of not only accessing the data in parallel but balancing the workload across different threads.
+This is done by controlling the number of queues that each process handles based on the rate at which packets are received in each of the queues.
+It is necessary to have different processes accessing the transmit and receive queues to maximize concurrent access to data.
 
-In generating the application graph as a sequence of operations it also makes it possible to process different stages of the operation in a pipelined manner.
-This when combined with stream optimizations and could be used to determine which steps of the graph is best suited to form the stages of a pipeline.
-The condition to balance here is the weight associated with computation in that stage with that of the overhead incurred in computing the data in two separate stages.
-It would be necessary to determine whether the additional over head involved in executing the task in parallel provides any significant improvement in performance.
+\subsubsection{Flow based Packet Batching}
+
+Packet batching is a common method used to boost the performance of network applications.
+Packet batching combines several packets and performing operations on them together, thus reducing the overhead involved in the I/O operations.
+
+In our framework we utilize batching but the number of packets that forms a batch is not fixed but varies based on the available space in the hardware queue.
+This number reflects the amount of data we are able to process before issuing hardware synchronization calls and forms a important means by which we reduce the usage of system calls.
+
+\subsubsection{Reduction of Per Packet System Calls}
+
+System calls are required in order to synchronize data between the NIC hardware and the buffer memory used for intermediate storage of packets.
+Previous work describe the overhead associated with per packet system calls and provide means of optimizing this\cite{1564468}\cite{Han:2010:PGS:1851275.1851207}\cite{Rizzo:2012:RNI:2090147.2103536}.
+
+We employ an \textit{adaptive scheme} described in section \ref{ppio} that uses the above principle.
+Our framework balances the system calls based on the communication rate and application processing speed.
+The transmission and reception are handled differently to optimize the usage of system calls in each of the case independently.
+During transmission we issue system calls to synchronize the available buffer space based on the rate at which we transmit data.
+This prevents degradation in latency for larger packets and improves performance for smaller packets given the same bandwidth at which we transmit in both the cases.
+During reception of packets the system reads packets at the rate at which it is able to process them.
+
+\subsection{Packet Processing Graph}
+
+\comment{Understanding the set of packet processing tasks associated with the application is essential to performing them in parallel.} We use a graph based representation of packet processing tasks to determine the dependency involved for the different stages and determine the relation with that of the different tasks of the application.
+The graph contains the stages through which the data, associated with the application, is transformed into network packets and vice versa.
+It shows which of the tasks can be performed in parallel with those of the application and highlights those that have data dependency on a specific task in the application.
+This graph also enables optimizations based on the characteristics of the application or the architecture used for execution.
+Figure \ref{fig2} shows a User Datagram protocol (UDP) packet generation represented in this graph.\\
+
+\begin{figure}[ht]
+\centering
+\includegraphics[width=3.5in]{pack-op}
+\caption{Simplified udp packet construction}
+\label{fig2}
+\end{figure}
+
+\subsection{Application Task Graph}
+
+\comment{Application task graph represents the operations associated in performing the computations.} The subset of the application that we are interested in parallelizing is the stream processing component.
+In such components the communication between tasks is predictable and there is correlation between the stages of the application.
+Both the packet processing and application tasks can be represent by a directed acyclic graph (DAG) similar to the packet processing graph.
+The packet processing tasks operate on packets and the application tasks operate on the data.
+Since we focus more on parallel packet processing, the optimization in relation to the application graph are not discussed.
+It is assumed that a static version of the application graph is available and we focus more on executing these tasks in parallel with that of the packet processing tasks.
+
+\subsection{Integration}
+
+The sequence of tasks that can be executed in parallel is determined at this stage.
+The integration stage is used to combine the application graph and the packet processing graph into a larger graph that unifies and represents the overall operations.
+The dependency relations determined on the previous cases are used in establishing links between the graphs.
+We determine the sequence of tasks that can be executed in parallel from both the application task graph and packet processing graph.
+The combined set of tasks is then used by the parallel processing engine.
+
+\subsection{Parallel Processing Engine}
+
+Two important aspects that the parallel processing engine must tackle is the scheduling of tasks and balancing the load across the available resources.
+The parallel processing engine consists of processes which are bound to specific core and act as engines to execute the workload.
+It is constructed in a manner in which its flexible and can adapt to utilize the resources efficiently.
+The tasks are scheduled across these execution engines based on the available resources and its requirements.
+In this manner the workload is distributed uniformly while retaining control over the different execution elements.
 
 \section{Implementation}
+\label{imple}
 
-Click was designed as a means to easily implement a software router on top of a Linux OS.
-This is implemented in two different ways one is by installing a kernel module and the other is a program that executes in the user level.
-Both of them accept a configuration file described in click's language specification.
-This is then used to construct the router by linking the virtual functions for the push and pull operations of the different elements in accordance with the specified graph.
+Most of the existing stream applications depend on the operating system for most of their communication tasks.
+In commodity systems, the interaction with the operating system for network operations is a necessary abstraction to handle multiple applications.
+However, in the case of specialized servers which cater to specific applications it would be beneficial to interact directly with the network device.
+Moreover, in the case of intensive network traffic the operating system could act as a bottleneck due to the overhead associated with data flowing through the kernel before being transmitted through the network device\cite{Wu:2007:PPB:1323954.1323957}\cite{Wu:2007:PAL:1227865.1228021}.
+Eliminating this would be possible by either executing the application in the kernel space which could compromise stability of the system or by using an interface for accessing the NIC buffers from the user space itself.
+The later case of accessing the buffers of network hardware in user space is supported by several interfaces such as \texttt{netmap} \cite{Rizzo:2012:RNI:2090147.2103536}, PF\_RING DNA \cite{1564468}, UIO-IXGBE \cite{Krasnyansky}, etc., Even though the exported APIs support user space packet access they lack the ability to be invoked in a efficient manner by the application.
+The lack of support for interacting with these APIs in order to read and write data in parallel from the different process or threads adds to this problem.
+Moreover since the application is responsible for handling network packets it has to perform the additional operations that would otherwise be carried out by the operating system.
 
-\subsection{Modifications to click system}
+The framework provides support for sending and receiving data in parallel over the network interface from user space and a common parallel execution engine that supports both the application and packet processing tasks.
 
-\subsubsection{Static Generation Of ???}
-The first step in the design was to overcome the limitation of using click's infrastructure to construct the software router which consists of dynamically initialized objects and virtual functions.
-This was necessary due to two important conditions, one was to eliminate the overhead associated with the dynamic construction of the objects and virtual functions and the other was to provide the compiler a chance to optimize the program during the compilation phase.
-This would however result in not being able to construct the software router using click's infrastructure but would require the application specific code to be recompiled.
-This could be a significant hurdle in the case of running the software router in kernel mode where the router configuration is installed by the click kernel module in which case the entire kernel module should be recompiled and reinitialized.
-Whereas in user space this would not be an issue since routers of different configurations are basically treated as separate applications and would just have to go through an additional compilation phase before being used.
-The added advantage of this would be to enable network applications that operate on L1 - L3 layers to use the packet processing ( L4 - L7 ) infrastructure provided by click elements.
+\subsection{Parallel Packet I/O}
+\label{ppio}
+The parallel packet I/O interface when combined with the application tasks provides concurrent communication by the application threads or processes over the network.
+To do this two important requirements are to be met.
+Firstly, it should provice access to network data from user space which we can write to in parallel and secondly, to integrate it with the application by providing necessary interfaces for performing the packet operations.
 
-\subsubsection{Implementation of data parallelism}
+\subsubsection{Userspace Packet Access}
+Applications executed in the kernel space for accessing hardware devices directly would lead to instability of the entire system.
+Instead, several previous work \cite{Rizzo:2012:RNI:2090147.2103536}, \cite{1564468}, \cite{Krasnyansky}, etc., propose access of network buffers from user space.
+We use \texttt{netmap} APIs \cite{Rizzo:2012:RNI:2090147.2103536} to access data from the network buffers.
+\texttt{Netmap} uses a set of ring buffers in the kernel space and user space to export the network hardware buffer's memory region.
+This enables the applications to execute in the user space and directly access the network hardware.
+It prevents the application from accessing memory outside the valid region by checking limits of the ring buffer.
+Packets can be written to the buffers when the space is available and an \textit{ioctl} call to the driver would enable the network controller to send the data.
+A \textit{poll} interface is implemented on the memory region mapped in the user space to check when new packets are available.
+On reading or writing the packets the available space is updated.
 
---Talk about MPI \& threads\\
+We leverage on the availability of multiple hardware queues present in modern NIC to read and write data concurrently to the hardware.
+\texttt{Netmap} by default supports assigning each hardware queue to an individual application process or thread, but we noticed that in order to be efficient in handling the interface we had to assign more than one queue to a particular thread or process.
+We modified \texttt{netmap}'s hardware queue assignment API to support this.
+Later, it also allowed the framework to balance the amount of packets handled by each process by varying the number of queues assigned to each of it in order to balance the workload.
 
---Our goal here is to focus on the network application as a workload and characterize it as such without involving any kind of overhead that might occur due to the network.\\
+We also found that it was essential to have a balanced control over issuing the system calls with respect to the amount of data being transferred.
+Frequent system calls seem to hamper the bandwidth and latency and in essence affect the peak performance of the application.
+To tackle this we propose the following algorithms for reading and writing data from the hardware queues using the \texttt{Netmap} APIs.
+The algorithms are designed to provide these features,
+
+\begin{itemize}
+
+\item Provide batching of packets based on available space on hardware rings
+\item Reduce the usage of system calls such as \textit{ioctl \& poll}
+\item Mechanisms for handling multiple queues through the application
+
+\end{itemize}
+
+By minimizing the synchronizations with the hardware rings we essentially limit the system calls issued and improve the overall system performance.
+
+\begin{algorithm}
+\caption{To send packets}
+\label{s_pack}
+\begin{algorithmic}[1]
+
+\State \textbf{struct} $ring[ n ]$\textbf{\{}
+\State \Comment{Stores information about hardware rings}
+\State \Comment{ $ n \gets \ $tx\ rings\ assigned\ to\ thread}
+\State \  \  $avb$, available space
+\State \  \  $curr$, current position
+\State \  \  $limit$, total space \ \textbf{\}}
+
+\Function{send\_data\_mqueue}{Ring $f(n)$, Data $packet$}
+\While{$!success$}
+	\State $i \gets f( n ) $ \label{s_pack_param}
+	\State $success = $ \texttt{SEND\_DATA}($ i, packet $)
+\EndWhile
+\EndFunction
+\Function{send\_data}{Ring $i$, Data $packet$}
+\If{ $ring[ i ].avb > 0$ }
+	\State $hw\_ring[ i ].slot( ring[ i ].curr ) \gets packet$
+	\State $ring[ n ].avb \gets ring[ i ].avb - 1$
+	\State $ring[ i ].curr++$
+	\If{ $ring[ i ].avb == 0$ }
+		\State $hw\_ring[ i ].avb \ -=\ ring[ i ].limit$
+		\State $hw\_ring[ i ].curr \gets ring[ i ].curr$
+	\EndIf
+	\State \textbf{return} $success$
+\EndIf
+
+\State \textbf{\textit{issue ioctl and poll}}
+\State $ring[ i ].avb \gets hw\_ring[ i ].avb$
+\State $ring[ i ].limit \gets hw\_ring[ i ].avb$
+\EndFunction
+
+\end{algorithmic}
+\end{algorithm}
+
+Algorithm \ref{s_pack} \& \ref{r_pack} uses a batch update for sending and receiving packets based on the rate at which data flows through the interface.
+This is done by maintaining a data structure that has the information about the hardware rings and updating it on a periodic basis.
+Once a fixed amount of space is determined we update this on our local structure, we then continue to read or write data from the hardware ring but update the actual limits and position on our local structure.
+When we have used up the space that was allocated we then update the hardware ring with the changes done.
+In cases where the application is generating sparse traffic it would be possible to frequently update the ring in order to maintain a consistent latency in transmission.
+Multi queue support is provided by managing the process of selecting the ring to send or receive packets based on a application determined function.
+The $f(n)$ present in Algorithm \ref{s_pack} line \ref{s_pack_param} and Algorithm \ref{r_pack} line \ref{r_pack_param} represents this.
+Choosing a random queue usually proves to be effective in our experiments but if the application requires a specific flow to be maintained a different function here could provide that support.
+Cases where this would be helpful is if the data operations are more predictable or if the transactions seem to follow a specific pattern.
+The $threshold$ specified in Algorithm \ref{r_pack} line \ref{r_pack_thres} is determined for each application based on the rate at which it is able to process data.
+By issuing the system calls such as \textit{ioctl} and \textit{poll} only when required and effectively handling the multiple queues we are able to achieve improved performance.
+
+\begin{algorithm}[h]
+\caption{To receive packets}
+\label{r_pack}
+\begin{algorithmic}[1]
+
+\State \textbf{struct} $ring[ n ]$\textbf{\{}
+\State \Comment{Stores information about hardware rings}
+\State \Comment{ $ n \gets \ $rx\ rings\ assigned\ to\ thread}
+\State \  \  $avb$, available packets
+\State \  \  $curr$, current position
+\State \  \  $limit$, total packets \ \textbf{\}}
+
+\Function{receive\_data\_mqueue}{Ring $f(n)$} 
+\While{$!packet$}
+	\State $i \gets f( n ) $ \label{r_pack_param}
+	\State $packet =$ \texttt{RECEIVE\_DATA}(Ring $i$)
+\EndWhile
+\EndFunction
+
+\Function{receive\_data}{Ring $i$}
+\If{ $ring[ i ].avb > 0$ }
+	\State $packet \gets hw\_ring[ i ].slot( ring[ i ].curr ) $
+	\State $ring[ i ].avb \gets ring[ i ].avb - 1$
+	\State $ring[ i ].curr++$
+	\If{ $ring[ i ].avb < threshold$ } \label{r_pack_thres}
+		\State $hw\_ring[ i ].avb -= ring[ i ].( limit - avb)$
+		\State $hw\_ring[ i ].curr \gets ring[ i ].curr$
+		\State $ring[ i ].avb \gets hw\_ring[ i ].avb$
+		\State $ring[ i ].limit \gets hw\_ring[ i ].avb$
+	\EndIf
+	\State \textbf{return} $packet$
+\EndIf
+
+\State \textbf{\textit{issue ioctl and poll}}
+\State $ring[ i ].avb \gets hw\_ring[ i ].avb$
+\State $ring[ i ].limit \gets hw\_ring[ i ].avb$
+\EndFunction
+
+\end{algorithmic}
+\end{algorithm}
+
+\subsubsection{Packet Processing Operations}
+
+Packet processing operations consists of converting the raw data that the application provides into network packets suitable for transmission.
+Since the application tasks are represented as a set of operations on a stream of data we include these operations as stages through which the data flows through to get transformed into data packets.
+By integrating them with the application task graph it makes it easier to extract parallelism later in the execution engine.
+Several user space libraries exist that perform packet related processing operations like libnids \cite{libnids}, lwip \cite{Dunkels01designand}, click \cite{Kohler2000} etc., Since our application are targeted at a closed environment, we chose the click modular router for its flexibility and vast library of elements as a means of operating on the network packets.
+Our current framework supports a simpler udp protocol but click's modular infrastructure would make it easier to construct more complex ones.
+
+Each of the element in the click contains a virtual function which is called from the previous element it is attached to.
+We had to overcome this limitation of using click's infrastructure to construct based on dynamically initialized objects and virtual functions.
+This was necessary due to two important conditions, one was to eliminate the overhead associated with the dynamic construction of the objects and virtual functions and the other was to optimize the task graph along with these operations.
+To achieve this our framework constructs a simpler support structure that can link elements supported in click and import them to integrate with our application.
+
+\subsection{Parallel Processing Engine}
+
+After integrating the packet operations by means of support from \texttt{click}'s infrastructure and data access to network hardware through \texttt{netmap}'s APIs, we integrate these operations with the application graph, we now have a complete task graph that does both the computation and communication operations.
+The next logical step was executing these tasks in parallel.
+Two basic forms in which we can extract this lies in data and task parallelism.
+Packet streams have an inherent nature to operate in a data parallel fashion i.e.
+it is possible to operate on two different streams simultaneously.
+Since the application tasks in turn depend on the data packets they could also be executed in parallel.
+So from the integrated graph constructed by combining the application and packet processing operations it becomes possible to extract data parallelism by operating on different data streams simultaneously.
+In executing different graphs that represent distinct operations on the data streams it is possible to exploit task parallelism.
+
+\subsubsection{Execution engine}
+\label{ppe}
+
+The execution engine consists of set of process and threads that the work is scheduled upon in order to be executed.
+When implementing threads to execute code which consists of system calls to the OS such as \textit{malloc(), ioctl(), poll(),} etc., we found a degradation in performance even when the work that has to be executed is completely independent.
+To support the light weight style of threads the OS implements most of the locking mechanisms for system calls coming from threads.
+Previous research such as \cite{Lever:2000:MPM:1267724.1267780} \cite{Dice:2002:MLM:773039.512451} \cite{Michael:2004:SLD:996893.996848} shows the degradation in performance due to the usage of \textit{malloc()} in threads.
+Since our framework has to deal with system calls and issue them in parallel we use \textit{unix process} to perform the tasks concurrently.
+We do this by using \texttt{Message Passing Interface (MPI)} \cite{TheMPIForum:1993:MMP:169627.169855} to construct a set of processes that is scheduled on to the different cpus on a multicore system.
+
+In order to prevent interference from the OS scheduler's process migration, we specify \textit{affinity} of the MPI process on to particular cpu cores.
+This prevents our process from being executed on a different cpu other than that specified.
+This makes our processes act as execution engines running on each of the cores on which we schedule the operations that are to be performed.
+Several previous research\cite{Salehi:1996:EAS:234766.234769}\cite{1409136} show the importance of affinity in network sensitive applications.
+Next, from utilizing our \textit{Parallel Packet I/O} we specify the hardware queues each of these process handle.
+Initially we divide the queues evenly among the different processes we create, but later during execution we could modify the allocation based on the application operations scheduled on them.
+Our modifications to \texttt{netmap}'s hardware queue assignment API help us here in redistributing the queues to different processes.
+In using \texttt{MPI's send \& receive} inter-process communication calls it makes it possible to modify the attributes of the process such as the cpus assigned or the number of queues it handles and even modify the operations the processes were initially assigned.
+
+\subsubsection{Parallel task processing}
+\label{ppt}
+
+Once we have scheduled our execution engines on the physical system, the next step is to execute the integrated task graph consisting of application and packet operations on them.
+To extract data parallelism we replicate the task graph on to the multiple execution engines.
+This is to ensure operations on the same data are performed on the same cpu thereby reusing local cache.
+By making sure that packets belonging to specific flow always end up in the same execution engine, we only perform operations conforming to that stream of data.
+We leverage on the ability of modern network to classify packets, based on attributes of the packet, on to different queues.
+
+\lstset{language=C,caption={Packet Payload Modification Function},label=payload_sub}
+\begin{lstlisting}
+bool payload_copy( const void *payload_data, uint32_t length )
+{
+    uint32_t headroom = _data - _head;
+    uint32_t tailroom = _end - _tail;
+    uint32_t orig_length = _tail - _data;
+    uint32_t n = length + headroom + tailroom;
+    if( n < min_buffer_length ) 
+    {
+        tailroom = min_buffer_length - length - headroom;
+        n = min_buffer_length;
+    }
+    _data = _head + headroom;
+    _tail = _data + length;
+    if( payload_data )
+        memcpy( _data, payload_data, length );
+    return true;
+}
+\end{lstlisting}
+
+In order to replicate the task graph we had to overcome some restrictions.
+First was to ensure that there was no intricate data dependency between different data streams.
+Although, click's library of elements are designed to be executed in parallel \cite{Chen:2001:FCP:647055.759948}, it was more oriented towards being executed in a serial fashion.
+Hence we had to ensure the code present in the elements were thread safe.
+One important modification we had to perform was in the packet library that click uses.
+Click by default destroys and re creates the packet's internal storage buffer by reallocating memory space, which is an expensive operation.
+In order to avoid this overhead we allocate a packet with a large static buffer and copy the data to be modified on it.
+We modify the pointers that determine the header and data addresses based on the required packet length.
+This is done by using a simple sub routine shown in Listing \ref{payload_sub} that we added to the click's packet library.
+Even though this incurs the overhead of a copy operation it does avoid performing invalid and dangerous memory operations directly on the network hardware queue's buffer.
+
+Click's packet library does provide a mechanism to take ownership of a specific address and pass a destructor to be called when the packet is killed.
+To prevent corruption of memory mapped from the hardware queue by the application tasks we use the above mechanism to copy the data before modifying it.
+In generating the application graph as a sequence of operations it also makes it possible to run different graphs on each individual execution engine.
+The MPI inter-process communication is used to schedule the application and packet operations on the processes representing the execution engines.
+Further more, we use it to also perform explicit data sharing on tasks in the application that requires it.
 
@@
 
+The evaluation of our system should test the scalability and the performance speedup that we could achieve on a multicore system for an application that is constructed in our framework.
+To do this we choose two generic operations such as compression and encryption, built simple stream applications based on them in our framework.
+We then measured the improvement in performance as we increased the number of execution engines that are dedicated for the application.
+
+\subsection{Hardware Setup}
+
+The multicore system that is used to evaluate the framework consists of two servers connected directly to each other.
+Each of the server consists of two Intel Xeon 5600 hexacore HT in a dual socket configuration with 24GB DDR3 memory.
+So there is a total of 2 Processors x 6 Cores x 2 HT = 24 cpus in each server.
+These servers are connected with each other using Intel 8299EB 10Gbe adapters over direct attach twinaxial cables.
+They are placed in PCI Express x8 slot in each system to maximize the available bandwidth for the cards.
+This forms a closed connection between two servers and limits the influence of any external devices such as switches or routers on the actual data transmission.
+
+The application is constructed as two parts, one acting as a source of data and other a sink.
+We measure the overall time it takes for application task operations, packet operations and the time to send from the source and receive at the sink.
+This is done for different packet sizes and applications.
+Then we plot the speedup achieved in each of the cases over the serial version where we have a single sender and receiver operating on the multiple queues.
+This is considered to be the baseline in our experiments since this is similar to how the OS behaves for transmitting and receiving using the multi queue network controller.
+
+The Intel 82599EB has limitations on the number of queues it can use for Receive Side Scaling (RSS) which is restricted to 16.
+So this limited our maximum amount of processes at the sink to be 16 even though there were 24 cpus available on the system.
+The source is not affected by this limitation and can scale to all the 24 cpus available on the system with the presence 24 hardware tx queues.
+Overall we at least have one hardware queue mapped to each of our processes or execution engines.
+We also ensure that each of the processes are bound to a particular cpu by issuing affinity call to the OS to ensure that we are not migrated between different cores.
+
+\subsection{Application Construction}
+\label{app-const}
+
+In order to evaluate our framework we needed distinct operations that would be suitable to be applied in a streaming framework.
+Since compression and encryption applications are used commonly we chose two applications that represent these operations.
+The integrated graph that represents both these applications are shown in fig \ref{fig-app-graph}.
+Since our intention is to test the overall performance of the stream applications we use simple protocol such as udp in order to send data between the source and the sink.
+
+\begin{figure}[htb]
+\centering
+\includegraphics[width=3.5in]{enc-dec}
+\caption{ Integrated Task Graph  }
+\label{fig-app-graph}
+\end{figure}
+
+Both the applications are constructed in a similar way.
+Two data sizes of 1.28 GB and 2.56 GB are selected as the workload for the compression and encryption applications respectively.
+Openssl library is used for the encryption and decryption and zlib is used for the compression and decompression operations respectively.
+The application consists of data generation phase after which it is compressed or encrypted depending on the application after which the packet operations are done.
+At this point the stream consisting of packets is sent over the network using the multiple queues.
+The sink side does the reverse of these operations.
+Once the data is received the packet operations are performed to extract the payload, after which the decompression or decryption operation is performed to retrieve the original data.
+These are some of the basic operations that would be required to construct a much more complex application.
+In choosing to compare these simple operations we avoid any limitations that might be present in a more complex application.
+This would enable us to evaluate the performance due to our framework and not that of any optimization done in the application.\\
+
+\begin{figure*}[tb]
+\centering
+\includegraphics[width=7in]{result}
+\caption{Performance Speedup; (a),(b),(c) - Application A.
+Encryption; (d),(e),(f) - Application B.
+Compression;\\ \newline Speedup based on time in takes to compute a fixed workload for different number of source and sink processes\\ \newline and for varying packet sizes over a single source and sink are shown above}
+\label{fig-res}
+\end{figure*}
+
+\section{Results}
+
+We see that there is considerable difference in the way each of the operations behave as we scale them.
+This is shown in Figure \ref{fig-res}.
+We can make three distinct observations from the results.
+\\
+
+One is that from fig\ref{fig-res}.(d,e,f) we find that the compression requires a lot more work than that of decompression, whereas from fig\ref{fig-res}.(a,b,c) encryption and decryption seems to be more or less balanced.
+This is evident from the continued speedup we get up to 4 senders in the case of compression \& decompression.
+In encryption \& decryption this trend is not seen and the performance only improves by increasing both the number of senders and receivers.
+So we conclude that the combination of parallel tasks for application and packet operations differs between different applications.
+Determining this for each scenario is essential to boost the performance of the overall system.\\
+
+We see that the packet sizes have some impact on the performance from fig\ref{fig-res}.(d,e,f) but there is not much change compared to the overall speedup we achieve in each of the case.
+It can also be noticed that the compression operation seems to be influenced more than that of encryption operation.
+There seems to be speedup of 12x from fig\ref{fig-res}.(d) in the case of smaller packets and 10x from fig\ref{fig-res}.(f) for larger packets for compression operation.
+Whereas in the encryption application there is not much change and we get a speedup of 8x for any packet size observed from fig\ref{fig-res}.(a,b,c).
+While, packet sizes seem to have a much larger effect when there are smaller number of senders or receivers, but for larger cases not much change is observed in fig\ref{fig-res}.(d,e,f).
+So we conclude that the influence of a parameter for the network system such as a packet size varies between different applications and it is essential to have a unified system that can tackle both the operations together.
+
+From fig\ref{fig-res}(a,b,c,d,e,f), we noticed the drop in speedup as we move from 12 senders to more.
+This could be due to the influence of NUMA architecture \cite{Awasthi:2010:HPO:1854273.1854314} on the performance of the system.
+This is the particular point in which we utilize all the cores of a particular cpu and move on to a different socket.
+Mainly, the combination of cache locality and limitation of the network controller to receive data through the PCI-E bus from both the processors could be the contributing factors.
+Still, we find that even in a NUMA architecture once our framework overcomes the initial drop, it continues to improve and eventually attain better performance in most cases.
+
+We can see that our framework scales very well for the different workloads provided and can handle both the computations and communication and operate on them in parallel.
+In some cases we find that our framework could be improved by having a much more control over the fine grained parallelism within the execution engines.
+This could be done by introducing pipeline stages for the different operations.
+
 \section{Related Work}
 
+Several prior research \cite{Dobrescu09routebricks:exploiting}\cite{Han:2010:PGS:1851275.1851207}\cite{Wang:2009:PPN:1542275.1542307}\cite{springerlink:10.1007/s11227-011-0579-3} have worked on using commodity multicore processors to perform network operations.
+
+Dobrescu et al.\cite{Dobrescu09routebricks:exploiting} shows the performance improvement achieved in the construction of software routers by distributing the related operations over multiple servers.
+
+Han et al.
+\cite{Han:2010:PGS:1851275.1851207} perform tasks related to high performance packet processing on commodity hardware using GPUs.
+They have developed a novel framework know as PacketShader that can do packet processing in user space by exploiting massively-parallel architecture of GPUs.
+
+Wang et al.
+\cite{Wang:2009:PPN:1542275.1542307} constructs a high performance connection-affinity based lock-free multicore network processing system that claims to achieve multiple Gbps network processing speed for complex tasks.
+
+Egi et al.\cite{springerlink:10.1007/s11227-011-0579-3} shows techniques on parallelizing network operations on multicore architectures by defining principles that consider multiple resources and not just the CPU alone.
+
+Stream programming languages such as StreamIt \cite{Thies:2002:SLS:647478.727935}, LUSTRE \cite{Halbwachs91thesynchronous}, etc., target stream applications but they focus more on parallelizing the computations related to the application alone on the multicore system.
+
+\comment{
+All of these applications are more focused on operating as a network application and primarily focus on parallelizing packet operations.
+In our work even though a part of it involves parallelizing network operations, the objective of this research is to provide support for parallelizing stream application and distributing their workloads over multiple systems.
+This also requires optimizing the underlying network operations by parallelizing it.}
+
+\texttt{Netmap} \cite{Rizzo:2012:RNI:2090147.2103536}, provides the APIs to access the hardware queue's buffer in user space by means of memory mapped rings.
+It implements an intermediate ring in kernel space that is used to synchronize the data from the user space to that of hardware queues in the NIC.
+
+Click modular router \cite{Kohler2000} is an extensively used tool for the construction of software routers using general purpose hardware and provides a massive infrastructure to support this.
+SMP Click \cite{Chen:2001:FCP:647055.759948} is an improved version, which supports parallel processing of these using threads.
+Click implements its own mechanism and polling driver in kernel space to access packets, bypassing the operating system's network infrastructure.
+
+\comment{
+In order to send data the programmer calls \textit{poll} on the tx ring in user space and writes to it when space is available.
+Then he calls \textit{ioctl} to synchronize the data with the kernel ring and transmit the data.
+To receive the packets \textit{poll} is again called on the rx ring in user space but this time to check for new data and after reading it, he issues \textit{ioctl} to let the driver know that packet has been read and space can be re used.
+
+, our framework uses techniques as described in Section \ref{imple} to integrate them with the application and access them in parallel from different application processes.
+Also, while some of click's \cite{Kohler2000} elements are used to provide support for packet processing operations, in our framework we use it for doing some of application's network tasks alone.
+Due to this we modified a subset of click's elements in order to support our system.
+
+ In our framework we target a different category of stream applications such as data mining, real time image analysis, etc., which tend to compute on large amounts of data.
+In order to improve overall performance in such cases, the operations are distributed over multiple systems.
+Our framework provides an easier methodology to construct such an application while managing both the parallelizing and communication operations.
+
+Our framework provides the necessary functionality to target communication intensive stream applications.
+We focus on improving the overall performance of such an application by integrating the network and application operations.
+We do this by providing a high-level programming framework that can be used to construct these applications, which can extract parallelism and improve communication performance.
+Using this it would be possible to develop stream applications that could be distributed over multiple systems and using our framework it would be possible to handle the highly intensive communication between them.}
+
 \section{Conclusions}
+In this paper we have proposed a high-level programming framework that can integrate packet processing within stream applications and perform both the operations in parallel.
+
+We conclude that by using our algorithms for sending and receiving data, which implicitly performs batching and reduces the usage of system calls, and in combining this with \texttt{netmap} APIs which minimizes the interaction with the operating system, communication performance in our framework never becomes the bottleneck in our system on the programs we have tested.
+Instead of performing the packet operations in the OS, by integrating it with the application using click's elements within our infrastructure we are able to construct a unified task graph that represents the computation and communication operations.
+Using this we are able to parallelize these operations together and achieve speedups of more than a factor of eight in all the applications we tested.
+
+In integrating these features into a programming framework, the system of multiple queues, parallel communication and computation are entirely hidden from the programmer, who merely specifies a standard stream graph for computation alone.
+The results shows that our system scales well to as many as 24 parallel processes on a multicore computer, and achieves speedups of more than a factor of ten in some cases compared to sequential versions of the stream applications.
 
@@
 \section*{Acknowledgment}
-This work is funded by the IRCSET Enterprise Partnership Scheme along with IBM Research Ireland.
+This work is funded by the IRCSET Enterprise Partnership Scheme in collaboration with IBM Research, Ireland.
 
@@
 % http://www.michaelshell.org/tex/ieeetran/bibtex/
-%\bibliographystyle{IEEEtran}
+\bibliographystyle{IEEEtran}
 % argument is your BibTeX string definitions and bibliography database(s)
-%\bibliography{IEEEabrv,../bib/paper}
+\bibliography{IEEEabrv,ref.bib}
 %
@@
 % (used to reserve space for the reference number labels box)
-\begin{thebibliography}{1}
+%\begin{thebibliography}{1}
 
-\bibitem{IEEEhowto:kopka}
-H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
-  0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.
-
-\end{thebibliography}
+%\end{thebibliography}
 

### 2013-03-18 script.sh

@@ -1,5 +1,10 @@
 #!/bin/sh
 while inotifywait paper.tex; do
 #    pdflatex -interaction=nonstopmode paper.tex
-     make
+#     make
+    pdflatex -interaction=nonstopmode paper.tex
+    bibtex paper
+    pdflatex -interaction=nonstopmode paper.tex
+    pdflatex -interaction=nonstopmode paper.tex
+    rm -rf *.log *.blg *.bbl *.aux
 done

### 2013-03-20 paper.tex
Kevin's changes

--- 2012-06-26 paper.tex
@@
 
+\usepackage{listings}
+\usepackage{color}
+
+\definecolor{dkgreen}{rgb}{0,0.6,0}
+\definecolor{gray}{rgb}{0.5,0.5,0.5}
+\definecolor{mauve}{rgb}{0.58,0,0.82}
+
+\lstset{ %
+  language=C,                % the language of the code
+  basicstyle=\footnotesize\ttfamily,           % the size of the fonts that are used for the code
+  numbers=right,                   % where to put the line-numbers
+  numberstyle=\tiny\color{gray},  % the style that is used for the line-numbers
+  stepnumber=0,                   % the step between two line-numbers. If it's 1, each line
+                                  % will be numbered
+  numbersep=0pt,                  % how far the line-numbers are from the code
+  %backgroundcolor=\color{white},      % choose the background color. You must add \usepackage{color}
+  showspaces=false,               % show spaces adding particular underscores
+  showstringspaces=false,         % underline spaces within strings
+  showtabs=false,                 % show tabs within strings adding particular underscores
+  %frame=single,                   % adds a frame around the code
+  rulecolor=\color{black},        % if not set, the frame-color may be changed on line-breaks within not-black text (e.g. commens (green here))
+  tabsize=2,                      % sets default tabsize to 2 spaces
+  captionpos=b,                   % sets the caption-position to bottom
+  breaklines=true,                % sets automatic line breaking
+  breakatwhitespace=false,        % sets if automatic breaks should only happen at whitespace
+  title=\lstname,                   % show the filename of files included with \lstinputlisting;
+                                  % also try caption instead of title
+  keywordstyle=\bfseries,          % keyword style
+  commentstyle=\color{dkgreen},       % comment style
+  stringstyle=\color{mauve},         % string literal style
+  escapeinside={\%*}{*)},            % if you want to add LaTeX within your code
+  morekeywords={*,...}               % if you want to add more keywords to the set
+}
+
 % *** SPECIALIZED LIST PACKAGES ***
 %
+\usepackage{algpseudocode}
+\usepackage{algorithm}
 %\usepackage{algorithmic}
+%\usepackage{algorithmicx}
 % algorithmic.sty was written by Peter Williams and Rogerio Brito.
@@
 
-%\usepackage[caption=false]{caption}
+\usepackage[font=small,center,belowskip=-10pt,aboveskip=5pt]{caption}
 %\usepackage[font=footnotesize]{subfig}
@@
 %\boldmath
-the abstract goes here.
+Stream applications are often limited in their performance by their underlying communication system.
+In a typical implementation, the Operating System (OS) handles the majority of network operations.
+In such cases, the communication stack, which was not designed to handle such tremendous amounts of data, acts as a bottleneck and restricts the performance of the application.
+Therefore, a framework for computing and communicating in parallel is essential for improving performance.
+
+We propose a high-level programming framework that integrates the packet processing operations with stream applications, and provides a common interface to execute both application and packet processing operations in parallel on multicore processors.
+
+We utilize the ability of the network interface controller (NIC) to classify packets onto multiple hardware queues.
+This is combined with system software that allows direct access to these queues.
+We develop an interface in our framework for application processes to read or write in parallel.
+Our system integrates this with the application by a series of packet processing operations that transform the data from the application to packets suitable for transmission.
+Representing these tasks as stages in the application, we are able to parallelize them using techniques similar to those used for parallelizing regular computations.
+
+All the communication operations in the framework are performed in userspace, bypassing the OS, to send and receive data directly over the network interface.
+By overcoming the OS bottleneck, and by parallelizing computations and packet operations together, we are able to boost the performance of stream applications substantially.
 
@@
 \IEEEpeerreviewmaketitle
+%Kevin - vertical spacing
+\setlength{\textfloatsep}{15pt}
+\section{Motivation}
 
-\section{Motivation}
-In the continued race for improvement of multi core and many core architectures, computer networks are becoming a fundamental area for performance bottleneck especially in systems which are used for distributed processing of parallel applications.
-In order to handle the requirements posed by these applications the interconnects are becoming more and more complex.
-Such systems utilize a wide range of custom interconnects ranging from infiniband, myrinet to the more readily available ones such as 1Gbe and 10Gbe Ethernet standards and sophisticated routers and switches to handle the added challenge in routing data.
-While a small group of the specialized systems employ custom routers and topology configurations the majority of them especially those that use 1Gbe or 10Gbe use the commercial routers.
-In addition the growing complexity of routing protocols has resulted in advanced switches such as OpenVSwitch \cite{openvswitch}, etc., has resulted in these routers being upgraded frequently to keep up with the growing requirements.
-Software routers such as click modular router have been gaining popularity due to their flexible nature and cheaper alternative to the more expensive commercial routers and are widely used to prototype newer standards.
-In using software routers in the above scenario it would be possible to not only benefit from the newer developments in switching and routing protocols but it would require little effort to implement it in existing systems.
-One might argue that it would be a challenge to handle such complex workloads in software routers but with the presence of multicore and many core architectures and stream processors it is feasible to parallelize the packet processing workloads to improve the overall performance of the system.
-Several research projects are working on improving the performance by parallelizing the packet processing tasks on these architectures \cite{routebricks} \cite{packetshader}.
+Stream applications are designed in a manner where they operate on data that flows as a \textit{stream}.
+This data stream passes through different stages of computations, to be transformed from query to result \cite{Thies:2002:SLS:647478.727935} \cite{Halbwachs91thesynchronous} \cite{Stephens95asurvey}.
+This differs from a conventional computation model where data is stored and then processed.
+By eliminating the need to store data, and processing it on the go instead, it is possible for stream applications to process large amounts of data.
 
-In a computing cluster which is used in the execution of parallel applications it is common to dedicate a core or processor in each node for operating system related tasks in order to minimize its effects on the application.
-Similarly, it should also be possible to utilize an additional core or processor to perform packet processing operations.
-While this is would be a challenging scenario it does provide the added benefits of a cheaper alternative to commercial routers.
-Also in bringing the network system into a software controlled domain it should be possible to perform additional optimizations from the application such as prioritized data flow, intelligent routing, etc., that would otherwise be difficult to implement due to the abstraction of the network system.
+Stream applications such as file storage systems, database query management, data mining, etc., operating on vast amounts of data, are often distributed over multiple systems.
+Traditional stream applications target multicore processors for performing their computations in parallel.
+As multicore processors grow in power, so do stream applications, processing ever-increasing amounts of data.
+However, stream applications that operate on large amounts of data at a very high bandwidth can be limited by the communication stack of the OS kernel\cite{Wu:2007:PPB:1323954.1323957}\cite{Wu:2007:PAL:1227865.1228021}\cite{Dobrescu09routebricks:exploiting}\cite{Han:2010:PGS:1851275.1851207}\cite{Kohler2000}.
+With the continuing improvement of multicore and many-core architectures, it is becoming essential to improve the processing speeds of packet operations along with those of the application in order to scale performance.
+\comment{
+When performing packet processing operations for communication intensive stream applications, a major challenge is the high levels of performance required.
+These applications process large numbers of data streams in real time.
+If the communication system cannot keep up with the rate at which packets arrive, then they must be dropped or remain unprocessed.
+}
 
-The focus of this paper is the development of a parallel processing engine that could be integrated with the application and can perform both the actual computations and the packet processing related tasks in parallel.
-The entire system can be divided into key parts as follows,
+In building a framework for such applications on general purpose computers, we can exploit two hardware features that could allow for dramatically improved performance, namely multicore processors and multi queue network controllers.
+Using these features makes it possible to do much of the packet processing in parallel using the multiple cores commonly available on modern processors.
+
+A major bottleneck for packet processing on general purpose computers is the overhead of the operating system's network stack.
+In particular, for applications that run as user processes, data must be copied between OS kernel space and user space.
+Previous research uses alternate mechanisms to access packets from the network hardware directly to overcome this problem\cite{Dobrescu09routebricks:exploiting}\cite{Han:2010:PGS:1851275.1851207}\cite{Kohler2000}.
+Most of these use modern network hardware in combination with particular system software that allows direct access to hardware-controlled packet queues, dramatically reducing overheads and improving performance.
+However, constructing stream applications that use these features is difficult.
+Furthermore, many stream applications have a very specific communication pattern, typically operating on data between a source and sink in the form of a \textit{data stream}.
+Data in a stream must be processed sequentially, whereas those from separate streams can often be processed in parallel.
+
+Modern network hardware typically has a mechanism for separating arriving packets into multiple queues, where packets that belong to the same flow are\comment{ guaranteed to be} placed in the same queue\cite{micro2008} \cite{intel2010}.
+We can utilize this feature to operate on multiple streams of data by processing each of the queues in parallel.
+However, building such applications is complex because it requires explicit parallel programming and interaction with low-level network interfaces.
+Research focuses on improving the performance of network applications by parallelizing the packet processing tasks on multicore architectures \cite{Dobrescu09routebricks:exploiting} \cite{Han:2010:PGS:1851275.1851207}.
+We parallelize both the application and packet operations to gain significant improvement in performance.
+
+\section{Contribution}
+
+In this paper we propose a high-level programming framework for building communication intensive stream applications.
+We focus on integrating the packet processing and application tasks into a single graph and using this to extract data parallelism.
+The developer describes their program as a graph of \textit{actors}.
+Each actor consumes one or more \textit{tokens} of data, which may be simple data such as integers, or more complex data such as packets, arrays, or objects and produces one or more tokens.
+The framework can be used to automatically map the application over multiple parallel cores, and interface with the \texttt{netmap} APIs \cite{Rizzo:2012:RNI:2090147.2103536} for accessing data through the network buffers.
+
+The system can be divided into three main parts:
+
+\begin{itemize}
+\item A subsystem for interfacing with the network hardware through the \texttt{netmap} APIs to send and receive data from the application in user space without going through the operating system.
+(Section \ref{ppio})
+\item Given a graph that represents the workload to be performed by the application, we construct and integrate the packet processing operations represented as a directed acyclic graph (DAG) to it.
+(Section \ref{ppe})
+\item A run-time system that executes these tasks in parallel and explicitly schedules them on the multicore processor. (Section \ref{ppt})
+\end{itemize}
+
+% Even though we integrate the system with the application, the techniques for extracting parallelism from the application is not in scope of this paper.
+The application is considered to be in a streaming model and the tasks corresponding to the various functions are considered to be parallel in nature.
+
+\section{Background}
+
+In this paper, we are concerned with applications that operate on data streams.
+Example operations of such an application are represented in the stream graph shown in figure \ref{fig1}.
+This sample application operates by performing encryption on the data that matches predefined rules like file names, access level of user, etc., otherwise it performs compression.
+Due to the nature of such stream applications, it is possible to distribute the component operations across different CPUs.
+We are particularly interested in a scenario where groups of networked servers are used to perform these kind of stream operations by distributing them evenly over different nodes.
+
+\begin{figure}[ht]
+\centering
+\subfigure
+{
+	\includegraphics[width=3.5in]{ip-data-proc}
+}
+
+\subfigure
+{
+	\includegraphics[width=3in]{op-data-proc}
+}
+\caption{Sample stream application}
+\label{fig1}
+\end{figure}
+
+The main challenge that arises in such a situation is how best to address the intensive computational and communication demands of such applications.
+Two important features enable us to tackle this challenge.
+Firstly, the operations of both the application and the network are parallel in nature.
+Secondly, multicore processors can be used to execute the application.
+
+In our system, we use an efficient parallel processing framework that could be leveraged on multicore processors and multi queue network controllers to achieve efficient communication.
+The performance of multicore processors continues to dramatically improve the capability of commodity hardware, thus making it possible to meet the computational demand of such applications.
+In addition, multi queue network interface controllers, designed to handle multiple streams of data, help achieve the necessary communication performance.
+
+An important part of our system is the tasks related to packet processing in protocols such as UDP, IP, etc., We utilize the modular router, Click, for these packet processing tasks.
+Click is a tool that includes a library of elements which is used in building flexible, re-configurable and modular software routers \cite{Kohler2000}.
+It offers a high-performance platform that is flexible for rapid development of newer network protocols.
+An improved version known as SMP Click supports thread-based parallel processing of network related tasks\cite{Chen:2001:FCP:647055.759948}.
+Until recently, Click's performance in the user space was inferior to its kernel space equivalent.
+This is mainly due to its extensive use of polling driver in the kernel space that was not possible in user space.
+Netmap I/O\cite{Rizzo:2012:RNI:2090147.2103536} provides a set of APIs that can directly access the packets from the NIC and was used to improve Click's user space performance.
+Although other other user space packet-access APIs could be used \cite{Rizzo:2012:RNI:2090147.2103536} \cite{1564468} \cite{Krasnyansky}, the simpler construction of \texttt{netmap} was more suitable to adapt to our framework.
+
+Section \ref{dppl} represents the design principles that we follow to conceptualize our framework.
+Section \ref{imple} describes the implementation details used in the construction of the system.
+In section \ref{ppe}, we discuss the challenges in using a system consisting of execution engines to manage the work.
+The extent to which we extract parallelism ranges, not only to the packet tasks, but also to the low level access of data I/O to the network interface hardware through custom additions to the \texttt{netmap}.
+Finally in section \ref{app-const}, we present a simple stream application constructed from operations such as compression and encryption in order to evaluate our framework.
+
+\section{Design Principles}
+\label{dppl}
+
+The design of a high-level programming framework that integrates packet processing tasks into the application requires overcoming several challenges.
+\comment{The entire system is driven by the extent to which parallelism can be extracted.
+Some of the techniques that can be applied to the packet processing tasks can be used for the tasks related to the application.} In this section, we describe these challenges and the principles that are used to address them.
+\comment{We focus on the methodologies related to the packet processing tasks and the parallel processing engine itself.}
+
+\subsection{Userspace Packet I/O}
+
+For a packet processing system to be integrated with an application, it is necessary to have an efficient mechanism for accessing the data from the NIC without interacting with the operating system.
+Rapid data access here is crucial for the extraction of data parallelism at later stages.
+\comment{In using the techniques described below we exploit the modern NIC architecture to its full potential.}
+
+\subsubsection{Handling Multiple Hardware Queues}
+
+The development of virtualization and improved network flow handling has led to hardware level multi queue support in NICs.
+During transmission, concurrent writes to different queues are possible, enabling multiple processes to send data simultaneously.
+During reception, the NIC classifies each packet onto one of the receive queues in a technique known as Receive Side Scaling (RSS)\cite{micro2008} \cite{intel2010}.
+This technique uses the header information or tuples to classify packets onto different hardware queues.
+Multi queue support can handle several streams of data in parallel and it is exploited extensively in our framework for this reason.
+By offloading the classification of packets to the NIC, we remove any overhead associated with packet classification in the application.
+We can also use these hardware queues as a means of balancing the workload across different threads by controlling the number of queues that each process handles.
+\comment{Since different processes can access these transmit and receive queues, concurrent access to data is possible.}
+
+\subsubsection{Flow based Packet Batching}
+
+Packet batching is a common method used to boost the performance of network applications where several packets are combined and operated upon together, thus reducing the overhead of I/O operations.
+In our framework, batching is utilized in a form where the number of packets in a batch is varied based on the available space in the hardware queue.
+This number reflects the amount of data we are able to process before issuing hardware synchronization calls and forms an important means by which we reduce the usage of system calls.
+
+\subsubsection{Reduction of Per Packet System Calls}
+
+System calls are required in order to synchronize data between the NIC hardware and the buffer memory used for intermediate storage of packets.
+Previous work describes the overhead associated with per-packet system calls and provides mechanisms for overcoming this overhead\cite{1564468}\cite{Han:2010:PGS:1851275.1851207}\cite{Rizzo:2012:RNI:2090147.2103536}.
+
+We employ an \textit{adaptive scheme} described in section \ref{ppio} that utilizes this strategy of reducing per-packet system calls.
+Our framework balances the system calls based on the communication rate and application processing speed.
+The transmission and reception are handled differently to optimize the usage of system calls in each of the cases independently.
+During transmission, we issue system calls to synchronize the available buffer space based on the rate at which we transmit data.
+This prevents degradation in latency for larger packets and improves performance for smaller packets.
+During reception of packets, the system reads packets at the rate at which it is able to process them.
+
+\subsection{Packet Processing Graph}
+
+\comment{Understanding the set of packet processing tasks associated with the application is essential to performing them in parallel.} We use a graph-based representation of packet processing tasks to determine the dependencies involved for the different stages and to determine the relation with the different tasks of the application.
+This graph contains the stages through which the data associated with the application is transformed into network packets and vice versa.
+It shows which of the tasks can be performed in parallel with those of the application, and highlights those that have a data dependency on a specific task in the application.
+The graph also enables optimizations based on specific characteristics of the application or the architecture used for execution.
+Figure \ref{fig2} shows a graph representing User Datagram Protocol (UDP) packet generation.
+
+\begin{figure}[ht]
+\centering
+\includegraphics[width=3.5in]{pack-op}
+\caption{Simplified UDP packet construction}
+\label{fig2}
+\end{figure}
+
+\subsection{Application Task Graph}
+
+\comment{Application task graph represents the operations associated in performing the computations.} The subset of the application that we are interested in parallelizing is the stream processing component.
+In such components, the communication flow between tasks is well-defined.
+Both the packet processing and application tasks can be represent by a directed acyclic graph (DAG) similar to the packet processing graph.
+The packet processing tasks operate on packets and the application tasks operate on the data.
+Since we focus more on parallel packet processing, optimizations in relation to the application graph are not discussed.
+It is assumed that a static version of the application graph is available and we focus more on executing these tasks in parallel with the packet processing tasks.
+
+\subsection{Integration}
+
+The sequence of tasks that can be executed in parallel is determined at this stage.
+The integration stage is used to combine the application graph and the packet processing graph into a larger graph that unifies and represents the overall operations.
+The dependency relations identified earlier are used in establishing links between the graphs.
+We determine the sequence of tasks that can be executed in parallel from both the application task graph and packet processing graph.
+The combined set of tasks is then used by the parallel processing engine.
+
+\subsection{Parallel Processing Engine}
+
+Two important aspects that the parallel processing engine must address is the scheduling of tasks and balancing the load across the available resources.
+The parallel processing engine consists of collaborating processes which are bound to specific CPUs and act as engines to execute the workload.
+Based on their requirements and the available resources, tasks are scheduled across these execution engines.
+\comment{In this manner, the workload can be distributed flexibly.}
+
+\section{Implementation}
+\label{imple}
+
+\comment{Most extant stream applications depend on the OS for their communication tasks.}In commodity systems, the interaction with the OS for network operations is a necessary abstraction to handle multiple applications.
+However, in the case of specialized servers which cater to specific applications, it would be beneficial to interact directly with the network device.
+Moreover, in the case of intensive network traffic, the OS could act as a bottleneck due to the overhead associated with data flowing through the kernel before being transmitted through the network device\cite{Wu:2007:PPB:1323954.1323957}\cite{Wu:2007:PAL:1227865.1228021}.
+Eliminating this would be possible, either by executing the application in the kernel space (which could compromise stability of the system), or by using an interface for accessing the NIC buffers from user space.
+The latter approach is supported by several interfaces such as \texttt{netmap} \cite{Rizzo:2012:RNI:2090147.2103536}, PF\_RING DNA \cite{1564468}, UIO-IXGBE \cite{Krasnyansky}, etc., Even though these interfaces support user space packet access, they lack the ability to be invoked in a efficient manner by the application.
+More specifically, they do not have any way of optimizing the rate of processing, lacking, for example, the concept of flows for application tasks.
+The lack of support for interacting with these interfaces in order to read and write data in parallel from concurrent processes or threads adds to the problem.
+Also, since the application is now responsible for handling network packets, it has to perform the additional operations that would otherwise be carried out by the OS.
+
+Our framework provides support for sending and receiving data in parallel over the NIC from user space and a common parallel execution engine that supports both application and packet processing tasks.
+
+\subsection{Parallel Packet I/O}
+\label{ppio}
+The parallel packet I/O interface, when combined with the application tasks, provides concurrent communication for the application threads or processes over the network.
+To do this, two important requirements have to be met.
+Firstly, access to network data from user space should be provided, which can be read from and written to in parallel.
+Secondly, it should be possible to integrate the packet operations with the application by providing the necessary interfaces for performing these packet operations, thereby allowing for some flexibility in the choice of communication protocol.
+
+\subsubsection{Userspace Packet Access}
+Applications executed in kernel space for accessing hardware devices directly would lead to system instability.
+Instead, previous researchers \cite{Rizzo:2012:RNI:2090147.2103536}, \cite{1564468}, \cite{Krasnyansky}, etc., propose access to network buffers from user space.
+We use \texttt{netmap} APIs \cite{Rizzo:2012:RNI:2090147.2103536} to access data from the network buffers.
+\texttt{Netmap} uses a set of ring buffers in kernel space and user space to export the network hardware buffer's memory region.
+This enables the applications in user space to directly access the network hardware.
+It also prevents the application from accessing memory outside valid regions by checking the limits of the ring buffer.
+Packets can be written to the buffers when space is available and an \textit{ioctl} call to the driver enables the network controller to send the data.
+A \textit{poll} interface is implemented on the memory region mapped in user space to check when new packets are available.
+On reading or writing packets, the available space is updated.
+
+We leverage the availability of multiple hardware queues present in modern NICs to read and write data concurrently.
+\texttt{Netmap}, by default, supports assigning each hardware queue to an individual application process or thread.
+However, in order to allow for more efficient communication in the framework, we required the ability to assign more than one queue to a particular thread or process.
+Our modifications to \texttt{netmap}'s hardware queue assignment API allow our framework to balance the amount of packets handled by each process by varying the number of queues assigned to it.
+
+During testing, it was found that it was essential to have control over the rate at which system calls were issued with respect to the amount of data being transferred.
+Frequent system calls hamper bandwidth and limit the peak performance of the application.
+To tackle this, we propose the following algorithms for reading and writing data from the hardware queues using the \texttt{Netmap} APIs.
+These algorithms are designed to provide the following features,
 
@@
 
-\item A suitable methodology by which we can access the data in the userspace without interacting with the operating system.
-To do this (1)We need to read the data from the Network Interface Controller (NIC) before the OS and (2)Using this, establish a efficient manner for sending and receiving the data.(Section \ref{sec:userspace-packet})
-
-\item Given a graph that represents the workload to be done on the packet flow, we divide it into tasks that can be executed in parallel and determine the regions that could be used to extract data and task parallelism.(Section \ref{sec:workload})
-
-\item The parallel processing engine that utilizes the above and does the actual computations on the multi core or many core processor.
-It also handles load balancing and scheduling of the work across the available computing resources.
-(Section \ref{sec:ppe})
+\item Provide batching of packets based on available space on hardware rings
+\item Reduce the usage of system calls such as \textit{ioctl} and \textit {poll}
+\item Provide mechanisms for handling multiple queues through the application
 
@@
 
-We concentrate more on the framework that can integrate the packet processing and application tasks into a parallel processing engine and the extraction of parallelism from the packet processing related tasks to improve the overall performance.
-Given an application graph and its partition consisting of parallel tasks we the combine it with the packet processing operations.
-In addition, Section \ref{sec:} we show the challeneges involved in using a system consisting of threads and process that are used to actually perform the work.
-The extent to which we extract parallelism ranges not only to the packet tasks but to the low level access of data I/O to the network interface hardware through certain additions to the netmap APIs.
-Finally in Section \ref{sec: } we present certain pseudo applications that involves operations such as compression and encryption to evaluate our parallel processing engine.
+\comment{By minimizing the synchronizations with the hardware rings, we essentially limit the system calls issued and consequently improve the overall system performance.}
 
-\section{Background}
+Algorithms \ref{s_pack} and \ref{r_pack} use a batch update for sending and receiving packets based on the rate at which data flows through the interface.
+This is done by maintaining a data structure that has information about the hardware rings, updating it on a periodic basis.
+Once a fixed amount of space is determined, we update this on our local structure.
+We then continue to read or write data from the hardware ring, but update the actual limits and position on our local structure.
+When we have used up the space that was allocated, we update the hardware ring with the changes.
+In cases where the application is generating sparse traffic, it is possible to frequently update the ring in order to maintain a consistent latency in transmission.
+Multi queue support is provided by managing the process of selecting the ring to send or receive based on an application determined function.
+The $f(n)$ present in Algorithm \ref{s_pack} line \ref{s_pack_param} and Algorithm \ref{r_pack} line \ref{r_pack_param} represents this.
+Choosing a random queue has proven to be effective in our experiments, but if an application requires a specific flow to be maintained, a different function here could provide that support.
+This would be helpful in cases where the data operations are more predictable, or if the transactions seem to follow a specific pattern.
+The $threshold$ specified in Algorithm \ref{r_pack} line \ref{r_pack_thres} is determined for each application based on the rate at which it is able to process data.
+By issuing the system calls such as \textit{ioctl} and \textit{poll} only when required, and effectively handling the multiple queues, we are able to achieve improved performance.
 
-An important part of our system is the tasks related to packets such as routing, forwarding, etc., Since our design focuses more on the aspects of the parallel framework we utilize the click modular router for the packet processing tasks.
-Click modular router \cite{click} is a commonly used software router that includes a library of elements which is used in building flexible and reconfigurable routers.
-It offers a platform that is flexible for rapid development of newer network protocols and achieves good performance.
-An improved version known as SMP Click \cite{smpclick} also supports thread based parallel processing of network related tasks.
-Until recently click's performance in the user space was not as good as its kernel space equivalent.
-This is mainly due to its extensive use of polling driver in the kernel space.
-Netmap I/O \cite{netmap} provides a set of APIs that could directly access the packets from the NIC and it helped in the recent improvement of click's user space performance.
+\setlength{\textfloatsep}{0pt}
 
-Click dynamically constructs routers in which the configuration file is passed as a parameter for the kernel module or user space binary.
-As a result most of the actual components are dynamically allocated and connected at run time.
-While this may be an advantage considering the alternate is to recompile the linux kernel, it does act as significant bottleneck in the case of user space.
-Moreover some of the restrictions on how the configuration can be parallelized limits the actual performance that could be achieved in combination with netmap.
-By integrating the ability to access click's elements within our parallel processing engine we were able to scale significantly better.
-This does require a limited amount of modifications to some of click's libraries and elements.
-In generating the actual code statically, compilers such as gnu c, etc., can perform optimizations such as loop fusions, in-lining, dead code optimization etc., which were not possible previously.
+\begin{algorithm}[h]
+\caption{To send packets}
+\label{s_pack}
+\begin{algorithmic}[1]
 
-Our framework primarily focuses on user space and netmap APIs forms a crucial part in this.
+\State \textbf{struct} $ring[ n ]$\textbf{\{}
+\State \Comment{Stores information about hardware rings}
+\State \Comment{ $ n \gets \ $tx\ rings\ assigned\ to\ thread}
+\State \  \  $avb$, available space
+\State \  \  $curr$, current position
+\State \  \  $limit$, total space \ \textbf{\}}
 
-\section{Background}
+\Function{send\_data\_mqueue}{Ring $f(n)$, Data $packet$}
+\While{$!success$}
+	\State $i \gets f( n ) $ \label{s_pack_param}
+	\State $success = $ \texttt{SEND\_DATA}($ i, packet $)
+\EndWhile
+\EndFunction
+\Function{send\_data}{Ring $i$, Data $packet$}
+\If{ $ring[ i ].avb > 0$ }
+	\State $hw\_ring[ i ].slot( ring[ i ].curr ) \gets packet$
+	\State $ring[ n ].avb \gets ring[ i ].avb - 1$
+	\State $ring[ i ].curr++$
+	\If{ $ring[ i ].avb == 0$ }
+		\State $hw\_ring[ i ].avb \ -=\ ring[ i ].limit$
+		\State $hw\_ring[ i ].curr \gets ring[ i ].curr$
+	\EndIf
+	\State \textbf{return} $success$
+\EndIf
 
-The constant advancement in server hardware has made systems with large number of hardware threads[] readily available.
-In addition to this the presence of co-processors[], network processors[], Graphics Processing Units ( GPUs ) \& General Purpose GPUs ( GPGPUs ) has dramatically improved the capability of commodity hardware.
-These developments has paved way to a platform with immense computing potential which existing applications already take advantage of.
-Along with the improvements in the Multi-Core \& Many-Core processor architecture the memory and I/O speeds have also seen significant improvement.
-In particular the network subsystem has seen significant growth, in combination with the improvements to optical communication it has lead to 10GbE interfaces being quite common in enterprise servers.
-While 40GbE and 100GbE are available for backbone infrastructure but would eventually follow into commodity hardware.
+\State \textbf{\textit{issue ioctl and poll}}
+\State $ring[ i ].avb \gets hw\_ring[ i ].avb$
+\State $ring[ i ].limit \gets hw\_ring[ i ].avb$
+\EndFunction
 
-Network applications represent those applications that operate on any of the seven network layers described in the OSI Model. \{describe more here\}
+\end{algorithmic}
+\end{algorithm}
 
-Several previous work have discussed the process of parallelizing network applications[][][] to enable it to run faster in such multi threaded systems.
-The development of Software routers[][][] is encouraged further because it costs only a fraction of the commodity hardware routers.
-Moreover the advancements in the network interfaces has made faster interfaces readily in combination with the multi core servers the performance  It has been shown that such routers can be used to process packets at line rate [][][] which has been a challenge on its own.
-Network applications including those that operate on few or all of the network layers have been designed with sequential processing as the target environment.
-The shift to multi core architectures in recent times has induced such applications to utilize parallel processing as a alternate means of computing.
-While this could lead to significant improvement in performance it also suffers from the same problems as any parallel application.
-These include scalability, load-balancing, data dependency and the presence of non parallel regions in the program apart from those that might arise due to that specific application.
-The scaling of software routers has been discussed extensively in RouteBricks[] which shows the problems suffered by software routers when they are scaled beyond a single system.
-One could argue that complexity of network applications such as Deep Packet Inspection ( DPI ), Packet Forwarding, etc., could be handled by a single many socket Multi-Core SMP system.
-However, as the complexity of network applications develops it would be inevitable but to move to a multi node environment in order to perform the computations and to handle more network bandwidth.
+\subsubsection{Packet Processing Operations}
 
-Large scale computing clusters has been used in high-performance scientific application to solve problems and applications of very large size.
-It uses a cluster of compute nodes and distributes the workload across them in order to solve it faster.
-These systems utilize parallel programs consisting of both threads and MPI to break massive problems into smaller workload that can be executed in parallel.
-In addition, the presence of GPUs and GPGPUs requires a mixture of parallelization constructs such as OpenCL, CUDA, etc., in order to achieve the best performance in the system.
-In this paper we discuss how some of the techniques used by these applications could be used to parallelize network applications.
+Packet processing operations consist of converting the raw data that the application provides into network packets suitable for transmission.
+Since the application tasks are represented as a set of operations on a data stream, we include these operations as stages through which the data flows through to get transformed into data packets.
+Integrating them with the application task graph makes it easier to extract parallelism later in the execution engine.
+Several user space libraries exist that perform packet related processing operations like libnids \cite{libnids}, lwip \cite{Dunkels01designand}, Click \cite{Kohler2000} etc.
+Since our application is targeted at a closed environment, we chose the Click modular router due to its flexibility and vast library of elements.
+Our current framework supports the UDP protocol but Click's modular infrastructure makes it easy to support more complex ones.
+\comment{Indeed, stream applications with more complex protocols would benefit even more from the framework due to the parallelizing of these complex tasks.}
 
-Click Modular Software Router[] consists of extensive elements which are used to construct software routers.
-Click primarily targets packet processing applications especially those that operate on L4 - L7.
-\{elaborate more here\}
+Each of the elements in Click contain a virtual function which is called from the previous element it is attached to.
+It was necessary to overcome the limitation of Click's infrastructure where elements are constructed based on dynamically initialized objects and virtual functions.
+This was required due to the need to eliminate the overhead associated with the dynamic construction of the objects and virtual functions, and also to allow for optimizing the task graph along with these operations.
+To achieve these goals, our framework provides a simple support structure to allow Click elements to be imported and to be integrated with applications.
+\begin{algorithm}[h]
+\caption{To receive packets}
+\label{r_pack}
+\begin{algorithmic}[1]
 
-Discuss about what each sections contains here.....
+\State \textbf{struct} $ring[ n ]$\textbf{\{}
+\State \Comment{Stores information about hardware rings}
+\State \Comment{ $ n \gets \ $rx\ rings\ assigned\ to\ thread}
+\State \  \  $avb$, available packets
+\State \  \  $curr$, current position
+\State \  \  $limit$, total packets \ \textbf{\}}
 
-\section{Background}
+\Function{receive\_data\_mqueue}{Ring $f(n)$}
+\While{$!packet$}
+	\State $i \gets f( n ) $ \label{r_pack_param}
+	\State $packet =$ \texttt{RECEIVE\_DATA}(Ring $i$)
+\EndWhile
+\EndFunction
 
-Talk about click...\\
+\Function{receive\_data}{Ring $i$}
+\If{ $ring[ i ].avb > 0$ }
+	\State $packet \gets hw\_ring[ i ].slot( ring[ i ].curr ) $
+	\State $ring[ i ].avb \gets ring[ i ].avb - 1$
+	\State $ring[ i ].curr++$
+	\If{ $ring[ i ].avb < threshold$ } \label{r_pack_thres}
+		\State $hw\_ring[ i ].avb -= ring[ i ].( limit - avb)$
+		\State $hw\_ring[ i ].curr \gets ring[ i ].curr$
+		\State $ring[ i ].avb \gets hw\_ring[ i ].avb$
+		\State $ring[ i ].limit \gets hw\_ring[ i ].avb$
+	\EndIf
+	\State \textbf{return} $packet$
+\EndIf
 
-Describe about what would be the best way to parallelize using threads and MPI.. provide reference.. \\
+\State \textbf{\textit{issue ioctl and poll}}
+\State $ring[ i ].avb \gets hw\_ring[ i ].avb$
+\State $ring[ i ].limit \gets hw\_ring[ i ].avb$
+\EndFunction
 
-\section{Design}
+\end{algorithmic}
+\end{algorithm}
+\subsection{Parallel Processing Engine}
 
-This section describes the design principles related to our methodology.
-The objective is to determine the requirements necessary to parallelize network applications.
-Our methods focus on ways to improve the click system in some cases even extend its functionality.
+After integrating the packet operations by means of support from \texttt{Click}'s infrastructure and access to network hardware through \texttt{netmap}'s APIs, we integrate these operations with the application graph.
+The results is a complete task graph that does both the computation and communication operations.
+The next logical step is executing these tasks in parallel by exploiting data and task parallelism.
+Packet streams are inherently operable in a data parallel fashion i.e.
+it is possible to operate on two different streams simultaneously.
+Since the application tasks depend on the data packets, they could also be executed in parallel.
+So from the integrated graph constructed by combining the application and packet processing operations, it becomes possible to extract data parallelism by operating on different data streams simultaneously.
+By executing different components, representing distinct operations on the data streams, it is also possible to exploit task parallelism.
 
-\subsection{Emphasize on user level execution environment}
+\subsubsection{Execution engine}
+\label{ppe}
 
-The existing software routers are predominantly implemented in the kernel space for the ease of access to resources and can bypass the OS overhead incurred when data is accessed through the network subsystem.
-While this alternative provides better performance it also leads to concerns regarding stability and security of the system.
-Newer techniques such as Netmap[] allows direct access to network buffers which makes it possible to write high performance packet processing applications in userspace itself.
-Typically network applications which operate on L4 - L7 are developed in the kernel space in order to intercept the network packets before it flows through the OS.
-Whereas those applications that utilize L1 - L3 are present in the user space.
-The objective here is to completely switch to user space to solve any of the stability or security concerns.
-In addition, scenarios where the applications that operate across most or all the layers could be improved by combining the operations across the different layers in user space itself.
+The execution engine consists of a set of processes and threads that the work is scheduled upon in order to be executed.
+When implementing threads to execute code which consist of system calls such as \textit{malloc(), ioctl(), poll(),} etc., we found a performance degradation, even when the threads are completely independent.
+To support the lightweight style of threads, the OS implements most of the locking mechanisms for system calls from threads.
+Previous research such as \cite{Lever:2000:MPM:1267724.1267780} \cite{Dice:2002:MLM:773039.512451} \cite{Michael:2004:SLD:996893.996848} shows the degradation in performance due to the usage of \textit{malloc()} in threads.
+Since our framework has to deal with system calls and issue them in parallel, we use \textit{unix process} to perform the tasks concurrently.
+We do this by using Message Passing Interface (MPI) \cite{TheMPIForum:1993:MMP:169627.169855} to construct a set of processes that is scheduled onto the different CPUs on a multicore system.
 
-\subsection{Eliminate the overhead associated with creation of dynamic objects}
+In order to prevent unwanted process migration due to the OS scheduler, we specify \textit{affinity} of the MPI processes to particular CPU cores.
+This prevents each process from being executed on a different CPU other than that specified.
+This makes the processes act as execution engines, running on each of the cores on which we schedule the operations that are to be performed.
+Previous research \cite{Salehi:1996:EAS:234766.234769}\cite{1409136} highlights the importance of affinity in network sensitive applications.
+Next, by utilizing our \textit{Parallel Packet I/O}, we specify the hardware queues each of these processes handle.
+\comment{Initially we divide the queues evenly among the different processes we create, but later during execution we could modify the allocation based on the application operations scheduled on them.
+Our modifications to \texttt{netmap}'s hardware queue assignment API can help us here in redistributing the queues to different processes.} By using MPI's \texttt{send} and \texttt{receive} inter-process communication calls to handle communication between execution engines located at each CPU, it is possible to modify the attributes of the process such as CPUs assigned or number of NIC queues handled, and even the operations the processes were initially assigned.
 
-Click constructs application graphs based on dynamic creation of objects and passing the data packets across the virtual functions.
-Though this is a very elegant approach in creating the software router it does incur the penalty of the virtual function tables.
-The click-tool[] addresses this issue but it does not solve the problem entirely.
-The construction of the software router at run time is the primary reason behind it.
-This mainly arises due to the presence of kernel module which installs the router as a kernel module based on the input configuration.
-However in the user level it would not be necessary for a separate process to initialize the router instead we could generate the entire program statically including the necessary flow associated with the specific application.
-This not only eliminates the overhead associated but also opens up means to optimize the graph further.
+\subsubsection{Parallel task processing}
+\label{ppt}
 
-\subsection{Optimize based on application graph}
+Once we have scheduled our execution engines on the physical system, the next step is to execute the integrated task graph consisting of application and packet operations.
+To extract data parallelism, we replicate the task graph onto multiple execution engines.
+This ensures operations on the same data are performed on the same CPU, thereby using local cache.
+By making sure that packets belonging to a specific flow always end up in the same execution engine, we only perform operations conforming to that stream of data.
+\comment{We use the ability of modern NICs to classify packets onto different queues, based on their attributes.}
 
-Network applications especially that operate on the lower layers are very sensitive to the overhead associated with operating on the data separately on each layer.
-Instead given an application graph if we were to provide a sequential flow of the different functions at the various layers then it would be possible for the compiler to optimize the regions of program into a single and optimized block of code.
-This would also improve the cache locality of the data accessed by consecutive functions.
-This also presents us with the ability to perform other types of stream graph optimizations[] which would not be possible if the specific graph was constructed dynamically at run time.
+\comment{
+\lstset{language=C,caption={Packet Payload Modification Function},label=payload_sub}
+\begin{lstlisting}
+bool payload_copy( const void *payload_data, uint32_t length )
+{
+    uint32_t headroom = _data - _head;
+    uint32_t tailroom = _end - _tail;
+    uint32_t orig_length = _tail - _data;
+    uint32_t n = length + headroom + tailroom;
+    if( n < min_buffer_length )
+    {
+        tailroom = min_buffer_length - length - headroom;
+        n = min_buffer_length;
+    }
+    _data = _head + headroom;
+    _tail = _data + length;
+    if( payload_data )
+        memcpy( _data, payload_data, length );
+    return true;
+}
+\end{lstlisting}
+}
 
-\subsection{Automated data parallelism}
+In order to replicate the task graph we had to overcome some restrictions.
+First was to ensure that there was no intricate data dependency between different data streams.
+Although Click's library of elements are designed to be executed in parallel \cite{Chen:2001:FCP:647055.759948}, they are more oriented towards serial execution.
+Hence we had to ensure the code present in the elements were thread-safe.
+One important modification we had to perform was in Click's packet library.
+By default, Click destroys and recreates the packet's internal storage buffer by reallocating memory space, an expensive operation.
+In order to avoid this, we allocate a packet with a large static buffer and copy the data to be modified on it.
+We then modify the pointers that determine the header and data addresses, based on the required packet length.
+\comment{This is done by using a simple sub-routine shown in Listing \ref{payload_sub} that we added to the packet library.} Although this addition incurs the overhead of a copy operation, it avoids performing invalid and dangerous memory operations directly on the network hardware queue's buffer.
 
-Network traffic is a classic example by which data parallelism can be utilized to achieve scalable parallelism.
-Click's support for parallelism in user level is described as experimental, it is implemented by using a task graph consisting of each work associated with each element which is executed in the threads.
-The support for parallelism in the kernel mode is better in comparison to that of user mode.
-Since our target environment is in the user space the proposed design involves a different approach to which we can extract data parallelism.
-We first split the data across multiple queues and then replicate the application graph across to process the data the queues in parallel.
-An important condition here is to make sure packets of the same flow should pass through the same queue.
-Since the application graph can be replicated it also makes it possible to dynamically load balance several applications based on the workload.
-
-\subsection{Efficient implementation of task parallelism}
-
-In generating the application graph as a sequence of operations it also makes it possible to process different stages of the operation in a pipelined manner.
-This when combined with stream optimizations and could be used to determine which steps of the graph is best suited to form the stages of a pipeline.
-The condition to balance here is the weight associated with computation in that stage with that of the overhead incurred in computing the data in two separate stages.
-It would be necessary to determine whether the additional over head involved in executing the task in parallel provides any significant improvement in performance.
-
-\section{Implementation}
-
-Click was designed as a means to easily implement a software router on top of a Linux OS.
-This is implemented in two different ways one is by installing a kernel module and the other is a program that executes in the user level.
-Both of them accept a configuration file described in click's language specification.
-This is then used to construct the router by linking the virtual functions for the push and pull operations of the different elements in accordance with the specified graph.
-
-\subsection{Modifications to click system}
-
-\subsubsection{Static Generation Of ???}
-The first step in the design was to overcome the limitation of using click's infrastructure to construct the software router which consists of dynamically initialized objects and virtual functions.
-This was necessary due to two important conditions, one was to eliminate the overhead associated with the dynamic construction of the objects and virtual functions and the other was to provide the compiler a chance to optimize the program during the compilation phase.
-This would however result in not being able to construct the software router using click's infrastructure but would require the application specific code to be recompiled.
-This could be a significant hurdle in the case of running the software router in kernel mode where the router configuration is installed by the click kernel module in which case the entire kernel module should be recompiled and reinitialized.
-Whereas in user space this would not be an issue since routers of different configurations are basically treated as separate applications and would just have to go through an additional compilation phase before being used.
-The added advantage of this would be to enable network applications that operate on L1 - L3 layers to use the packet processing ( L4 - L7 ) infrastructure provided by click elements.
-
-\subsubsection{Implementation of data parallelism}
-
---Talk about MPI \& threads\\
-
---Our goal here is to focus on the network application as a workload and characterize it as such without involving any kind of overhead that might occur due to the network.\\
+Click's packet library also provides a mechanism to take ownership of a specific address and pass a destructor to be called when the packet is killed.
+To prevent corruption of memory mapped from the hardware queue by the application tasks, we use the above mechanism to copy the data before modifying it.
+In generating the application graph as a sequence of operations, it also makes it possible to run different graphs on each individual execution engine.
+MPI communication is used to schedule the application and packet operations on the processes representing the execution engines.
+Furthermore, we use it to also perform explicit data sharing between tasks in the application that require it.
 
@@
 
+In order to evaluate our framework, we test the scalability and the performance speedup on a multicore system for an application that is constructed using the framework.
+Using two generic operations: compression and encryption, we build simple stream applications using our framework.
+We then measure the performance improvement as we increase the number of execution engines dedicated to the application.
+
+\subsection{Hardware Setup}
+
+The evaluation system consists of two servers connected directly to each other.
+Each server consists of two Intel Xeon 5600 hexacore HT in a dual socket configuration with 24GB DDR3 memory.
+In total there is 2 Processors x 6 Cores x 2 HT = 24 CPUs in each server.
+These servers are connected using Intel 8299EB 10Gbe adapters over directly attached twin-axial cables.
+The adapters are located in the PCI Express x8 slot to maximize the available bandwidth.
+This forms a closed connection between two servers and limits the influence of external devices such as switches or routers on data transmission.
+
+Each application is constructed as two parts, one a source of data and the other a sink.
+We measure the overall time it takes for application task operations, packet operations, and the time to send from the source and receive at the sink.
+This is done for different packet sizes and applications.
+Then we plot the speedup achieved in each of the cases over the serial version where we have a single sender and receiver operating on the multiple queues.
+This is the baseline in our experiments and is similar to how the OS behaves for transmitting and receiving using the multi queue network controller.
+
+The Intel 82599EB is restricted to 16 queues when using Receive Side Scaling (RSS).
+This limited the maximum amount of processes at the sink to be 16, even though 24 CPUs were available.
+The source is not affected by this limitation and can use all the 24 CPUs available, along with 24 hardware transmission queues.
+Overall, we require at least one hardware queue mapped to each of our processes or execution engines.
+\comment{We also ensure that each of the processes are bound to a particular CPU by issuing affinity calls to the OS to ensure that they are not migrated between different cores.}
+
+\subsection{Application Construction}
+\label{app-const}
+
+In order to evaluate our framework, distinct operations were required that would be suitable for a streaming framework.
+Since compression and encryption applications are relatively common, we chose two applications composed from these operations.
+The integrated graph that represents both these applications is shown in fig \ref{fig-app-graph}.
+UDP is used to send the data between source and sink.
+
+\begin{figure}[htb]
+\centering
+\includegraphics[width=3.5in]{enc-dec}
+\caption{ Integrated Task Graph  }
+\label{fig-app-graph}
+\end{figure}
+
+Both applications are constructed similarly.
+Data sizes of 1.28GB and 2.56GB are selected as the workload for the compression and encryption applications respectively.
+The OpenSSL library is used for encryption/decryption and zlib is used for compression/decompression operations.
+Each application consists of a data generation phase, then data is compressed or encrypted depending on the application.
+Packet operations are performed after the data is generated.
+The stream consisting of packets is then sent over the network using multiple queues.
+The sink side performs the reverse of these operations.
+Once data is received, packet operations are performed to extract the payload, followed by decompression or decryption to retrieve the original data.
+These are representative of some of the basic operations that would be involved in a much more complex application.
+\comment{Choosing to compare these simple operations avoids any extraneous factors that may be present in a more complex application.}Use of these simpler stream applications enables us to evaluate the performance of our framework and not that of any optimizations specific to the application.
+
+\begin{figure*}[tb]
+\centering
+\includegraphics[width=7in]{result}
+\caption{Performance Speedup; (a),(b),(c) - Application A.
+Encryption; (d),(e),(f) - Application B.
+Compression;\\ \newline Speedup based on time in takes to compute a fixed workload for different number of source and sink processes\\ \newline and for varying packet sizes over a single source and sink are shown above}
+\label{fig-res}
+\end{figure*}
+
+\section{Results}
+\label{results}
+It can be seen in Figure \ref{fig-res} that there is considerable difference in the way each of the operations scale. Three distinct observations can be made from the results.
+
+One is that, from fig\ref{fig-res}.(d,e,f), we note that compression requires more work than decompression, whereas from fig\ref{fig-res}.(a,b,c) encryption and decryption are more balanced.
+This is evident from the continued speedup we get up to 4 senders in the case of compression/decompression.
+Only a slight asymmetry can be seen in the encryption/decryption case, where adding more senders than receivers only yields a slight increase in performance.
+In this case, the likely explanation is support for AES instructions in the architecture, making decryption a marginally easier task.
+In general though, the performance here only improves by increasing both the number of senders and receivers simultaneously.
+Even from these preliminary results, it is clear that the optimal sender/receiver balance varies from application to application and highlights the need for effective load-balancing.
+
+Packet size also impacts performance as seen in fig\ref{fig-res}.(d,e,f).
+The compression operation is influenced more than that of the encryption application.
+There is a speedup of 12x in fig\ref{fig-res}.(d) with smaller packets and 10x in fig\ref{fig-res}.(f) with larger packets.
+This discrepancy appears to be greater in cases where there is a large number of senders and a high number of receivers.
+A possible explanation for this relates to the CPU utilization on the faster receivers.
+With large packet sizes, they are more likely to block waiting for data to be streamed from the slower senders.
+Smaller packets allow the work to be distributed more evenly among the available receivers.
+In the encryption application, there is not as much asymmetry and packet size does not matter as much.
+The fact that packet size impacts applications differently suggests that is essential to incorporate dynamic packet-size selection into the framework itself.\comment{ to respond to the unique requirements of each application.}
+
+From fig\ref{fig-res}(a,b,c,d,e,f), a drop in speedup is apparent when employing 12 senders or more.
+The most likely explanation for this is the influence of the NUMA architecture \cite{Awasthi:2010:HPO:1854273.1854314}.
+The 12 sender point is when all the cores of a particular CPU are utilized and moved onto a different socket.
+This causes issues in relation to cache locality and the limitation of the network controller to receive data through the PCI-E bus from more than one CPU.
+Despite these NUMA architecture issues, once the initial drop is overcome, the application continues to scale and eventually attain better performance in most cases.
+
+In summary, it can be seen that the framework scales well for different workloads, can handle both computations and communication, and can operate on them in parallel.
+In some cases, the framework can be improved by having a greater degree of control over the fine grained parallelism within the execution engines.
+This can be achieved by introducing pipeline stages for different operations.
+
 \section{Related Work}
 
+Several prior researchers \cite{Dobrescu09routebricks:exploiting}\cite{Han:2010:PGS:1851275.1851207}\cite{Wang:2009:PPN:1542275.1542307}\cite{springerlink:10.1007/s11227-011-0579-3} have worked on using commodity multicore processors to perform network operations.
+
+Dobrescu et al.\cite{Dobrescu09routebricks:exploiting} shows the performance improvement achieved in the construction of software routers by distributing related operations over multiple servers.
+
+Han et al.
+\cite{Han:2010:PGS:1851275.1851207} perform tasks related to high performance packet processing on commodity hardware using GPUs.
+They have developed a novel framework known as PacketShader that can do packet processing in user space by exploiting the massively-parallel architecture of GPUs.
+
+Wang et al.
+\cite{Wang:2009:PPN:1542275.1542307} constructs a high performance connection-affinity based lock-free multicore network processing system that claims to achieve multiple Gbps network processing speed for complex tasks.
+
+Egi et al.\cite{springerlink:10.1007/s11227-011-0579-3} shows techniques on parallelizing network operations on multicore architectures by defining principles that consider multiple resources and not just the CPU alone.
+
+Stream programming languages such as StreamIt \cite{Thies:2002:SLS:647478.727935}, LUSTRE \cite{Halbwachs91thesynchronous}, etc., target stream applications but the focus is more on parallelizing the computations related to the application alone on the multicore system.
+
+\comment{
+All of these applications are more focused on operating as a network application and primarily focus on parallelizing packet operations.
+In our work even though a part of it involves parallelizing network operations, the objective of this research is to provide support for parallelizing stream application and distributing their workloads over multiple systems.
+This also requires optimizing the underlying network operations by parallelizing it.}
+
+\texttt{Netmap} \cite{Rizzo:2012:RNI:2090147.2103536}, provides the APIs to access the hardware queue's buffer in user space by means of memory mapped rings.
+It implements an intermediate ring in kernel space that is used to synchronize the data from the user space to that of hardware queues in the NIC.
+
+The Click modular router \cite{Kohler2000} is an extensively used tool for the construction of software routers using general purpose hardware and provides a comprehensive infrastructure to support this.
+SMP Click \cite{Chen:2001:FCP:647055.759948} is an improved version, which supports parallel processing of these software routers using threads.
+Click implements its own mechanism and polling driver in kernel space to access packets, bypassing the OS's network infrastructure.
+
+\comment{
+In order to send data the programmer calls \textit{poll} on the tx ring in user space and writes to it when space is available.
+Then he calls \textit{ioctl} to synchronize the data with the kernel ring and transmit the data.
+To receive the packets \textit{poll} is again called on the rx ring in user space but this time to check for new data and after reading it, he issues \textit{ioctl} to let the driver know that packet has been read and space can be re used.
+
+, our framework uses techniques as described in Section \ref{imple} to integrate them with the application and access them in parallel from different application processes.
+Also, while some of click's \cite{Kohler2000} elements are used to provide support for packet processing operations, in our framework we use it for doing some of application's network tasks alone.
+Due to this we modified a subset of click's elements in order to support our system.
+
+ In our framework we target a different category of stream applications such as data mining, real time image analysis, etc., which tend to compute on large amounts of data.
+In order to improve overall performance in such cases, the operations are distributed over multiple systems.
+Our framework provides an easier methodology to construct such an application while managing both the parallelizing and communication operations.
+
+Our framework provides the necessary functionality to target communication intensive stream applications.
+We focus on improving the overall performance of such an application by integrating the network and application operations.
+We do this by providing a high-level programming framework that can be used to construct these applications, which can extract parallelism and improve communication performance.
+Using this it would be possible to develop stream applications that could be distributed over multiple systems and using our framework it would be possible to handle the highly intensive communication between them.}
+
 \section{Conclusions}
+In this paper we have proposed a high-level programming framework that can integrate packet processing within stream applications and perform both types of operation in parallel.
+
+We conclude that by using our algorithms for sending and receiving data, which implicitly perform batching and reduce the usage of system calls, and by combining this with \texttt{netmap} APIs which minimizes the interaction with the OS, communication performance in our framework never becomes the bottleneck in our system on the programs we have tested.
+Instead of performing the packet operations in the OS, by integrating it with the application using Click's elements within our infrastructure, we are able to construct a unified task graph that represents the computation and communication operations.
+Using this, we are able to parallelize these operations together and achieve speedups of more than a factor of eight in all the applications we tested.
+
+In integrating these features into a programming framework, the system of multiple queues, parallel communication and computation are entirely hidden from the programmer, who merely specifies a standard stream graph for computation alone.
+This frees up the programmer from the onerous task of having to decide how their application should be distributed, leaving it instead to the execution engine of our framework, which also looks after the ensuing communication requirements after application distribution.
+
+The results show that our system scales to as many as 24 parallel processes on a multicore computer, and achieves speedups of more than a factor of ten in some cases compared to sequential implementations.
+Results also suggest where efforts should be made in developing the framework further.
+Until now, the focus has been on deploying the execution engines in order to execute stream applications and to enable optimal communication between separate nodes.
+
+Future work will address some of the opportunities presented by these execution engines and the collaboration between them.
+Dynamic load balancing is likely to be the most benificial addition to the framework in terms of performance of the stream applications built using it.
+It is also possible to add the capability to dynamically adjust the number of network queues allocated to each core as stream application requirements change over time.
+Finally, as seen in section \ref {results}, being able to adjust packet size to an optimal value for each network link, again dependent on stream application requirements, would also be a useful addition to the framework.
 
@@
 \section*{Acknowledgment}
-This work is funded by the IRCSET Enterprise Partnership Scheme along with IBM Research, Ireland.
+This work is funded by the IRCSET Enterprise Partnership Scheme in collaboration with IBM Research, Ireland.
 

### 2013-03-21 paper.tex
Paper updates

@@
 %
-\documentclass[conference]{IEEEtran}
-%% INFOCOM 2013 addition:
-\makeatletter
-\def\ps@headings{%
-\def\@oddhead{\mbox{}\scriptsize\rightmark \hfil \thepage}%
-\def\@evenhead{\scriptsize\thepage \hfil \leftmark\mbox{}}%
-\def\@oddfoot{}%
-\def\@evenfoot{}}
-\makeatother
-\pagestyle{headings}
+\documentclass[10pt, conference, compsocconf, reqno]{IEEEtran}
+
 % Add the compsoc option for Computer Society conferences.
@@
 
+%squeeze
+
+\renewcommand{\arraystretch}{1}
+
+\renewcommand\floatpagefraction{.9}
+\renewcommand\topfraction{.9}
+\renewcommand\bottomfraction{.9}
+\renewcommand\textfraction{.1}
+\setcounter{totalnumber}{50}
+\setcounter{topnumber}{50}
+\setcounter{bottomnumber}{50}
+
+\raggedbottom
+
+\addtolength{\floatsep}{-5mm}
+\addtolength{\textfloatsep}{-3mm}
+%~ \addtolength{\intextsep}{-2mm}
+\addtolength{\dbltextfloatsep}{-2mm}
+\addtolength{\dblfloatsep}{-1mm}
+\addtolength{\abovecaptionskip}{-1mm}
+\addtolength{\belowcaptionskip}{-1mm}
+\addtolength{\abovedisplayskip}{-3mm}
+\addtolength{\belowdisplayskip}{-3mm}
+\addtolength{\arraycolsep}{-1mm}
+
+\let\oldthebibliography=\thebibliography
+\let\endoldthebibliography=\endthebibliography
+\renewenvironment{thebibliography}[1]{%
+  \begin{oldthebibliography}{#1}%
+    \setlength{\parskip}{0.2ex}%
+    \setlength{\itemsep}{0.2ex}%
+}%
+{%
+  \end{oldthebibliography}%
+}
+
+\captionsetup{justification=centering}
+
 % correct bad hyphenation here
@@
 % can use linebreaks \\ within to get better formatting as desired
-\title{A high-level programming framework to support parallel processing of communication intensive \\ stream applications}
+\title{A parallel runtime framework for communication intensive stream applications}
 
@@
 % affiliations
-\author{\IEEEauthorblockN{Servesh Muralidharan}
-\IEEEauthorblockA{School of Computer Science and\\Statistics\\
-Trinity College Dublin\\
-Dublin 2, Ireland\\
-Email: muralis@tcd.ie}
-\and
-\IEEEauthorblockN{David Gregg}
-\IEEEauthorblockA{School of Computer Science and\\Statistics\\
-Trinity College Dublin\\
-Dublin 2, Ireland\\
-Email: David.Gregg@cs.tcd.ie}}
+\author{\IEEEauthorblockN{Servesh Muralidharan\IEEEauthorrefmark{2},
+Kevin Casey\IEEEauthorrefmark{4}
+and
+David Gregg\IEEEauthorrefmark{2}}
+\IEEEauthorblockA{\IEEEauthorrefmark{2}School of Computer Science and Statistics,
+Trinity College Dublin, Dublin 2, Ireland\\
+Email: muralis@scss.tcd.ie, David.Gregg.cs.tcd.ie }
+\IEEEauthorblockA{\IEEEauthorrefmark{4}DCU\\
+Email: }
+}
+%~
+%~ \author{\IEEEauthorblockN{Servesh Muralidharan}
+%~ \IEEEauthorblockA{School of Computer Science and\\Statistics\\
+%~ Trinity College Dublin\\
+%~ Dublin 2, Ireland\\
+%~ Email: muralis@tcd.ie}
+%~ \and
+%~ \IEEEauthorblockN{David Gregg}
+%~ \IEEEauthorblockA{School of Computer Science and\\Statistics\\
+%~ Trinity College Dublin\\
+%~ Dublin 2, Ireland\\
+%~ Email: David.Gregg@cs.tcd.ie}}
 
@@
 Stream applications are often limited in their performance by their underlying communication system.
-In a typical implementation, the Operating System (OS) handles the majority of network operations.
+In a typical implementation, the underlying communication relies on operating system to handle majority of the network operations.
 In such cases, the communication stack, which was not designed to handle such tremendous amounts of data, acts as a bottleneck and restricts the performance of the application.
-Therefore, a framework for computing and communicating in parallel is essential for improving performance.
-
-We propose a high-level programming framework that integrates the packet processing operations with stream applications, and provides a common interface to execute both application and packet processing operations in parallel on multicore processors.
-
-We utilize the ability of the network interface controller (NIC) to classify packets onto multiple hardware queues.
-This is combined with system software that allows direct access to these queues.
-We develop an interface in our framework for application processes to read or write in parallel.
-Our system integrates this with the application by a series of packet processing operations that transform the data from the application to packets suitable for transmission.
-Representing these tasks as stages in the application, we are able to parallelize them using techniques similar to those used for parallelizing regular computations.
-
-All the communication operations in the framework are performed in userspace, bypassing the OS, to send and receive data directly over the network interface.
-By overcoming the OS bottleneck, and by parallelizing computations and packet operations together, we are able to boost the performance of stream applications substantially.
+
+In this paper, we propose a parallel runtime framework that integrates the communication operations with stream applications, and provides a common parallel processing engine that can execute both the communication and computation operations in parallel on multicore processors.
+
+Our runtime system does this by using a series of communication filters that transform the data from the application to packets suitable for transmission.
+These communication filters are built using system software that allows direct access to network hardware.
+We also utilize the ability of the modern network controller to classify packets onto multiple hardware queues by integrating it with the communication filter.
+Using this our runtime framework is able to construct a unified stream graph based on the application graph provided.
+This when used with a static schedule based on a given architecture, can process the data streams in parallel.
+
+We are able to parallelize the stream application and achieve speedups of more than a factor of eight in all the applications we tested.
+The results show that our system scales to as many as 24 parallel processes on a multicore computer, and achieves speedups of more than a factor of ten in some cases compared to sequential implementations.
 
@@
 \IEEEpeerreviewmaketitle
-%Kevin - vertical spacing
-\setlength{\textfloatsep}{15pt}
+%Kevin - vertical spacing % Servesh - Fixed it on global command
+%\setlength{\textfloatsep}{15pt}
 \section{Motivation}
 
-Stream applications are designed in a manner where they operate on data that flows as a \textit{stream}.
-This data stream passes through different stages of computations, to be transformed from query to result \cite{Thies:2002:SLS:647478.727935} \cite{Halbwachs91thesynchronous} \cite{Stephens95asurvey}.
-This differs from a conventional computation model where data is stored and then processed.
-By eliminating the need to store data, and processing it on the go instead, it is possible for stream applications to process large amounts of data.
-
-Stream applications such as file storage systems, database query management, data mining, etc., operating on vast amounts of data, are often distributed over multiple systems.
+Applications that follow streaming model~\cite{Thies:2002:SLS:647478.727935,Halbwachs91thesynchronous,Stephens95asurvey,muthukrishnan2005data} are expressed as a set of filters that operate on one or more tokens of data.
+The flow of these data tokens through the filters is referred to as a \textit{stream}.
+Since the data is accessed sequentially in the form of a stream the memory requirement to process the data is considerably less compared to the size of the enitre stream, thereby making it possible to operate on very large data sizes.
+Typical applications based on this model operate on data ranges in the order of terabytes or even bigger.
+Applications such as financial trading systems~\cite{agarwal2009faster}, database systems~\cite{Golab03datastream}, network data analysis~\cite{Gilbert01quicksand:quick,Sullivan:1998:TSM:1268256.1268258} are few examples that follow the streaming model and operate on very large data sets.
+
 Traditional stream applications target multicore processors for performing their computations in parallel.
-As multicore processors grow in power, so do stream applications, processing ever-increasing amounts of data.
-However, stream applications that operate on large amounts of data at a very high bandwidth can be limited by the communication stack of the OS kernel\cite{Wu:2007:PPB:1323954.1323957}\cite{Wu:2007:PAL:1227865.1228021}\cite{Dobrescu09routebricks:exploiting}\cite{Han:2010:PGS:1851275.1851207}\cite{Kohler2000}.
-With the continuing improvement of multicore and many-core architectures, it is becoming essential to improve the processing speeds of packet operations along with those of the application in order to scale performance.
+As multicore processors grow in size, so do stream applications, processing more streams of data in parallel.
+However, in such cases the applications are typically limited by the traditional communication systems~\cite{Wu:2007:PPB:1323954.1323957,Wu:2007:PAL:1227865.1228021,Dobrescu09routebricks:exploiting,Han:2010:PGS:1851275.1851207,Kohler2000}.
+The reason being these systems have limited support for operating on parallel streams and are burdened by abstraction layers that provide compatibility for non-streaming applications.
+
+Several high speed networks~\cite{fredj2011survey} exists that could handle such communication requirements but these require expensive hardware and specialized programming skills.
+Commodity network adapters based on 10-gigabit ethernet (10GbE) are comparatively cheaper alternative and are being widely deployed in clusters and grids in data centers.
+Previous research\cite{Romanow03anoverview,Feng:2003:OEN:1048935.1050200} have been done on optimizing ethernet for improving communication in such cases.
+However, they do not use many of the features provided by modern day network adapters such as hardware queues, flow control, etc., Also, they do not focus specifically on streaming applications and hence making it difficult to utilize them properly.
+
+\section{Contribution}
+
+In this paper we propose a parallel processing framework for building stream applications that are bound by very large communication.
+A simplistic framework where the filters are described as \textit{actors} and the application is represented by connecting such \textit{actors} into a directed acyclic graph (DAG).
+Then our framework is used to establish, (1)the communication system that allows these \textit{actors} to send and receive data across each other and (2)the computation system that executes these tasks in parallel on a given multicore system.
+The main contributions can be described as,
+
+\begin{enumerate}
+\item Parallel multi point communication layer, which provides the ability for stream actors to communication across the network.
+\item Execution system that can run the given data parallel filters on the multicore system
+\item Utilizes network hardware features such as queues and flow control to support stream applications
+
+\end{enumerate}
+
 \comment{
-When performing packet processing operations for communication intensive stream applications, a major challenge is the high levels of performance required.
-These applications process large numbers of data streams in real time.
-If the communication system cannot keep up with the rate at which packets arrive, then they must be dropped or remain unprocessed.
-}
-
-In building a framework for such applications on general purpose computers, we can exploit two hardware features that could allow for dramatically improved performance, namely multicore processors and multi queue network controllers.
-Using these features makes it possible to do much of the packet processing in parallel using the multiple cores commonly available on modern processors.
-
-A major bottleneck for packet processing on general purpose computers is the overhead of the operating system's network stack.
-In particular, for applications that run as user processes, data must be copied between OS kernel space and user space.
-Previous research uses alternate mechanisms to access packets from the network hardware directly to overcome this problem\cite{Dobrescu09routebricks:exploiting}\cite{Han:2010:PGS:1851275.1851207}\cite{Kohler2000}.
-Most of these use modern network hardware in combination with particular system software that allows direct access to hardware-controlled packet queues, dramatically reducing overheads and improving performance.
-However, constructing stream applications that use these features is difficult.
-Furthermore, many stream applications have a very specific communication pattern, typically operating on data between a source and sink in the form of a \textit{data stream}.
-Data in a stream must be processed sequentially, whereas those from separate streams can often be processed in parallel.
-
-Modern network hardware typically has a mechanism for separating arriving packets into multiple queues, where packets that belong to the same flow are\comment{ guaranteed to be} placed in the same queue\cite{micro2008} \cite{intel2010}.
-We can utilize this feature to operate on multiple streams of data by processing each of the queues in parallel.
-However, building such applications is complex because it requires explicit parallel programming and interaction with low-level network interfaces.
-Research focuses on improving the performance of network applications by parallelizing the packet processing tasks on multicore architectures \cite{Dobrescu09routebricks:exploiting} \cite{Han:2010:PGS:1851275.1851207}.
-We parallelize both the application and packet operations to gain significant improvement in performance.
-
-\section{Contribution}
-
-In this paper we propose a high-level programming framework for building communication intensive stream applications.
 We focus on integrating the packet processing and application tasks into a single graph and using this to extract data parallelism.
-The developer describes their program as a graph of \textit{actors}.
+The developer describes their program as a graph of \textit{filters}.
 Each actor consumes one or more \textit{tokens} of data, which may be simple data such as integers, or more complex data such as packets, arrays, or objects and produces one or more tokens.
@@
 \end{itemize}
+}
 
@@
 
-\section{Background}
-
-In this paper, we are concerned with applications that operate on data streams.
-Example operations of such an application are represented in the stream graph shown in figure \ref{fig1}.
+\comment{
+With the continuing improvement of multicore and many-core architectures, it is becoming essential to improve the processing speeds of packet operations along with those of the application in order to scale performance.
+}
+\comment{
+When performing packet processing operations for communication intensive stream applications, a major challenge is the high levels of performance required.
+These applications process large numbers of data streams in real time.
+If the communication system cannot keep up with the rate at which packets arrive, then they must be dropped or remain unprocessed.
+}
+
+\section{Introduction}
+
+Our main objective is to optimize performance of streaming applications by First, providing a speaclized communication link between the filters and Second, extracting data parallel actors and executing them in parallel.
+Stream applications usually consists of stateless and stateful filters.
+Stateful filters are those that are depedendent on the previous execution and hence making it difficult to replicate them.
+Stateless filters do not pose such limitations and can be freely replicated based on the available data streams.
+In this paper, we focus on extracting data parallelism based on stateless filters and streams.
+Example operations of such an application are represented in the stream graph shown in Figure \ref{fig1}.
+All the filters used in this application are essentially stateless filters.
 This sample application operates by performing encryption on the data that matches predefined rules like file names, access level of user, etc., otherwise it performs compression.
-Due to the nature of such stream applications, it is possible to distribute the component operations across different CPUs.
-We are particularly interested in a scenario where groups of networked servers are used to perform these kind of stream operations by distributing them evenly over different nodes.
+Due to the nature of such stream applications, it is possible to replicate and distribute the filters across different CPUs.
+We are particularly interested in a scenario where groups of networked servers are used to perform these kind of stream operations by distributing data streams over multiple nodes.
 
@@
 The main challenge that arises in such a situation is how best to address the intensive computational and communication demands of such applications.
-Two important features enable us to tackle this challenge.
-Firstly, the operations of both the application and the network are parallel in nature.
-Secondly, multicore processors can be used to execute the application.
-
-In our system, we use an efficient parallel processing framework that could be leveraged on multicore processors and multi queue network controllers to achieve efficient communication.
+Two important features enable us to tackle this challenge in our framework.
+Firstly, the operations of both the application and the network are designed to be stateless filters.
+Secondly, communication techniques are employed that would enable multicore processors to perform operations related to computation and communication in parallel.
+
+In building a framework for such applications on general purpose computers, we can exploit two hardware features that could allow for dramatically improved performance, namely multicore processors and multi queue network controllers.
+Using the multiple queues it makes it possible to do much of the communication in parallel and using the multiple cores the filters from the stream application can be executed concurrently.
+
+A major bottleneck for communication on general purpose computers is the overhead of the operating system's network stack.
+In particular, for applications that run as user processes, data must be copied between OS kernel space and user space.
+Previous research uses alternate mechanisms to access data from the network hardware directly to overcome this problem\cite{Dobrescu09routebricks:exploiting}\cite{Han:2010:PGS:1851275.1851207}\cite{Kohler2000}.
+Most of these use modern network hardware in combination with specialized system software that allows direct access to hardware-controlled packet queues, dramatically reducing overheads and improving performance.
+However, constructing stream applications that use these features is difficult.
+Furthermore, many stream applications have a very specific communication pattern, typically operating on a \textit{flow}, a specific \textit{stream} between two filters.
+Data in a \textit{stream} must be processed sequentially, whereas different \textit{flows} can often be processed in parallel.
+
+Newer network hardware has introduced a mechanism for separating arriving data into multiple queues, where data that belong to the same \textit{flow} are\comment{ guaranteed to be} placed in the same queue\cite{micro2008} \cite{intel2010}.
+We can utilize this feature to operate on multiple streams of data by processing each of the queues in parallel.
+However, building applications that can use this feature directly is a challenging task because it requires explicit parallel programming and interaction with low-level network interfaces.
+Previous research~\cite{Dobrescu09routebricks:exploiting,Han:2010:PGS:1851275.1851207} that focuses on improving the performance of network applications in such a manner shows the complexity involved in building such a system.
+
+\comment{We parallelize both the application and packet operations to gain significant improvement in performance.}
+
+In our framework, we use a more elegant approach.
+We define a runtime framework that already consists of the communication filters that are pre-built.
+So inorder to implement a stream application such as that shown in Figure~\ref{fig1}, all we need to do is define the computation filters and connect them in the framework to the pre-built communication filters.
+On execution, our runtime framework is able to replicate the stateless filters and the communication filters on to the different multiple cores, thereby executing the operations belonging to different \textit{flows} in parallel.
+
+\comment{
+ parallel processing framework that could be leveraged on multicore processors and multi queue network controllers to achieve efficient communication.
 The performance of multicore processors continues to dramatically improve the capability of commodity hardware, thus making it possible to meet the computational demand of such applications.
@@
 Although other other user space packet-access APIs could be used \cite{Rizzo:2012:RNI:2090147.2103536} \cite{1564468} \cite{Krasnyansky}, the simpler construction of \texttt{netmap} was more suitable to adapt to our framework.
+}
 
@@
 In section \ref{ppe}, we discuss the challenges in using a system consisting of execution engines to manage the work.
-The extent to which we extract parallelism ranges, not only to the packet tasks, but also to the low level access of data I/O to the network interface hardware through custom additions to the \texttt{netmap}.
-Finally in section \ref{app-const}, we present a simple stream application constructed from operations such as compression and encryption in order to evaluate our framework.
+\comment{The extent to which we extract parallelism ranges, not only to the packet tasks, but also to the low level access of data I/O to the network interface hardware through custom additions to the \texttt{netmap}.} Finally in section \ref{app-const}, we present a simple stream application constructed from operations such as compression and encryption and evaluate our framework.
 
@@
 
-\subsection{Userspace Packet I/O}
-
-For a packet processing system to be integrated with an application, it is necessary to have an efficient mechanism for accessing the data from the NIC without interacting with the operating system.
+\subsection{Communication filters}
+
+\comment{Understanding the set of packet processing tasks associated with the application is essential to performing them in parallel.} We use a graph-based representation of filters related to packet processing to determine the dependencies involved for the different stages and to determine the relation with the different tasks of the application.
+This graph contains the stages through which the data associated with the application is transformed into network packets and vice versa.
+It shows which of the tasks can be performed in parallel with those of the application, and highlights those that have a data dependency on a specific task in the application.
+The graph also enables optimizations based on specific characteristics of the application or the architecture used for execution.
+Figure \ref{fig2} shows a graph representing User Datagram Protocol (UDP) packet generation.
+
+\begin{figure}[ht]
+\centering
+\includegraphics[width=3.5in]{pack-op}
+\caption{Simplified UDP packet construction}
+\label{fig2}
+\end{figure}
+
+\subsection{Userspace communication}
+
+For a communication system to be integrated with an application, it is necessary to have an efficient mechanism for reading and writing packets to and from the NIC without interacting with the operating system.
 Rapid data access here is crucial for the extraction of data parallelism at later stages.
@@
 By offloading the classification of packets to the NIC, we remove any overhead associated with packet classification in the application.
-We can also use these hardware queues as a means of balancing the workload across different threads by controlling the number of queues that each process handles.
+We can also use these hardware queues as a means of balancing the workload across different threads by controlling the number of queues that each thread handles.
 \comment{Since different processes can access these transmit and receive queues, concurrent access to data is possible.}
 
-\subsubsection{Flow based Packet Batching}
-
-Packet batching is a common method used to boost the performance of network applications where several packets are combined and operated upon together, thus reducing the overhead of I/O operations.
+\subsubsection{Flow based communication}
+
+Combining multiple packets that belong to a flow known as \texttt{batching} is a common method used to boost the performance of network applications where several packets are combined and operated upon together, thus reducing the overhead of I/O operations.
 In our framework, batching is utilized in a form where the number of packets in a batch is varied based on the available space in the hardware queue.
@@
 
-System calls are required in order to synchronize data between the NIC hardware and the buffer memory used for intermediate storage of packets.
+System calls are required in order to synchronize data between the NIC hardware and the buffer memory used for intermediate storage of data.
 Previous work describes the overhead associated with per-packet system calls and provides mechanisms for overcoming this overhead\cite{1564468}\cite{Han:2010:PGS:1851275.1851207}\cite{Rizzo:2012:RNI:2090147.2103536}.
@@
 
-\subsection{Packet Processing Graph}
-
-\comment{Understanding the set of packet processing tasks associated with the application is essential to performing them in parallel.} We use a graph-based representation of packet processing tasks to determine the dependencies involved for the different stages and to determine the relation with the different tasks of the application.
-This graph contains the stages through which the data associated with the application is transformed into network packets and vice versa.
-It shows which of the tasks can be performed in parallel with those of the application, and highlights those that have a data dependency on a specific task in the application.
-The graph also enables optimizations based on specific characteristics of the application or the architecture used for execution.
-Figure \ref{fig2} shows a graph representing User Datagram Protocol (UDP) packet generation.
-
-\begin{figure}[ht]
-\centering
-\includegraphics[width=3.5in]{pack-op}
-\caption{Simplified UDP packet construction}
-\label{fig2}
-\end{figure}
-
-\subsection{Application Task Graph}
-
-\comment{Application task graph represents the operations associated in performing the computations.} The subset of the application that we are interested in parallelizing is the stream processing component.
-In such components, the communication flow between tasks is well-defined.
-Both the packet processing and application tasks can be represent by a directed acyclic graph (DAG) similar to the packet processing graph.
-The packet processing tasks operate on packets and the application tasks operate on the data.
+\subsection{Computation filters}
+
+\comment{Computation filters represents the operations associated .
+The subset of the application that we are interested in parallelizing is the stream processing component.} In stream processing model, the communication flow between tasks is well-defined.
+\comment{This is usually represented by a synchronous data flow graph (SDF).} The filters associated with the computation can be connected to the communication filters in the same directed acyclic graph (DAG) used to represent the application.
+\comment{The packet processing tasks operate on packets and the application tasks operate on the data.
 Since we focus more on parallel packet processing, optimizations in relation to the application graph are not discussed.
-It is assumed that a static version of the application graph is available and we focus more on executing these tasks in parallel with the packet processing tasks.
+It is assumed that a static version of the application graph is available and we focus more on executing these tasks in parallel with the packet processing tasks.} The computation filters are designed by the user and are programmed into the runtime framework which already consists of the communication filters.
 
@@
 
-The sequence of tasks that can be executed in parallel is determined at this stage.
+\comment{The sequence of tasks that can be executed in parallel is determined at this stage.
 The integration stage is used to combine the application graph and the packet processing graph into a larger graph that unifies and represents the overall operations.
@@
 We determine the sequence of tasks that can be executed in parallel from both the application task graph and packet processing graph.
-The combined set of tasks is then used by the parallel processing engine.
+The combined set of tasks is then used by the parallel processing engine.}
+
+This stage essentially connects the corresponding communication filter to the computation filter defined earlier.
+Based on the stream graph that is provided, the necessary communication filters are added.
+This also determines the unique \textit{flows} that exists between the filters and is used later to replicate and execute them in parallel.
 
@@
 
-Two important aspects that the parallel processing engine must address is the scheduling of tasks and balancing the load across the available resources.
+Two important aspects that the parallel processing engine must address is the scheduling of filters and balancing the load across the available resources.
 The parallel processing engine consists of collaborating processes which are bound to specific CPUs and act as engines to execute the workload.
-Based on their requirements and the available resources, tasks are scheduled across these execution engines.
+Based on their requirements and the available resources, filters are scheduled across these execution engines.
+We assume a static schedule of the filters on to the given multiple cores is provided for execution.
 \comment{In this manner, the workload can be distributed flexibly.}
@@
 More specifically, they do not have any way of optimizing the rate of processing, lacking, for example, the concept of flows for application tasks.
+The application tasks are essentially the computation filters defined in our design principles.
 The lack of support for interacting with these interfaces in order to read and write data in parallel from concurrent processes or threads adds to the problem.
@@
 The $f(n)$ present in Algorithm \ref{s_pack} line \ref{s_pack_param} and Algorithm \ref{r_pack} line \ref{r_pack_param} represents this.
-Choosing a random queue has proven to be effective in our experiments, but if an application requires a specific flow to be maintained, a different function here could provide that support.
+Choosing a random queue ($f(n)=X \sim U([0,n])$ n=no of queues)has proven to be effective in our experiments, but if an application requires a specific flow to be maintained, a different function here could provide that support.
 This would be helpful in cases where the data operations are more predictable, or if the transactions seem to follow a specific pattern.
@@
 
-After integrating the packet operations by means of support from \texttt{Click}'s infrastructure and access to network hardware through \texttt{netmap}'s APIs, we integrate these operations with the application graph.
-The results is a complete task graph that does both the computation and communication operations.
-The next logical step is executing these tasks in parallel by exploiting data and task parallelism.
+After defining the communication filters by means of \texttt{Click}'s infrastructure and access to network hardware through \texttt{netmap}'s APIs, we integrate these filters with the application tasks.
+The results is a complete stream graph that does both the computation and communication operations.
+The next logical step is executing these filters in parallel by exploiting data parallelism.
 Packet streams are inherently operable in a data parallel fashion i.e.
 it is possible to operate on two different streams simultaneously.
-Since the application tasks depend on the data packets, they could also be executed in parallel.
+\comment{Since the application tasks depend on the data packets, they could also be executed in parallel.
 So from the integrated graph constructed by combining the application and packet processing operations, it becomes possible to extract data parallelism by operating on different data streams simultaneously.
-By executing different components, representing distinct operations on the data streams, it is also possible to exploit task parallelism.
+By executing different components, representing distinct operations on the data streams, it is also possible to exploit task parallelism.}
 
@@
 
-Once we have scheduled our execution engines on the physical system, the next step is to execute the integrated task graph consisting of application and packet operations.
-To extract data parallelism, we replicate the task graph onto multiple execution engines.
-This ensures operations on the same data are performed on the same CPU, thereby using local cache.
-By making sure that packets belonging to a specific flow always end up in the same execution engine, we only perform operations conforming to that stream of data.
+Once we have scheduled our execution engines on the physical system, the next step is to execute the stream graph for different \textit{flows}.
+To extract data parallelism, we replicate the stream graph onto multiple execution engines.
+This ensures operations on the same \textit{flow} are performed on the same CPU, thereby using local cache.
+By making sure that packets belonging to a specific f\textit{flow} always end up in the same execution engine, we only perform operations conforming to that stream of data.
 \comment{We use the ability of modern NICs to classify packets onto different queues, based on their attributes.}
@@
 
-In order to replicate the task graph we had to overcome some restrictions.
+\comment{In order to replicate the task graph we had to overcome some restrictions.
 First was to ensure that there was no intricate data dependency between different data streams.
@@
 MPI communication is used to schedule the application and packet operations on the processes representing the execution engines.
-Furthermore, we use it to also perform explicit data sharing between tasks in the application that require it.
+Furthermore, we use it to also perform explicit data sharing between tasks in the application that require it.}
 
@@
 These servers are connected using Intel 8299EB 10Gbe adapters over directly attached twin-axial cables.
-The adapters are located in the PCI Express x8 slot to maximize the available bandwidth.
+The adapters are located in the PCI Express x8 slot to provide required bandwidth for the network cards.
 This forms a closed connection between two servers and limits the influence of external devices such as switches or routers on data transmission.
@@
 Each application is constructed as two parts, one a source of data and the other a sink.
-We measure the overall time it takes for application task operations, packet operations, and the time to send from the source and receive at the sink.
+We measure the overall time it takes for application task operations, communication operations, and the time to send from the source and receive at the sink.
 This is done for different packet sizes and applications.
 Then we plot the speedup achieved in each of the cases over the serial version where we have a single sender and receiver operating on the multiple queues.
-This is the baseline in our experiments and is similar to how the OS behaves for transmitting and receiving using the multi queue network controller.
+This is the baseline in our experiments and is similar to how the application would behave if it were implemented using just direct access to the network hardware.
 
@@
 Since compression and encryption applications are relatively common, we chose two applications composed from these operations.
-The integrated graph that represents both these applications is shown in fig \ref{fig-app-graph}.
+The stream graph that represents both these applications is shown in fig \ref{fig-app-graph}.
 UDP is used to send the data between source and sink.
@@
 In the encryption application, there is not as much asymmetry and packet size does not matter as much.
-The fact that packet size impacts applications differently suggests that is essential to incorporate dynamic packet-size selection into the framework itself.\comment{ to respond to the unique requirements of each application.}
+The fact that packet size impacts applications differently suggests that it is essential to incorporate dynamic packet-size selection into the framework itself.\comment{ to respond to the unique requirements of each application.}
 
@@
 In summary, it can be seen that the framework scales well for different workloads, can handle both computations and communication, and can operate on them in parallel.
+We see that we can achieve a speedup $10\times$ for compression and $12\times$ for encryption sample stream application respectively on the architecture consisting of two nodes of 24 logical cores each.
 In some cases, the framework can be improved by having a greater degree of control over the fine grained parallelism within the execution engines.
@@
 
+\comment{Results also suggest where efforts should be made in developing the framework further.
+Until now, the focus has been on deploying the execution engines in order to execute stream applications and to enable optimal communication between separate nodes.}
+
+\comment{Future work will address some of the opportunities presented by these execution engines and the collaboration between them.}
+
+Since we work with a static schedule provided by the user in executing the applications.
+Dynamic load balancing is likely to be the most benificial addition to the framework in terms of performance of the stream applications built using it.
+It is also possible to add the capability to dynamically adjust the number of network queues allocated to each core as stream application requirements change over time.
+Finally, as seen in Figure~\ref{fig-res}, being able to adjust packet sizes dynamically to an optimal value for each network link, dependant on the stream application, would also be a useful addition to the framework.
+
 \section{Related Work}
 
+Stream programming languages such as StreamIt~\cite{Thies:2002:SLS:647478.727935}, LUSTRE~\cite{Halbwachs91thesynchronous} and extensions such as Carpenter et al.~\cite{Carpenter:2007:SMD:1776200.1776217}, Buck et al.~\cite{Buck:2004:BGS:1015706.1015800} target stream applications but the focus is more on parallelizing the computations related to the application alone on the multicore system unlike our runtime framework which parallelizes communication operations also.
+
+MPICH-G2~\cite{Karonis02mpich-g2:a} targets large grid networks and for MPI programming model unlike us which is focused more on streaming model.
+Moreover, we do not employ any of the MPI communication primitives provided and use our own runtime system to communicate data between the filters in the stream graph across systems.
+
+Wagner et al.~\cite{5160944} propose a stream processing framework that completely utilizes the MPI library.
+They make use of MPI groups and communicators to improve the flexibility of MPI library to support stream-processing.
+Our runtime framework is designed to provide a low overhead and parallel communication framework by directly utilizing the network link.
+Whereas they employ a library using MPI primitives to construct their stream-processing system and rely on the underlying abstraction layer to communicate over any network the MPI library supports.
+
+Mancini et al.~\cite{5493465} propose a hybrid approach of Stream and MPI programming models and use a whole MPI program as building block in a stream application to improve processing speeds of computational units in hetrogeneous computational systems.
+
+While we use MPI to execute the computation filters, our runtime framework employs a direct and parallel communication interface for the actual transmission of reception of data between filters.
+This is the key feature that enables us to parallelize the communication along with the computation for a given streaming application, which is essential in improving its performance.
+
+\comment{
 Several prior researchers \cite{Dobrescu09routebricks:exploiting}\cite{Han:2010:PGS:1851275.1851207}\cite{Wang:2009:PPN:1542275.1542307}\cite{springerlink:10.1007/s11227-011-0579-3} have worked on using commodity multicore processors to perform network operations.
@@
 Egi et al.\cite{springerlink:10.1007/s11227-011-0579-3} shows techniques on parallelizing network operations on multicore architectures by defining principles that consider multiple resources and not just the CPU alone.
-
-Stream programming languages such as StreamIt \cite{Thies:2002:SLS:647478.727935}, LUSTRE \cite{Halbwachs91thesynchronous}, etc., target stream applications but the focus is more on parallelizing the computations related to the application alone on the multicore system.
+}
 
@@
 In our work even though a part of it involves parallelizing network operations, the objective of this research is to provide support for parallelizing stream application and distributing their workloads over multiple systems.
-This also requires optimizing the underlying network operations by parallelizing it.}
+This also requires optimizing the underlying network operations by parallelizing it.
 
@@
 Click implements its own mechanism and polling driver in kernel space to access packets, bypassing the OS's network infrastructure.
+}
 
@@
 \section{Conclusions}
-In this paper we have proposed a high-level programming framework that can integrate packet processing within stream applications and perform both types of operation in parallel.
-
-We conclude that by using our algorithms for sending and receiving data, which implicitly perform batching and reduce the usage of system calls, and by combining this with \texttt{netmap} APIs which minimizes the interaction with the OS, communication performance in our framework never becomes the bottleneck in our system on the programs we have tested.
-Instead of performing the packet operations in the OS, by integrating it with the application using Click's elements within our infrastructure, we are able to construct a unified task graph that represents the computation and communication operations.
-Using this, we are able to parallelize these operations together and achieve speedups of more than a factor of eight in all the applications we tested.
-
-In integrating these features into a programming framework, the system of multiple queues, parallel communication and computation are entirely hidden from the programmer, who merely specifies a standard stream graph for computation alone.
-This frees up the programmer from the onerous task of having to decide how their application should be distributed, leaving it instead to the execution engine of our framework, which also looks after the ensuing communication requirements after application distribution.
+In this paper we have proposed a parallel runtime framework that can integrate communication operations with the stream applications and perform both types of operation in parallel.
+
+Our algorithms for sending and receiving data, which implicitly perform batching and reduce the usage of system calls, and by combining this with \texttt{netmap} APIs which minimizes the interaction with the OS, communication performance in our framework is able to scale with available processors.
+Instead of performing the communication operations separately in the OS, by integrating them with the stream application \comment{using Click's elements within our infrastructure,} we are able to construct a unified stream graph that represents the computation and communication filters.
+Using this, we are able to parallelize the stream application and achieve speedups of more than a factor of eight in all the applications we tested.
+
+In integrating these features into a runtime framework, the system of multiple queues, parallel communication and computation are entirely hidden from the programmer, who merely specifies a standard stream graph for computation alone.
+\comment{This frees up the programmer from the onerous task of having to decide how their application should be distributed, leaving it instead to the execution engine of our framework, which also looks after the ensuing communication requirements after application distribution.}
 
 The results show that our system scales to as many as 24 parallel processes on a multicore computer, and achieves speedups of more than a factor of ten in some cases compared to sequential implementations.
-Results also suggest where efforts should be made in developing the framework further.
-Until now, the focus has been on deploying the execution engines in order to execute stream applications and to enable optimal communication between separate nodes.
-
-Future work will address some of the opportunities presented by these execution engines and the collaboration between them.
-Dynamic load balancing is likely to be the most benificial addition to the framework in terms of performance of the stream applications built using it.
-It is also possible to add the capability to dynamically adjust the number of network queues allocated to each core as stream application requirements change over time.
-Finally, as seen in section \ref {results}, being able to adjust packet size to an optimal value for each network link, again dependent on stream application requirements, would also be a useful addition to the framework.
 
@@
 % argument is your BibTeX string definitions and bibliography database(s)
-\bibliography{IEEEabrv,ref.bib}
+\bibliography{ref.bib}
 %

### 2013-03-21 paper.tex
david's morning changes

@@
 
-Applications that follow streaming model~\cite{Thies:2002:SLS:647478.727935,Halbwachs91thesynchronous,Stephens95asurvey,muthukrishnan2005data} are expressed as a set of filters that operate on one or more tokens of data.
-The flow of these data tokens through the filters is referred to as a \textit{stream}.
-Since the data is accessed sequentially in the form of a stream the memory requirement to process the data is considerably less compared to the size of the enitre stream, thereby making it possible to operate on very large data sizes.
-Typical applications based on this model operate on data ranges in the order of terabytes or even bigger.
-Applications such as financial trading systems~\cite{agarwal2009faster}, database systems~\cite{Golab03datastream}, network data analysis~\cite{Gilbert01quicksand:quick,Sullivan:1998:TSM:1268256.1268258} are few examples that follow the streaming model and operate on very large data sets.
-
-Traditional stream applications target multicore processors for performing their computations in parallel.
-As multicore processors grow in size, so do stream applications, processing more streams of data in parallel.
-However, in such cases the applications are typically limited by the traditional communication systems~\cite{Wu:2007:PPB:1323954.1323957,Wu:2007:PAL:1227865.1228021,Dobrescu09routebricks:exploiting,Han:2010:PGS:1851275.1851207,Kohler2000}.
-The reason being these systems have limited support for operating on parallel streams and are burdened by abstraction layers that provide compatibility for non-streaming applications.
+Applications that follow streaming model~\cite{Thies:2002:SLS:647478.727935,Halbwachs91thesynchronous,Stephens95asurvey,muthukrishnan2005data}
+are expressed as a network of \textit{filters} connected by communication
+\textit{channels}. Tokens of data flow along the communication channels
+between filters. Each filter consumes on one or more \textit{streams}
+of data from its input channels, and produces streams of data on its
+ouput channels. For applications that fit the model well, stream processing
+can greatly reduce the working memory required. Since the data flows through
+a sequence of filters in a pipelined style, it is often possible to
+operate on a relatively small window of the data at any given time,
+thereby making it possible to operate on conceptually infinite
+data streams.
+
+Traditionally, the stream processing model was used primarily
+signal processing type applications, where the application
+processes a continuous stream of signal inputs. For example,
+many video processing applications are modelled as a set of
+image filters operating on a stream of video frame, where the
+entire video may be hundreds of gigabytes in size, but the
+processing can be done with just a few frames at a time.
+However, in more recent years it has been found that the stream
+model is highly suitable for many applications that process
+very large data sets, in the order of terabytes or even larger.
+These include applications such as financial trading
+systems~\cite{agarwal2009faster},
+database systems~\cite{Golab03datastream},
+network data analysis~\cite{Gilbert01quicksand:quick,Sullivan:1998:TSM:1268256.1268258} are few examples that follow the streaming model and operate on very large data sets.
+
+Stream processing is highly suited to multicore and parallel
+computing, because the filters can typically execute in parallel.
+Smaller stream applications often achieve high levels of parallelism
+on multicore machines.  However, as the stream processing model
+is used to address larger data processing problems, more computational
+resources are needed, and a common solution is to distribute the
+filters across a number of machines.
+In such cases the applications are typically limited by the
+traditional communication
+~\cite{Wu:2007:PPB:1323954.1323957,Wu:2007:PAL:1227865.1228021,Dobrescu09routebricks:exploiting,Han:2010:PGS:1851275.1851207,Kohler2000}.
+The reason being these systems have limited support for operating on
+parallel streams and are burdened by abstraction layers that provide
+compatibility for non-streaming applications.
 

### 2013-03-21 paper.tex
kevins last min changes

@@
 Trinity College Dublin, Dublin 2, Ireland\\
-Email: muralis@scss.tcd.ie, David.Gregg.cs.tcd.ie }
-\IEEEauthorblockA{\IEEEauthorrefmark{4}DCU\\
-Email: }
+Email: muralis@scss.tcd.ie, David.Gregg@cs.tcd.ie }
+\IEEEauthorblockA{\IEEEauthorrefmark{4}School of Computing, Dublin City University, Glasnevin, Dublin 9, Ireland\\
+Email: Kevin.Casey@computing.dcu.ie}
 }
@@
 Stream applications are often limited in their performance by their underlying communication system.
-In a typical implementation, the underlying communication relies on operating system to handle majority of the network operations.
+A typical implementation relies on the operating system to handle the majority of network operations.
 In such cases, the communication stack, which was not designed to handle such tremendous amounts of data, acts as a bottleneck and restricts the performance of the application.
@@
 We also utilize the ability of the modern network controller to classify packets onto multiple hardware queues by integrating it with the communication filter.
-Using this our runtime framework is able to construct a unified stream graph based on the application graph provided.
-This when used with a static schedule based on a given architecture, can process the data streams in parallel.
-
-We are able to parallelize the stream application and achieve speedups of more than a factor of eight in all the applications we tested.
+Using this, our runtime framework is able to construct a unified stream graph based on the application graph provided.
+This, when used with a static schedule based on a given architecture, can process the data streams in parallel.
+
+We are able to parallelize stream applications and achieve speedups of more than a factor of eight in all the applications we tested.
 The results show that our system scales to as many as 24 parallel processes on a multicore computer, and achieves speedups of more than a factor of ten in some cases compared to sequential implementations.
@@
 
-Applications that follow streaming model~\cite{Thies:2002:SLS:647478.727935,Halbwachs91thesynchronous,Stephens95asurvey,muthukrishnan2005data}
+Applications that follow a streaming model~\cite{Thies:2002:SLS:647478.727935,Halbwachs91thesynchronous,Stephens95asurvey,muthukrishnan2005data}
 are expressed as a network of \textit{filters} connected by communication
@@
 Traditionally, the stream processing model was used primarily
-signal processing type applications, where the application
+for signal processing type applications, where the application
 processes a continuous stream of signal inputs. For example,
@@
 very large data sets, in the order of terabytes or even larger.
-These include applications such as financial trading
+Applications such as financial trading
 systems~\cite{agarwal2009faster},
 database systems~\cite{Golab03datastream},
-network data analysis~\cite{Gilbert01quicksand:quick,Sullivan:1998:TSM:1268256.1268258} are few examples that follow the streaming model and operate on very large data sets.
+network data analysis~\cite{Gilbert01quicksand:quick,Sullivan:1998:TSM:1268256.1268258} are a few examples that follow the streaming model and also operate on very large data sets.
 
@@
 filters across a number of machines.
-In such cases the applications are typically limited by the
+In such cases, the applications are typically limited by the
 traditional communication
 ~\cite{Wu:2007:PPB:1323954.1323957,Wu:2007:PAL:1227865.1228021,Dobrescu09routebricks:exploiting,Han:2010:PGS:1851275.1851207,Kohler2000}.
-The reason being these systems have limited support for operating on
+The main reason for this is that these systems have limited support for operating on
 parallel streams and are burdened by abstraction layers that provide
@@
 
-Several high speed networks~\cite{fredj2011survey} exists that could handle such communication requirements but these require expensive hardware and specialized programming skills.
-Commodity network adapters based on 10-gigabit ethernet (10GbE) are comparatively cheaper alternative and are being widely deployed in clusters and grids in data centers.
-Previous research\cite{Romanow03anoverview,Feng:2003:OEN:1048935.1050200} have been done on optimizing ethernet for improving communication in such cases.
+Several high speed networks~\cite{fredj2011survey} exist that could handle such communication requirements but these require expensive hardware and specialized programming skills.
+Commodity network adapters based on 10-gigabit ethernet (10GbE) are a comparatively cheaper alternative and are being widely deployed in clusters and grids in data centers.
+Previous research\cite{Romanow03anoverview,Feng:2003:OEN:1048935.1050200} has been done on optimizing ethernet for improving communication in such cases.
 However, they do not use many of the features provided by modern day network adapters such as hardware queues, flow control, etc., Also, they do not focus specifically on streaming applications and hence making it difficult to utilize them properly.
@@
 In this paper we propose a parallel processing framework for building stream applications that are bound by very large communication.
-A simplistic framework where the filters are described as \textit{actors} and the application is represented by connecting such \textit{actors} into a directed acyclic graph (DAG).
+A simple framework is used, where the filters are described as \textit{actors} and the application is represented by connecting such \textit{actors} into a directed acyclic graph (DAG).
 Then our framework is used to establish, (1)the communication system that allows these \textit{actors} to send and receive data across each other and (2)the computation system that executes these tasks in parallel on a given multicore system.
@@
 \begin{enumerate}
-\item Parallel multi point communication layer, which provides the ability for stream actors to communication across the network.
-\item Execution system that can run the given data parallel filters on the multicore system
-\item Utilizes network hardware features such as queues and flow control to support stream applications
+\item A parallel multi point communication layer, which provides the ability for stream actors to communicate across the network
+\item An execution system that can run the given data-parallel filters on the multicore system
+\item The utilization of network hardware features such as queues and flow control to support stream applications.
 
@@
 
-Our main objective is to optimize performance of streaming applications by First, providing a speaclized communication link between the filters and Second, extracting data parallel actors and executing them in parallel.
+Our main objective is to optimize performance of streaming applications by first, providing a specialized communication link between the filters and second, extracting data-parallel actors and executing them in parallel.
 Stream applications usually consists of stateless and stateful filters.
-Stateful filters are those that are depedendent on the previous execution and hence making it difficult to replicate them.
+Stateful filters are those that are depedendent on the previous execution which makes it difficult to replicate them.
 Stateless filters do not pose such limitations and can be freely replicated based on the available data streams.
@@
 All the filters used in this application are essentially stateless filters.
-This sample application operates by performing encryption on the data that matches predefined rules like file names, access level of user, etc., otherwise it performs compression.
+This application operates by performing encryption on the data that matches predefined rules like file names, access level of user, etc., otherwise it performs compression.
 Due to the nature of such stream applications, it is possible to replicate and distribute the filters across different CPUs.
@@
 In building a framework for such applications on general purpose computers, we can exploit two hardware features that could allow for dramatically improved performance, namely multicore processors and multi queue network controllers.
-Using the multiple queues it makes it possible to do much of the communication in parallel and using the multiple cores the filters from the stream application can be executed concurrently.
+Using the multiple queues makes it possible to do much of the communication in parallel and by using the multiple cores, the filters from the stream application can be executed concurrently.
 
@@
 In our framework, we use a more elegant approach.
-We define a runtime framework that already consists of the communication filters that are pre-built.
-So inorder to implement a stream application such as that shown in Figure~\ref{fig1}, all we need to do is define the computation filters and connect them in the framework to the pre-built communication filters.
-On execution, our runtime framework is able to replicate the stateless filters and the communication filters on to the different multiple cores, thereby executing the operations belonging to different \textit{flows} in parallel.
+We define a runtime framework that provides pre-built communication filters.
+In order to implement a stream application such as that shown in Figure~\ref{fig1}, all we need to do is define the computation filters and connect them in the framework to the pre-built communication filters.
+On execution, our runtime framework is able to replicate the stateless filters and the communication filters onto the different multiple cores, thereby executing the operations belonging to different \textit{flows} in parallel.
 
@@
 The graph also enables optimizations based on specific characteristics of the application or the architecture used for execution.
-Figure \ref{fig2} shows a graph representing User Datagram Protocol (UDP) packet generation.
+Figure \ref{fig2} shows a simple graph representing User Datagram Protocol (UDP) packet generation.
 
@@
 Rapid data access here is crucial for the extraction of data parallelism at later stages.
-\comment{In using the techniques described below we exploit the modern NIC architecture to its full potential.}
+We use a number of techniques to support this rapid data access.\comment{In using the techniques described below we exploit the modern NIC architecture to its full potential.}
 
@@
 
-The development of virtualization and improved network flow handling has led to hardware level multi queue support in NICs.
+The development of virtualization and improved network flow handling has led to hardware-level multi queue support in NICs.
 During transmission, concurrent writes to different queues are possible, enabling multiple processes to send data simultaneously.
@@
 
-Combining multiple packets that belong to a flow known as \texttt{batching} is a common method used to boost the performance of network applications where several packets are combined and operated upon together, thus reducing the overhead of I/O operations.
+A common technique known as \texttt{batching} is used to boost the performance of network applications, where several packets are combined and operated upon together, thus reducing the overhead of I/O operations.
 In our framework, batching is utilized in a form where the number of packets in a batch is varied based on the available space in the hardware queue.
@@
 Since we focus more on parallel packet processing, optimizations in relation to the application graph are not discussed.
-It is assumed that a static version of the application graph is available and we focus more on executing these tasks in parallel with the packet processing tasks.} The computation filters are designed by the user and are programmed into the runtime framework which already consists of the communication filters.
+It is assumed that a static version of the application graph is available and we focus more on executing these tasks in parallel with the packet processing tasks.} The computation filters are designed by the user and are programmed into the runtime framework which already provides the communication filters.
 
@@
 
-Two important aspects that the parallel processing engine must address is the scheduling of filters and balancing the load across the available resources.
+Two important aspects that the parallel processing engine must address is the scheduling of filters and the balancing of the load across the available resources.
 The parallel processing engine consists of collaborating processes which are bound to specific CPUs and act as engines to execute the workload.
 Based on their requirements and the available resources, filters are scheduled across these execution engines.
-We assume a static schedule of the filters on to the given multiple cores is provided for execution.
+We assume a static schedule of the filters onto the given multiple cores is provided for execution.
 \comment{In this manner, the workload can be distributed flexibly.}
@@
 More specifically, they do not have any way of optimizing the rate of processing, lacking, for example, the concept of flows for application tasks.
-The application tasks are essentially the computation filters defined in our design principles.
+These application tasks are represented as computation filters similar to those described in figure~\ref{fig1}.
 The lack of support for interacting with these interfaces in order to read and write data in parallel from concurrent processes or threads adds to the problem.
@@
 \texttt{Netmap} uses a set of ring buffers in kernel space and user space to export the network hardware buffer's memory region.
-This enables the applications in user space to directly access the network hardware.
-It also prevents the application from accessing memory outside valid regions by checking the limits of the ring buffer.
-Packets can be written to the buffers when space is available and an \textit{ioctl} call to the driver enables the network controller to send the data.
+This enables applications in user space to directly access the network hardware.
+It also prevents applications from accessing memory outside valid regions by checking the limits of the ring buffer.
+Packets can be written to the buffers when space is available, and an \textit{ioctl} call to the driver enables the network controller to send the data.
 A \textit{poll} interface is implemented on the memory region mapped in user space to check when new packets are available.
@@
 \item Reduce the usage of system calls such as \textit{ioctl} and \textit {poll}
-\item Provide mechanisms for handling multiple queues through the application
+\item Provide mechanisms for handling multiple queues through the application.
 
@@
 The $f(n)$ present in Algorithm \ref{s_pack} line \ref{s_pack_param} and Algorithm \ref{r_pack} line \ref{r_pack_param} represents this.
-Choosing a random queue ($f(n)=X \sim U([0,n])$ n=no of queues)has proven to be effective in our experiments, but if an application requires a specific flow to be maintained, a different function here could provide that support.
+Choosing a random queue ($f(n)=X \sim U([0,n])$ n=no of queues) has proven to be effective in our experiments, but if an application requires a specific flow to be maintained, a different function here could provide that support.
 This would be helpful in cases where the data operations are more predictable, or if the transactions seem to follow a specific pattern.
@@
 
-In order to evaluate our framework, we test the scalability and the performance speedup on a multicore system for an application that is constructed using the framework.
-Using two generic operations: compression and encryption, we build simple stream applications using our framework.
+In order to evaluate our framework, we test the scalability and the performance speedup on a multicore system for an application that is constructed with the framework.
+Using two generic operations: compression and encryption, we build simple stream applications.
 We then measure the performance improvement as we increase the number of execution engines dedicated to the application.
@@
 
-Each application is constructed as two parts, one a source of data and the other a sink.
+Each application is constructed in two parts, one a source of data and the other a sink.
 We measure the overall time it takes for application task operations, communication operations, and the time to send from the source and receive at the sink.
@@
 Once data is received, packet operations are performed to extract the payload, followed by decompression or decryption to retrieve the original data.
-These are representative of some of the basic operations that would be involved in a much more complex application.
+These are representative of some of the typical operations that would be involved in a much more complex application.
 \comment{Choosing to compare these simple operations avoids any extraneous factors that may be present in a more complex application.}Use of these simpler stream applications enables us to evaluate the performance of our framework and not that of any optimizations specific to the application.
@@
 One is that, from fig\ref{fig-res}.(d,e,f), we note that compression requires more work than decompression, whereas from fig\ref{fig-res}.(a,b,c) encryption and decryption are more balanced.
-This is evident from the continued speedup we get up to 4 senders in the case of compression/decompression.
+This is evident from the continued speedup obtained from up to 4 senders in the case of compression/decompression.
 Only a slight asymmetry can be seen in the encryption/decryption case, where adding more senders than receivers only yields a slight increase in performance.
@@
 
-Stream programming languages such as StreamIt~\cite{Thies:2002:SLS:647478.727935}, LUSTRE~\cite{Halbwachs91thesynchronous} and extensions such as Carpenter et al.~\cite{Carpenter:2007:SMD:1776200.1776217}, Buck et al.~\cite{Buck:2004:BGS:1015706.1015800} target stream applications but the focus is more on parallelizing the computations related to the application alone on the multicore system unlike our runtime framework which parallelizes communication operations also.
-
-MPICH-G2~\cite{Karonis02mpich-g2:a} targets large grid networks and for MPI programming model unlike us which is focused more on streaming model.
-Moreover, we do not employ any of the MPI communication primitives provided and use our own runtime system to communicate data between the filters in the stream graph across systems.
+Stream programming languages such as StreamIt~\cite{Thies:2002:SLS:647478.727935}, LUSTRE~\cite{Halbwachs91thesynchronous} and extensions such as Carpenter et al.~\cite{Carpenter:2007:SMD:1776200.1776217}, Buck et al.~\cite{Buck:2004:BGS:1015706.1015800} target stream applications, but the focus is more on parallelizing the computations related to the application alone on the multicore system unlike our runtime framework which also parallelizes communication operations.
+
+MPICH-G2~\cite{Karonis02mpich-g2:a} targets large grid networks and applications written using the MPI programming model, unlike our focus on the streaming model.
+Moreover, we do not employ any of the MPI communication primitives provided to handle streaming data.
+Instead we use our own fast runtime system to communicate data between the filters in the stream graph across systems.
 
 Wagner et al.~\cite{5160944} propose a stream processing framework that completely utilizes the MPI library.
-They make use of MPI groups and communicators to improve the flexibility of MPI library to support stream-processing.
-Our runtime framework is designed to provide a low overhead and parallel communication framework by directly utilizing the network link.
-Whereas they employ a library using MPI primitives to construct their stream-processing system and rely on the underlying abstraction layer to communicate over any network the MPI library supports.
+They make use of MPI groups and communicators to improve the flexibility of the MPI library to support stream-processing.
+Their approach employs a library using MPI primitives to construct their stream-processing system and relies on the underlying abstraction layer to communicate over any network the MPI library supports.
+Our runtime framework is designed to provide a low overhead, parallel communication framework by directly utilizing the network link.
 
@@
 
-While we use MPI to execute the computation filters, our runtime framework employs a direct and parallel communication interface for the actual transmission of reception of data between filters.
+While we use MPI to execute the computation filters, our runtime framework employs a direct and parallel communication interface for the actual transmission and reception of data between filters.
 This is the key feature that enables us to parallelize the communication along with the computation for a given streaming application, which is essential in improving its performance.
@@
 \section{Conclusions}
-In this paper we have proposed a parallel runtime framework that can integrate communication operations with the stream applications and perform both types of operation in parallel.
-
-Our algorithms for sending and receiving data, which implicitly perform batching and reduce the usage of system calls, and by combining this with \texttt{netmap} APIs which minimizes the interaction with the OS, communication performance in our framework is able to scale with available processors.
-Instead of performing the communication operations separately in the OS, by integrating them with the stream application \comment{using Click's elements within our infrastructure,} we are able to construct a unified stream graph that represents the computation and communication filters.
-Using this, we are able to parallelize the stream application and achieve speedups of more than a factor of eight in all the applications we tested.
-
-In integrating these features into a runtime framework, the system of multiple queues, parallel communication and computation are entirely hidden from the programmer, who merely specifies a standard stream graph for computation alone.
-\comment{This frees up the programmer from the onerous task of having to decide how their application should be distributed, leaving it instead to the execution engine of our framework, which also looks after the ensuing communication requirements after application distribution.}
+In this paper we have proposed a parallel runtime framework that can integrate communication and computational operations in stream applications and perform both types of operation in parallel.
+
+Our algorithms for sending and receiving data implicitly perform batching and reduce the usage of system calls, and we combine this with \texttt{netmap} APIs to minimize the interaction with the OS.
+This allows communication performance in our framework to better scale with the number of available processors.
+Instead of performing the communication operations separately in the OS, by integrating them within the stream application\comment{using Click's elements within our infrastructure,}, we are able to construct a unified stream graph that represents the entire application, including computation and communication filters.
+Using this, we are able to parallelize these stream applications and achieve speedups of more than a factor of eight in all the applications we tested.
+
+By integrating these features into a runtime framework, the system of multiple queues, parallel communication and computation are entirely hidden from the programmer, who merely specifies a standard stream graph for computation alone.
+This frees up the programmer from the onerous task of dealing with the substantial communication requirements that result from stream application distribution.
 

### 2013-05-10 paper (copy).tex
final submission for ispa 2013

%% bare_conf.tex
%% V1.3
%% 2007/01/11
%% by Michael Shell
%% See:
%% http://www.michaelshell.org/
%% for current contact information.
%%
%% This is a skeleton file demonstrating the use of IEEEtran.cls
%% (requires IEEEtran.cls version 1.7 or later) with an IEEE conference paper.
%%
%% Support sites:
%% http://www.michaelshell.org/tex/ieeetran/
%% http://www.ctan.org/tex-archive/macros/latex/contrib/IEEEtran/
%% and
%% http://www.ieee.org/

%%*************************************************************************
%% Legal Notice:
%% This code is offered as-is without any warranty either expressed or
%% implied; without even the implied warranty of MERCHANTABILITY or
%% FITNESS FOR A PARTICULAR PURPOSE! 
%% User assumes all risk.
%% In no event shall IEEE or any contributor to this code be liable for
%% any damages or losses, including, but not limited to, incidental,
%% consequential, or any other damages, resulting from the use or misuse
%% of any information contained here.
%%
%% All comments are the opinions of their respective authors and are not
%% necessarily endorsed by the IEEE.
%%
%% This work is distributed under the LaTeX Project Public License (LPPL)
%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
%% distributed and modified. A copy of the LPPL, version 1.3, is included
%% in the base LaTeX documentation of all distributions of LaTeX released
%% 2003/12/01 or later.
%% Retain all contribution notices and credits.
%% ** Modified files should be clearly indicated as such, including  **
%% ** renaming them and changing author support contact information. **
%%
%% File list of work: IEEEtran.cls, IEEEtran_HOWTO.pdf, bare_adv.tex,
%%                    bare_conf.tex, bare_jrnl.tex, bare_jrnl_compsoc.tex
%%*************************************************************************

% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
% *** with production work. IEEE's font choices can trigger bugs that do  ***
% *** not appear when using other class files.                            ***
% The testflow support page is at:
% http://www.michaelshell.org/tex/testflow/

% Note that the a4paper option is mainly intended so that authors in
% countries using A4 can easily print to A4 and see how their papers will
% look in print - the typesetting of the document will not typically be
% affected with changes in paper size (but the bottom and side margins will).
% Use the testflow package mentioned above to verify correct handling of
% both paper sizes by the user's LaTeX system.
%
% Also note that the "draftcls" or "draftclsnofoot", not "draft", option
% should be used if it is desired that the figures are to be displayed in
% draft mode.
%
\documentclass[conference]{IEEEtran}
%% INFOCOM 2013 addition:
\makeatletter
\def\ps@headings{%
\def\@oddhead{\mbox{}\scriptsize\rightmark \hfil \thepage}%
\def\@evenhead{\scriptsize\thepage \hfil \leftmark\mbox{}}%
\def\@oddfoot{}%
\def\@evenfoot{}}
\makeatother
\pagestyle{headings}
% Add the compsoc option for Computer Society conferences.
%
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
% \documentclass[conference]{../sty/IEEEtran}

% Some very useful LaTeX packages include:
% (uncomment the ones you want to load)

% *** MISC UTILITY PACKAGES ***
%
%\usepackage{ifpdf}
% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
% compilation based on whether the output is pdf or dvi.
% usage:
% \ifpdf
%   % pdf code
% \else
%   % dvi code
% \fi
% The latest version of ifpdf.sty can be obtained from:
% http://www.ctan.org/tex-archive/macros/latex/contrib/oberdiek/
% Also, note that IEEEtran.cls V1.7 and later provides a builtin
% \ifCLASSINFOpdf conditional that works the same way.
% When switching from latex to pdflatex and vice-versa, the compiler may
% have to be run twice to clear warning/error messages.

% *** CITATION PACKAGES ***
%
\usepackage{cite}
% cite.sty was written by Donald Arseneau
% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
% \cite{} output to follow that of IEEE. Loading the cite package will
% result in citation numbers being automatically sorted and properly
% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
% \cite will automatically add leading space, if needed. Use cite.sty's
% noadjust option (cite.sty V3.8 and later) if you want to turn this off.
% cite.sty is already installed on most LaTeX systems. Be sure and use
% version 4.0 (2003-05-27) and later if using hyperref.sty. cite.sty does
% not currently provide for hyperlinked citations.
% The latest version can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/cite/
% The documentation is contained in the cite.sty file itself.

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
  \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  \graphicspath{{./eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  % \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.eps}
\fi
% graphicx was written by David Carlisle and Sebastian Rahtz. It is
% required if you want graphics, photos, etc. graphicx.sty is already
% installed on most LaTeX systems. The latest version and documentation can
% be obtained at: 
% http://www.ctan.org/tex-archive/macros/latex/required/graphics/
% Another good source of documentation is "Using Imported Graphics in
% LaTeX2e" by Keith Reckdahl which can be found as epslatex.ps or
% epslatex.pdf at: http://www.ctan.org/tex-archive/info/
%
% latex, and pdflatex in dvi mode, support graphics in encapsulated
% postscript (.eps) format. pdflatex in pdf mode supports graphics
% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
% not a bitmapped formats (.jpeg, .png). IEEE frowns on bitmapped formats
% which can result in "jaggedy"/blurry rendering of lines and letters as
% well as large increases in file sizes.
%
% You can find documentation about the pdfTeX application at:
% http://www.tug.org/applications/pdftex

% *** MATH PACKAGES ***
%
%\usepackage[cmex10]{amsmath}
% A popular package from the American Mathematical Society that provides
% many useful and powerful commands for dealing with mathematics. If using
% it, be sure to load this package with the cmex10 option to ensure that
% only type 1 fonts will utilized at all point sizes. Without this option,
% it is possible that some math symbols, particularly those within
% footnotes, will be rendered in bitmap form which will result in a
% document that can not be IEEE Xplore compliant!
%
% Also, note that the amsmath package sets \interdisplaylinepenalty to 10000
% thus preventing page breaks from occurring within multiline equations. Use:
%\interdisplaylinepenalty=2500
% after loading amsmath to restore such page breaks as IEEEtran.cls normally
% does. amsmath.sty is already installed on most LaTeX systems. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/required/amslatex/math/

% *** SPECIALIZED LIST PACKAGES ***
%
\usepackage{algorithmic}
% algorithmic.sty was written by Peter Williams and Rogerio Brito.
% This package provides an algorithmic environment fo describing algorithms.
% You can use the algorithmic environment in-text or within a figure
% environment to provide for a floating algorithm. Do NOT use the algorithm
% floating environment provided by algorithm.sty (by the same authors) or
% algorithm2e.sty (by Christophe Fiorio) as IEEE does not use dedicated
% algorithm float types and packages that provide these will not provide
% correct IEEE style captions. The latest version and documentation of
% algorithmic.sty can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithms/
% There is also a support site at:
% http://algorithms.berlios.de/index.html
% Also of interest may be the (relatively newer and more customizable)
% algorithmicx.sty package by Szasz Janos:
% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithmicx/

% *** ALIGNMENT PACKAGES ***
%
%\usepackage{array}
% Frank Mittelbach's and David Carlisle's array.sty patches and improves
% the standard LaTeX2e array and tabular environments to provide better
% appearance and additional user controls. As the default LaTeX2e table
% generation code is lacking to the point of almost being broken with
% respect to the quality of the end results, all users are strongly
% advised to use an enhanced (at the very least that provided by array.sty)
% set of table tools. array.sty is already installed on most systems. The
% latest version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/required/tools/

%\usepackage{mdwmath}
%\usepackage{mdwtab}
% Also highly recommended is Mark Wooding's extremely powerful MDW tools,
% especially mdwmath.sty and mdwtab.sty which are used to format equations
% and tables, respectively. The MDWtools set is already installed on most
% LaTeX systems. The lastest version and documentation is available at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/mdwtools/

% IEEEtran contains the IEEEeqnarray family of commands that can be used to
% generate multiline equations as well as matrices, tables, etc., of high
% quality.

%\usepackage{eqparbox}
% Also of notable interest is Scott Pakin's eqparbox package for creating
% (automatically sized) equal width boxes - aka "natural width parboxes".
% Available at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/eqparbox/

% *** SUBFIGURE PACKAGES ***
%\usepackage[tight,footnotesize]{subfigure}
% subfigure.sty was written by Steven Douglas Cochran. This package makes it
% easy to put subfigures in your figures. e.g., "Figure 1a and 1b". For IEEE
% work, it is a good idea to load it with the tight package option to reduce
% the amount of white space around the subfigures. subfigure.sty is already
% installed on most LaTeX systems. The latest version and documentation can
% be obtained at:
% http://www.ctan.org/tex-archive/obsolete/macros/latex/contrib/subfigure/
% subfigure.sty has been superceeded by subfig.sty.

%\usepackage[caption=false]{caption}
%\usepackage[font=footnotesize]{subfig}
% subfig.sty, also written by Steven Douglas Cochran, is the modern
% replacement for subfigure.sty. However, subfig.sty requires and
% automatically loads Axel Sommerfeldt's caption.sty which will override
% IEEEtran.cls handling of captions and this will result in nonIEEE style
% figure/table captions. To prevent this problem, be sure and preload
% caption.sty with its "caption=false" package option. This is will preserve
% IEEEtran.cls handing of captions. Version 1.3 (2005/06/28) and later 
% (recommended due to many improvements over 1.2) of subfig.sty supports
% the caption=false option directly:
\usepackage[caption=false,font=footnotesize]{subfig}
%
% The latest version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/subfig/
% The latest version and documentation of caption.sty can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/caption/

% *** FLOAT PACKAGES ***
%
%\usepackage{fixltx2e}
% fixltx2e, the successor to the earlier fix2col.sty, was written by
% Frank Mittelbach and David Carlisle. This package corrects a few problems
% in the LaTeX2e kernel, the most notable of which is that in current
% LaTeX2e releases, the ordering of single and double column floats is not
% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
% single column figure to be placed prior to an earlier double column
% figure. The latest version and documentation can be found at:
% http://www.ctan.org/tex-archive/macros/latex/base/

%\usepackage{stfloats}
% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
% the ability to do double column floats at the bottom of the page as well
% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
% LaTeX2e). It also provides a command:
%\fnbelowfloat
% to enable the placement of footnotes below bottom floats (the standard
% LaTeX2e kernel puts them above bottom floats). This is an invasive package
% which rewrites many portions of the LaTeX2e float routines. It may not work
% with other packages that modify the LaTeX2e float routines. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/sttools/
% Documentation is contained in the stfloats.sty comments as well as in the
% presfull.pdf file. Do not use the stfloats baselinefloat ability as IEEE
% does not allow \baselineskip to stretch. Authors submitting work to the
% IEEE should note that IEEE rarely uses double column equations and
% that authors should try to avoid such use. Do not be tempted to use the
% cuted.sty or midfloat.sty packages (also by Sigitas Tolusis) as IEEE does
% not format its papers in such ways.

% *** PDF, URL AND HYPERLINK PACKAGES ***
%
%\usepackage{url}
% url.sty was written by Donald Arseneau. It provides better support for
% handling and breaking URLs. url.sty is already installed on most LaTeX
% systems. The latest version can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/misc/
% Read the url.sty source comments for usage information. Basically,
% \url{my_url_here}.

% *** Do not adjust lengths that control margins, column widths, etc. ***
% *** Do not use packages that alter fonts (such as pslatex).         ***
% There should be no need to do such things with IEEEtran.cls V1.6 and later.
% (Unless specifically asked to do so by the journal or conference you plan
% to submit to, of course. )

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{Title goes here}

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
\author{\IEEEauthorblockN{Servesh Muralidharan}
\IEEEauthorblockA{School of Computer Science and\\Statistics\\
Trinity College Dublin\\
Dublin 2, Ireland\\
Email: muralis@tcd.ie}
\and
\IEEEauthorblockN{David Gregg}
\IEEEauthorblockA{School of Computer Science and\\Statistics\\
Trinity College Dublin\\
Dublin 2, Ireland\\
Email: David.Gregg@cs.tcd.ie}}

% conference papers do not typically use \thanks and this command
% is locked out in conference mode. If really needed, such as for
% the acknowledgment of grants, issue a \IEEEoverridecommandlockouts
% after \documentclass

% for over three affiliations, or if they all won't fit within the width
% of the page, use this alternative format:
% 
%\author{\IEEEauthorblockN{Michael Shell\IEEEauthorrefmark{1},
%Homer Simpson\IEEEauthorrefmark{2},
%James Kirk\IEEEauthorrefmark{3}, 
%Montgomery Scott\IEEEauthorrefmark{3} and
%Eldon Tyrell\IEEEauthorrefmark{4}}
%\IEEEauthorblockA{\IEEEauthorrefmark{1}School of Electrical and Computer Engineering\\
%Georgia Institute of Technology,
%Atlanta, Georgia 30332--0250\\ Email: see http://www.michaelshell.org/contact.html}
%\IEEEauthorblockA{\IEEEauthorrefmark{2}Twentieth Century Fox, Springfield, USA\\
%Email: homer@thesimpsons.com}
%\IEEEauthorblockA{\IEEEauthorrefmark{3}Starfleet Academy, San Francisco, California 96678-2391\\
%Telephone: (800) 555--1212, Fax: (888) 555--1212}
%\IEEEauthorblockA{\IEEEauthorrefmark{4}Tyrell Inc., 123 Replicant Street, Los Angeles, California 90210--4321}}

% use for special paper notices
%\IEEEspecialpapernotice{(Invited Paper)}

% make the title area
\maketitle

\begin{abstract}
%\boldmath
the abstract goes here.

\end{abstract}
% IEEEtran.cls defaults to using nonbold math in the Abstract.
% This preserves the distinction between vectors and scalars. However,
% if the conference you are submitting to favors bold math in the abstract,
% then you can use LaTeX's standard command \boldmath at the very start
% of the abstract to achieve this. Many IEEE journals/conferences frown on
% math in the abstract anyway.

% no keywords

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

\section{Motivation}
In the continued race for improvement of multi core and many core architectures, computer networks are becoming a fundamental area for performance bottleneck especially in systems which are used for distributed processing of parallel applications. In order to handle the requirements posed by these applications the interconnects are becoming more and more complex. Such systems utilize a wide range of custom interconnects ranging from infiniband, myrinet to the more readily available ones such as 1Gbe and 10Gbe Ethernet standards and sophisticated routers and switches to handle the added challenge in routing data. While a small group of the specialized systems employ custom routers and topology configurations the majority of them especially those that use 1Gbe or 10Gbe use the commercial routers. In addition the growing complexity of routing protocols has resulted in advanced switches such as OpenVSwitch \cite{openvswitch}, etc., has resulted in these routers being upgraded frequently to keep up with the growing requirements. Software routers such as click modular router have been gaining popularity due to their flexible nature and cheaper alternative to the more expensive commercial routers and are widely used to prototype newer standards. In using software routers in the above scenario it would be possible to not only benefit from the newer developments in switching and routing protocols but it would require little effort to implement it in existing systems. One might argue that it would be a challenge to handle such complex workloads in software routers but with the presence of multicore and many core architectures and stream processors it is feasible to parallelize the packet processing workloads to improve the overall performance of the system. Several research projects are working on improving the performance by parallelizing the packet processing tasks on these architectures \cite{routebricks} \cite{packetshader}.

In a computing cluster which is used in the execution of parallel applications it is common to dedicate a core or processor in each node for operating system related tasks in order to minimize its effects on the application. Similarly, it should also be possible to utilize an additional core or processor to perform packet processing operations. While this is would be a challenging scenario it does provide the added benefits of a cheaper alternative to commercial routers. Also in bringing the network system into a software controlled domain it should be possible to perform additional optimizations from the application such as prioritized data flow, intelligent routing, etc., that would otherwise be difficult to implement due to the abstraction of the network system.

The focus of this paper is the development of a parallel processing engine that could be integrated with the application and can perform both the actual computations and the packet processing related tasks in parallel. The entire system can be divided into key parts as follows,

\begin{itemize}

\item A suitable methodology by which we can access the data in the userspace without interacting with the operating system. To do this (1)We need to read the data from the Network Interface Controller (NIC) before the OS and (2)Using this, establish a efficient manner for sending and receiving the data.(Section \ref{sec:userspace-packet})

\item Given a graph that represents the workload to be done on the packet flow, we divide it into tasks that can be executed in parallel and determine the regions that could be used to extract data and task parallelism.(Section \ref{sec:workload})

\item The parallel processing engine that utilizes the above and does the actual computations on the multi core or many core processor. It also handles load balancing and scheduling of the work across the available computing resources. (Section \ref{sec:ppe})

\end{itemize}

We concentrate more on the framework that can integrate the packet processing and application tasks into a parallel processing engine and the extraction of parallelism from the packet processing related tasks to improve the overall performance. Given an application graph and its partition consisting of parallel tasks we then combine it with the packet processing operations. In addition, Section \ref{sec:} we show the challenges involved in using a system consisting of threads and process that are used to actually perform the work. The extent to which we extract parallelism ranges not only to the packet tasks but to the low level access of data I/O to the network interface hardware through certain additions to the netmap APIs. Finally in Section \ref{sec: } we present certain pseudo applications that involves operations such as compression and encryption to evaluate our parallel processing engine.

\section{Background}

Let us consider a simple file storage system that is based in a data center. The application handles sensitive information by performing encryption on the data that matches predefined rules like file names, access level of user, etc., otherwise the default operation is to perform compression and store the information in the location specified. For the purpose of this example let us assume that the filesystem is located on a NFS server and it is responsible for concurrent operations on the data. A group of servers act as the interface to the storage system and the requests are evenly distributed over these nodes. The important operations that these nodes perform on the data are represented in the stream graph shown in figure \ref{fig1}. In the data retrieval part the information fetched from the filesystem is checked to see if it matches a set of predefined rules to check if it has to be encrypted otherwise passes it to the compression operation. During data storage the input data is checked whether it has to be decrypted or decompressed and then stored on to the filesystem.

\begin{figure}[ht]
\centering
\subfigure
{
	\includegraphics[width=3.5in]{ip-data-proc}
	\label{fig1.1}
}

\subfigure
{
	\includegraphics[width=3in]{op-data-proc}
	\label{fig1.2}
}
\caption{Operations performed on the data stored and retrieved}
\label{fig1}
\end{figure}

The nodes that are connected in the cluster are linked through a switch or router. In such a scenario the number of routers increase with the size of the cluster and the bandwidth required by the application. Due to this the performance of the application is directly dependent on the network especially the functionality of the routers. Moreover with a fixed infrastructure such as this it is not only difficult to upgrade the routers but it would also be a expensive proposition. In such a situation it makes more sense to use software routers which are much more flexible and cheaper and could prove to be a more effective than or in addition to commercial switches or routers.

In implementing the software routers if one were to dedicate several nodes for this purpose it might prove to be inefficient use of resources since the network traffic generated is highly dependent on the applications being executed in such a closed environment. Moreover such systems might act as a communication bottleneck and the routing operations are focused on these nodes. A more effective approach would be to bring the software routers into the nodes that are used for executing the applications itself. In doing this we dedicate a specific set of resources of each node for the packet processing operations and in essence make each node act as a router. All the nodes in such a cluster are connected with each other based on network topology suitable for its requirements. This would make it possible to divide the workload generated due to the network operations across all the nodes in the cluster. In this paper we propose such a system that not only brings the software routers as separate entity into the actual application nodes but takes it a step further and integrates the network operations with the application itself.

There are several challenges that arises due to this, mainly how toe tackle such large number of computations both from the application and the network effectively. Two important constraints would enable us to satisfy the requirements posed by such a system. First, the operations of both the application and the network is parallel in nature and second, the presence of co-processors \cite{coproc}, network processors \cite{NP}, Graphics Processing Units ( GPUs ) \& General Purpose GPUs ( GPGPUs ) \cite{gpu} has dramatically improved the capability of commodity hardware. With this theoretically, we should be able to meet the computational demand but if such a system were to succeed an efficient parallel processing engine is crucial.\\

\textit{write a para forward referencing other sections}\\

An important part of our system is the tasks related to packets such as routing, forwarding, etc., Since our design focuses more on the aspects of the parallel framework we utilize the click modular router for the packet processing tasks. Click modular router \cite{click} is a commonly used software router that includes a library of elements which is used in building flexible and reconfigurable routers. It offers a platform that is flexible for rapid development of newer network protocols and achieves good performance. An improved version known as SMP Click \cite{smpclick} also supports thread based parallel processing of network related tasks. Until recently click's performance in the user space was not as good as its kernel space equivalent. This is mainly due to its extensive use of polling driver in the kernel space. Netmap I/O \cite{netmap} provides a set of APIs that could directly access the packets from the NIC and it helped in the recent improvement of click's user space performance. We found the simpler construction of netmap easier to adapt to our framework  compared to other user space packet access APIs.

Click dynamically constructs routers in which the configuration file is passed as a parameter for the kernel module or user space binary. As a result most of the actual components are dynamically allocated and connected at run time. While this may be an advantage considering the alternate is to recompile the linux kernel, it does act as significant bottleneck in the case of user space. Moreover some of the restrictions on how the configuration can be parallelized and limits the actual performance that could be achieved in combination with netmap. By integrating the ability to access click's elements within our parallel processing engine we were able to scale significantly better. This does require a limited amount of modifications to some of click's libraries and elements. In generating the actual code statically, compilers such as gnu c, etc., can perform optimizations such as loop fusions, in-lining, dead code optimization etc., which would not have been possible previously.

\section{Design Principles}

The design of the parallel processing engine that integrates packet processing tasks into the application requires overcoming several challenges. The entire system is driven by the extent to which parallelism can be extracted. Some of the techniques that can be applied to the packet processing tasks can be used for the tasks related to the application. Even though we integrate the system with the application the techniques for extracting parallelism from the application is not in scope of this paper. We focus more on the methodologies related to the packet processing tasks and the parallel processing engine itself. The application is considered to be in a streaming model and the tasks corresponding to the various functions are considered to be parallel in nature. Below, we describe the various principles that are required for the conception of such a system.

\subsection{Userspace Packet I/O}

For a packet processing system that can be integrated with the application it is necessary to have an efficient manner for accessing the data from the NIC without interacting with the operating system. Moreover added flexibility and rapid data access here is crucial for the extraction of data parallelism at later stages. In using the techniques described below we exploit the modern NIC architecture to its full potential.

\subsubsection{Handling Hardware Queues}

The development of virtualization and improved network flow handling has lead to the hardware level multi queue support in the NICs. During transmission concurrent writes on to the different queues is possible, enabling multiple process to send data simultaneously. During reception the NIC classifies the packet on to one of the receive queues by a technique known as Receive Side Scaling (RSS). It uses the header information or tuples which specifies the queue the packet would be received at. The multi queue support helps in handling several streams of data in parallel and it is exploited extensively in our framework to process data concurrently. By offloading the classification of packets to the NIC, it removes any overhead associated in doing it in the application. We use these hardware queues as a means of not only accessing the data in parallel but essentially balance the workload across the different threads. This is done by controlling the number of queues that each thread handles based on the rate at which packets are received in each of the queues. It is also essential to have different threads accessing the transmit and receive queues to maximize concurrent access to data.

\subsubsection{Flow based Batching of Packets}

Packet batching is a common method used in boosting the performance of network applications. It involves combining several packets and performing operations on them together reducing the overhead involved in the I/O operations. In our framework we utilize batching but the number of packets that forms a batch is not a fixed number but is modified at run time to optimize processing rate. This is achieved by varying the number of packets that forms the batch based on the available threads that can handle the packets. The number of packets that we can operate together reflects the amount of data we are able to process and forms a important performance metric which is used later in the load balancing stage.

\subsubsection{Reduction of Per Packet System Calls}

System calls are required in order to synchronize data between the NIC hardware and the buffer memory used for intermediate storage of packets. Several work \cite{overhead} describe the overhead associated with per packet system calls and provide means of optimizing this. One of the common solution to this problem is issue system call to only a group of packets. To improve the performance of our framework we employ an \textit{adaptive scheme} based on both the rate of data flow and the type of application. The transmission and reception are handled differently in this scheme in order to optimize the usage of system calls in each of the case independently. During transmission we issue the synchronization call between the transmitted data and available buffer space in the NIC based on the rate at which we fill the buffer space. In doing this we can prevent degradation in latency for larger packets and improve performance for smaller packets given a bandwidth at which we transmit in both the cases. During receiving packets we try to read packets at the rate at which we are able to process them. If it is determined that we are trying to process at much higher rate than receiving packets we reduce the number of threads operating on the receive queues and dedicate these to processing tasks thereby reducing the number of system calls issued. If the rate at which we receive data increases we schedule more threads to handle it. In this manner we balance the system calls based on the data rate and application processing speed.

\subsection{Packet Processing Graph}

Understanding the set of packet processing tasks associated with the application is essential to performing them in parallel. We use a graph based representation of these tasks to determine the dependency involved for the different stages and determine the relation with that of the different tasks of the application. This graph contains the stages through which the data associated with the application is transformed into network packets and vice versa. By analyzing this graph it would make it possible in certain cases to even perform the operations related to the application directly on the packets itself. It also shows which of the tasks could be performed in parallel with that of the application and highlights those that have dependency on a specific task of the application. On analyzing these dependencies we are able to partition the packet processing graph into distinct set of tasks that can be used by our parallel processing engine to schedule on to available resources. This graph could also enable us to optimize based on the characteristics of the application or the architecture that is used for its execution.

\subsection{Application Task Graph}

Application task graph represents the operations of the associated in performing the computations. The specific subset of applications that we are interested in parallelizing are considered to be stream processing in nature. In such an application the communication between different tasks is predictable and there is tight correlation between the different stages of the application. In order to use the same parallel processing engine to handle both the packet processing and application tasks we represent the application tasks as a graph similar to the packet processing graph. An important difference is that while the packet processing tasks mainly operate on packets the application tasks are designed to operate on the data. For this paper since we focus more on the packet processing the optimizations in relation to the application graph are not discussed. Instead it is assumed that an optimal version of the application graph is provided and the focus is more on executing these tasks in parallel to that of the packet processing tasks. Similar to that of the packet processing graph we partition the application graph into finite task units that can be executed by our parallel processing engine. Also the tasks that can be operated on the packets directly are identified and are represented by relations between the application graph and the packet processing graph.

\subsection{Integration}

The integration stage is used in combining the application graph and the packet processing graph into a larger graph that unifies and represents the overall operations. The several dependency relations determined on the previous cases are used in establishing links between the graphs. We also determine the sequence of tasks that can be executed in parallel from both application task graph and packet processing graph. The combined set of tasks is then used by the parallel processing engine for execution. The sequence of tasks that can be executed in parallel and the data parallelism that can be extracted at the different stages is also determined at this stage.

\subsection{Parallel Processing Engine}

The parallel processing engine is the fundamental framework which handles the execution of the various tasks. It is constructed in a manner in which its flexible to utilize the resources efficiently. To do this it consists of both threads and process which are used to execute the workload. The overall group of tasks is partitioned or replicated across the different process and then each process schedules the specific workload using one of its worker threads. In this manner the workload is distributed uniformly and still retain fine grained control over the different execution elements. Two important aspects that the parallel processing engine must tackle is the scheduling of tasks and balancing the load across the available resources.

\subsubsection{Scheduling Tasks}

\subsubsection{Load Balancing}

\section{Implementation}

In this section we use the above design principles to construct the framework. We also discuss how we tackle some of the design challenges involved during the construction of the system.

\section{Old Background}

The constant advancement in server hardware has made systems with large number of hardware threads[] readily available. In addition to this the presence of co-processors[], network processors[], Graphics Processing Units ( GPUs ) \& General Purpose GPUs ( GPGPUs ) has dramatically improved the capability of commodity hardware. These developments has paved way to a platform with immense computing potential which existing applications already take advantage of. Along with the improvements in the Multi-Core \& Many-Core processor architecture the memory and I/O speeds have also seen significant improvement. In particular the network subsystem has seen significant growth, in combination with the improvements to optical communication it has lead to 10GbE interfaces being quite common in enterprise servers. While 40GbE and 100GbE are available for backbone infrastructure but would eventually follow into commodity hardware.

Network applications represent those applications that operate on any of the seven network layers described in the OSI Model. \{describe more here\}

Several previous work have discussed the process of parallelizing network applications[][][] to enable it to run faster in such multi threaded systems. The development of Software routers[][][] is encouraged further because it costs only a fraction of the commodity hardware routers. Moreover the advancements in the network interfaces has made faster interfaces readily in combination with the multi core servers the performance  It has been shown that such routers can be used to process packets at line rate [][][] which has been a challenge on its own. Network applications including those that operate on few or all of the network layers have been designed with sequential processing as the target environment. The shift to multi core architectures in recent times has induced such applications to utilize parallel processing as a alternate means of computing. While this could lead to significant improvement in performance it also suffers from the same problems as any parallel application. These include scalability, load-balancing, data dependency and the presence of non parallel regions in the program apart from those that might arise due to that specific application. The scaling of software routers has been discussed extensively in RouteBricks[] which shows the problems suffered by software routers when they are scaled beyond a single system. One could argue that complexity of network applications such as Deep Packet Inspection ( DPI ), Packet Forwarding, etc., could be handled by a single many socket Multi-Core SMP system. However, as the complexity of network applications develops it would be inevitable but to move to a multi node environment in order to perform the computations and to handle more network bandwidth.

Large scale computing clusters has been used in high-performance scientific application to solve problems and applications of very large size. It uses a cluster of compute nodes and distributes the workload across them in order to solve it faster. These systems utilize parallel programs consisting of both threads and MPI to break massive problems into smaller workload that can be executed in parallel. In addition, the presence of GPUs and GPGPUs requires a mixture of parallelization constructs such as OpenCL, CUDA, etc., in order to achieve the best performance in the system. In this paper we discuss how some of the techniques used by these applications could be used to parallelize network applications.

Click Modular Software Router[] consists of extensive elements which are used to construct software routers. Click primarily targets packet processing applications especially those that operate on L4 - L7. \{elaborate more here\}

Discuss about what each sections contains here.....

\section{Old Design}

This section describes the design principles related to our methodology. The objective is to determine the requirements necessary to parallelize network applications. Our methods focus on ways to improve the click system in some cases even extend its functionality.

\subsection{Emphasize on user level execution environment}

The existing software routers are predominantly implemented in the kernel space for the ease of access to resources and can bypass the OS overhead incurred when data is accessed through the network subsystem. While this alternative provides better performance it also leads to concerns regarding stability and security of the system. Newer techniques such as Netmap[] allows direct access to network buffers which makes it possible to write high performance packet processing applications in userspace itself. Typically network applications which operate on L4 - L7 are developed in the kernel space in order to intercept the network packets before it flows through the OS. Whereas those applications that utilize L1 - L3 are present in the user space. The objective here is to completely switch to user space to solve any of the stability or security concerns. In addition, scenarios where the applications that operate across most or all the layers could be improved by combining the operations across the different layers in user space itself.

\subsection{Eliminate the overhead associated with creation of dynamic objects}

Click constructs application graphs based on dynamic creation of objects and passing the data packets across the virtual functions. Though this is a very elegant approach in creating the software router it does incur the penalty of the virtual function tables. The click-tool[] addresses this issue but it does not solve the problem entirely. The construction of the software router at run time is the primary reason behind it. This mainly arises due to the presence of kernel module which installs the router as a kernel module based on the input configuration. However in the user level it would not be necessary for a separate process to initialize the router instead we could generate the entire program statically including the necessary flow associated with the specific application. This not only eliminates the overhead associated but also opens up means to optimize the graph further.

\subsection{Optimize based on application graph}

Network applications especially that operate on the lower layers are very sensitive to the overhead associated with operating on the data separately on each layer. Instead given an application graph if we were to provide a sequential flow of the different functions at the various layers then it would be possible for the compiler to optimize the regions of program into a single and optimized block of code. This would also improve the cache locality of the data accessed by consecutive functions. This also presents us with the ability to perform other types of stream graph optimizations[] which would not be possible if the specific graph was constructed dynamically at run time.

\subsection{Automated data parallelism}

Network traffic is a classic example by which data parallelism can be utilized to achieve scalable parallelism. Click's support for parallelism in user level is described as experimental, it is implemented by using a task graph consisting of each work associated with each element which is executed in the threads. The support for parallelism in the kernel mode is better in comparison to that of user mode. Since our target environment is in the user space the proposed design involves a different approach to which we can extract data parallelism. We first split the data across multiple queues and then replicate the application graph across to process the data the queues in parallel. An important condition here is to make sure packets of the same flow should pass through the same queue. Since the application graph can be replicated it also makes it possible to dynamically load balance several applications based on the workload.

\subsection{Efficient implementation of task parallelism}

In generating the application graph as a sequence of operations it also makes it possible to process different stages of the operation in a pipelined manner. This when combined with stream optimizations and could be used to determine which steps of the graph is best suited to form the stages of a pipeline. The condition to balance here is the weight associated with computation in that stage with that of the overhead incurred in computing the data in two separate stages. It would be necessary to determine whether the additional over head involved in executing the task in parallel provides any significant improvement in performance.

\section{Old Implementation}

Click was designed as a means to easily implement a software router on top of a Linux OS. This is implemented in two different ways one is by installing a kernel module and the other is a program that executes in the user level. Both of them accept a configuration file described in click's language specification. This is then used to construct the router by linking the virtual functions for the push and pull operations of the different elements in accordance with the specified graph.

\subsection{Modifications to click system}

\subsubsection{Static Generation Of ???}
The first step in the design was to overcome the limitation of using click's infrastructure to construct the software router which consists of dynamically initialized objects and virtual functions. This was necessary due to two important conditions, one was to eliminate the overhead associated with the dynamic construction of the objects and virtual functions and the other was to provide the compiler a chance to optimize the program during the compilation phase. This would however result in not being able to construct the software router using click's infrastructure but would require the application specific code to be recompiled. This could be a significant hurdle in the case of running the software router in kernel mode where the router configuration is installed by the click kernel module in which case the entire kernel module should be recompiled and reinitialized. Whereas in user space this would not be an issue since routers of different configurations are basically treated as separate applications and would just have to go through an additional compilation phase before being used. The added advantage of this would be to enable network applications that operate on L1 - L3 layers to use the packet processing ( L4 - L7 ) infrastructure provided by click elements.

\subsubsection{Implementation of data parallelism}

--Talk about MPI \& threads\\

--Our goal here is to focus on the network application as a workload and characterize it as such without involving any kind of overhead that might occur due to the network.\\

\section{Experimental Evaluation}

\section{Related Work}

\section{Conclusions}

% An example of a floating figure using the graphicx package.
% Note that \label must occur AFTER (or within) \caption.
% For figures, \caption should occur after the \includegraphics.
% Note that IEEEtran v1.7 and later has special internal code that
% is designed to preserve the operation of \label within \caption
% even when the captionsoff option is in effect. However, because
% of issues like this, it may be the safest practice to put all your
% \label just after \caption rather than within \caption{}.
%
% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
% option should be used if it is desired that the figures are to be
% displayed while in draft mode.
%
%\begin{figure}[!t]
%\centering
%\includegraphics[width=2.5in]{myfigure}
% where an .eps filename suffix will be assumed under latex, 
% and a .pdf suffix will be assumed for pdflatex; or what has been declared
% via \DeclareGraphicsExtensions.
%\caption{Simulation Results}
%\label{fig_sim}
%\end{figure}

% Note that IEEE typically puts floats only at the top, even when this
% results in a large percentage of a column being occupied by floats.

% An example of a double column floating figure using two subfigures.
% (The subfig.sty package must be loaded for this to work.)
% The subfigure \label commands are set within each subfloat command, the
% \label for the overall figure must come after \caption.
% \hfil must be used as a separator to get equal spacing.
% The subfigure.sty package works much the same way, except \subfigure is
% used instead of \subfloat.
%
%\begin{figure*}[!t]
%\centerline{\subfloat[Case I]\includegraphics[width=2.5in]{subfigcase1}%
%\label{fig_first_case}}
%\hfil
%\subfloat[Case II]{\includegraphics[width=2.5in]{subfigcase2}%
%\label{fig_second_case}}}
%\caption{Simulation results}
%\label{fig_sim}
%\end{figure*}
%
% Note that often IEEE papers with subfigures do not employ subfigure
% captions (using the optional argument to \subfloat), but instead will
% reference/describe all of them (a), (b), etc., within the main caption.

% An example of a floating table. Note that, for IEEE style tables, the 
% \caption command should come BEFORE the table. Table text will default to
% \footnotesize as IEEE normally uses this smaller font for tables.
% The \label must come after \caption as always.
%
%\begin{table}[!t]
%% increase table row spacing, adjust to taste
%\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
%\caption{An Example of a Table}
%\label{table_example}
%\centering
%% Some packages, such as MDW tools, offer better commands for making tables
%% than the plain LaTeX2e tabular which is used here.
%\begin{tabular}{|c||c|}
%\hline
%One & Two\\
%\hline
%Three & Four\\
%\hline
%\end{tabular}
%\end{table}

% Note that IEEE does not put floats in the very first column - or typically
% anywhere on the first page for that matter. Also, in-text middle ("here")
% positioning is not used. Most IEEE journals/conferences use top floats
% exclusively. Note that, LaTeX2e, unlike IEEE journals/conferences, places
% footnotes above bottom floats. This can be corrected via the \fnbelowfloat
% command of the stfloats package.

% conference papers do not normally have an appendix

% use section* for acknowledgement
\section*{Acknowledgment}
This work is funded by the IRCSET Enterprise Partnership Scheme along with IBM Research, Ireland.

% trigger a \newpage just before the given reference
% number - used to balance the columns on the last page
% adjust value as needed - may need to be readjusted if
% the document is modified later
%\IEEEtriggeratref{8}
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

% references section

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://www.ctan.org/tex-archive/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
%\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
%\bibliography{IEEEabrv,../bib/paper}
%
% <OR> manually copy in the resultant .bbl file
% set second argument of \begin to the number of references
% (used to reserve space for the reference number labels box)
%\begin{thebibliography}{1}

%\end{thebibliography}

% that's all folks
\end{document}

### 2013-05-10 paper.tex
final submission for ispa 2013

@@
 
-In this paper we propose a parallel processing framework for building stream applications that are bound by very large communication.
+In this paper we propose a parallel runtime framework for building stream applications that are bound by very large communication.
 A simple framework is used, where the filters are described as \textit{actors} and the application is represented by connecting such \textit{actors} into a directed acyclic graph (DAG).
@@
 In building a framework for such applications on general purpose computers, we can exploit two hardware features that could allow for dramatically improved performance, namely multicore processors and multi queue network controllers.
-Using the multiple queues makes it possible to do much of the communication in parallel and by using the multiple cores, the filters from the stream application can be executed concurrently.
+By using the multiple queues makes it possible to the communication in parallel and by using the multiple cores, the filters from the stream application can be executed concurrently.
 
@@
 However, constructing stream applications that use these features is difficult.
-Furthermore, many stream applications have a very specific communication pattern, typically operating on a \textit{flow}, a specific \textit{stream} between two filters.
+Furthermore, many stream applications have a very specific communication pattern, typically operating on a \textit{flow}, which is a specific \textit{stream} between two filters.
 Data in a \textit{stream} must be processed sequentially, whereas different \textit{flows} can often be processed in parallel.
 
-Newer network hardware has introduced a mechanism for separating arriving data into multiple queues, where data that belong to the same \textit{flow} are\comment{ guaranteed to be} placed in the same queue\cite{micro2008} \cite{intel2010}.
-We can utilize this feature to operate on multiple streams of data by processing each of the queues in parallel.
+Newer network hardware have introduced mechanisms for separating arriving data into multiple queues, where data that belong to the same \textit{flow} are\comment{ guaranteed to be} placed in the same queue\cite{micro2008} \cite{intel2010}.
+We can utilize this feature to operate on \textit{flows} by processing each of the queues in parallel.
 However, building applications that can use this feature directly is a challenging task because it requires explicit parallel programming and interaction with low-level network interfaces.
@@
 
-Our framework provides support for sending and receiving data in parallel over the NIC from user space and a common parallel execution engine that supports both application and packet processing tasks.
-
-\subsection{Parallel Packet I/O}
+Our framework provides support for sending and receiving data in parallel over the NIC from user space and a common parallel execution engine that supports both computation and communication filters.
+
+\subsection{Parallel communication interface}
 \label{ppio}
-The parallel packet I/O interface, when combined with the application tasks, provides concurrent communication for the application threads or processes over the network.
+The parallel communication interface, when combined with the computation filters, provides concurrent communication for the application threads or processes over the network.
 To do this, two important requirements have to be met.
 Firstly, access to network data from user space should be provided, which can be read from and written to in parallel.
-Secondly, it should be possible to integrate the packet operations with the application by providing the necessary interfaces for performing these packet operations, thereby allowing for some flexibility in the choice of communication protocol.
-
-\subsubsection{Userspace Packet Access}
-Applications executed in kernel space for accessing hardware devices directly would lead to system instability.
+Secondly, it should be possible to integrate this with the computation by providing the necessary communication filters that can perform packet operations.
+The ability to choose communication filters allows the flexibility in the choice of communication protocol.
+
+\subsubsection{Userspace data access}
+If applications are executed in kernel space for accessing data from network devices directly would lead to scenario where the operating system can become unstable.
 Instead, previous researchers \cite{Rizzo:2012:RNI:2090147.2103536}, \cite{1564468}, \cite{Krasnyansky}, etc., propose access to network buffers from user space.
@@
 
-\subsubsection{Packet Processing Operations}
+\subsubsection{Packet processing operations}
 
@@
 Since the application tasks are represented as a set of operations on a data stream, we include these operations as stages through which the data flows through to get transformed into data packets.
+This is done by representing these operations as communication filters pre-built into our runtime system.
 Integrating them with the application task graph makes it easier to extract parallelism later in the execution engine.
@@
 Previous research \cite{Salehi:1996:EAS:234766.234769}\cite{1409136} highlights the importance of affinity in network sensitive applications.
-Next, by utilizing our \textit{Parallel Packet I/O}, we specify the hardware queues each of these processes handle.
+Next, by utilizing our \textit{Parallel communication interface}~(\ref{ppio}), we specify the hardware queues each of these processes handle.
 \comment{Initially we divide the queues evenly among the different processes we create, but later during execution we could modify the allocation based on the application operations scheduled on them.
@@
 
-\subsubsection{Parallel task processing}
+\subsubsection{Parallel processing of computation filters}
 \label{ppt}
@@
 This ensures operations on the same \textit{flow} are performed on the same CPU, thereby using local cache.
-By making sure that packets belonging to a specific f\textit{flow} always end up in the same execution engine, we only perform operations conforming to that stream of data.
+By making sure that data belonging to a specific \textit{flow} always end up in the same execution engine, we only perform operations conforming to that stream of data.
 \comment{We use the ability of modern NICs to classify packets onto different queues, based on their attributes.}
@@
 Instead of performing the communication operations separately in the OS, by integrating them within the stream application\comment{using Click's elements within our infrastructure,}, we are able to construct a unified stream graph that represents the entire application, including computation and communication filters.
-Using this, we are able to parallelize these stream applications and achieve speedups of more than a factor of eight in all the applications we tested.
+Using this, we are able to parallelize the stream applications and achieve speedups of more than a factor of eight in all the applications we tested.
 
 By integrating these features into a runtime framework, the system of multiple queues, parallel communication and computation are entirely hidden from the programmer, who merely specifies a standard stream graph for computation alone.
-This frees up the programmer from the onerous task of dealing with the substantial communication requirements that result from stream application distribution.
+This frees the programmer from the onerous task of dealing with the implementation of the substantial communication requirements that result from the distribution of stream applications across multiple systems.
 
